### 2017-02-19 paper/paper.tex
added paper latex doc

\documentclass[]{article}
\usepackage[hidelinks]{hyperref}
%opening
\title{Comparative Evaluation of Spark and Flink Stream Processing}

\begin{document}

\maketitle
\newpage
\tableofcontents

\newpage

\begin{abstract}

\end{abstract}
Recent years have witnessed a big development improvements of Big Data frameworks such as Apache Spark, Apache Flink. For example, both frameworks support real-time
streaming processing. While which platform is the best still an open question. 
In this paper, we provide a performance comparison of streaming data processing in Spark and Flink, by measuring different performance metrics, namely latency and throughput. The experiment is done using multiple streaming workloads over real-world datasets.
\section{Introduction}

\par Big Data is now one of the hot topic in Information Technology, it has gained increasing attention of professionals in industry and academics, since the amount of data we generate is increasing. Moreover, recent years showed a rapid development of Big Data frameworks such Hadoop, Spark, Flink.    
 \par The Big Data is a method to process not just only just large datasets (volume), but also that arrives in high rate (velocity), and it comes in all kind of formats (variety)\cite{svs}.

\par The purpose of this paper is to provide a comparative experimental evaluation of throughput, latency of stream processing in Spark and Flink. We use a dataset of airplanes trajectories provided in
the context of Datacron project\footnote{\url{http://www.datacron\-project.eu/}}, in order to simulate real-world use cases. Using multiple streaming data workloads we try to cover the main API differences between the both frameworks.

\section{Technical Background}
This section presents the main characteristics, APIs, and main data abstraction of Spark and Flink.

\section{Experiment Setup and Implementation}
This section describes the experiment workloads and their implementation in each framework.
\section{Performance Results}

This section provides and compares the performance results of the different workloads.

\section{Conclusion}
This section gives the overall conclusion.

\begin{thebibliography}{[MT1]}
%
\bibitem[1]{svs} 
Laney, D 2001 3D Data Management: Controlling Data Volume, Velocity, and Variety. META Group.

%
\end{thebibliography}
\end{document}

### 2017-02-19 flink_project/src/main/java/kafka/datacorn/DatacornKafkaStreamProducer.java

@@ -82,7 +82,7 @@ public class DatacornKafkaStreamProducer {
 		/**
 		 * Generates a new random line every LINE_SLIDE_TIME_MS
 		 */
-		private static final int LINE_SLIDE_TIME_MS = 1000;
+		private static final int LINE_SLIDE_TIME_MS = 2*1000;
 
 		private static final long serialVersionUID = 2174904787118597072L;
 
@@ -100,6 +100,13 @@ public class DatacornKafkaStreamProducer {
 						line = new StreamRecord(line).toString();
 						ctx.collect(line);
 						Thread.sleep(LINE_SLIDE_TIME_MS);
+						ctx.collect(line);
+
+						Thread.sleep(LINE_SLIDE_TIME_MS);
+
+						ctx.collect(line);
+						Thread.sleep(LINE_SLIDE_TIME_MS);
+						
 					}
 				} catch (Exception e) {
 

### 2017-02-19 paper/paper.tex
added paper orga paragraph

@@
 Moreover, recent years showed a rapid development of Big Data frameworks such Hadoop, Spark, Flink.
- \par The Big Data is a method to process not just only just large datasets (volume), but also that arrives in high rate (velocity), and it comes in all kind of formats (variety)\cite{svs}.
+ \par The Big Data is a method to process not just only just large datasets (volume), but also that arrives in high rate (velocity), and it comes in all kind of formats (variety) \cite{svs}.
 
-\par The purpose of this paper is to provide a comparative experimental evaluation of throughput, latency of stream processing in Spark and Flink.
-We use a dataset of airplanes trajectories provided in
-the context of Datacron project\footnote{\url{http://www.datacron\-project.eu/}}, in order to simulate real-world use cases.
-Using multiple streaming data workloads we try to cover the main API differences between the both frameworks.
+\par The purpose of this paper is to provide a comparative experimental evaluation of throughput, latency (i.e., performance) of stream processing in Spark and Flink.
+We use a dataset of airplanes trajectories provided in the context of Datacron project\footnote{\url{http://www.datacron\-project.eu/}}, in order to simulate real-world use cases.
+Using multiple streaming data workloads we try to cover the main programming API differences between the both frameworks.
+Informally,  this work aims to help developer to determine which platform to chose in production.
+
+\par The rest of this paper is organized as follows: Section 2
+presents the main characteristics, APIs, and main data abstraction of Spark and Flink. Section 3  describes the experiment workloads and their implementation in each framework. Section 4
+provides and compares the performance results of the different workloads. Finally, Section 5 gives the overall conclusion.
 

### 2017-02-19 paper/paper.tex
first version of introduction section

@@
 \end{abstract}
-Recent years have witnessed a big development improvements of Big Data frameworks such as Apache Spark, Apache Flink. For example, both frameworks support real-time
+Recent years have witnessed a big development improvements of Big Data frameworks such as Apache Spark and Apache Flink. For example, both frameworks support real-time
 streaming processing. While which platform is the best still an open question. 
@@
 
-\par Big Data is now one of the hot topic in Information Technology, it has gained increasing attention of professionals in industry and academics, since the amount of data we generate is increasing.
-Moreover, recent years showed a rapid development of Big Data frameworks such Hadoop, Spark, Flink.
- \par The Big Data is a method to process not just only just large datasets (volume), but also that arrives in high rate (velocity), and it comes in all kind of formats (variety) \cite{svs}.
+\par Big Data is now one of the hot topics in Information Technology, it has gained increasing attention of professionals in industry and academics, since the amount of data we generate is increasing exponentially.
+Moreover, recent years showed a rapid development of Big Data frameworks such as Spark and Flink that provide more and efficient data processing capabilities.
+ \par The Big Data is a method to process not just only just large datasets (volume), but also that arrives in high rate (velocity), and it comes in all kinds of data format (variety) (i.e., 3Vs of Big Data (Volume, Velocity, Variety))\cite{svs}.
 
-\par The purpose of this paper is to provide a comparative experimental evaluation of throughput, latency (i.e., performance) of stream processing in Spark and Flink.
-We use a dataset of airplanes trajectories provided in the context of Datacron project\footnote{\url{http://www.datacron\-project.eu/}}, in order to simulate real-world use cases.
+\par The processing models of Big Data are Batch and Stream processing, Batch processing is a way to handle large volumes of data (e.g., processing of historical transactions), is more concerned with throughput.
+While the Stream processing  is to manages fast and continues data (real-time) and processes it as it arrives (e.g., fraud detection and network monitoring applications), where the low latency is required.
+
+\par The purpose of this paper is to provide a comparative experimental evaluation of throughput, latency (i.e., performance) of stream processing in Apache Spark and Apache Flink.
+We use a dataset of airplanes trajectories provided in the context of Datacron project\footnote{\url{http://www.datacron-project.eu/}}, in order to simulate real-world use cases.
 Using multiple streaming data workloads we try to cover the main programming API differences between the both frameworks.
-Informally,  this work aims to help developer to determine which platform to chose in production.
+Informally,  this work aims to help developer to determine which platform to choose in production.
 

### 2017-02-19 paper/paper.tex
first version of introduction section

@@
 
-\par Big Data is now one of the hot topics in Information Technology, it has gained increasing attention of professionals in industry and academics, since the amount of data we generate is increasing exponentially.
+\par Big Data is now one of the hot topics in Information Technology, it has gained increasing attention of professionals in industry and academics, since the amount of data we generate is increasing exponentially; generated from social media platforms, mobile phones, IoT and finical transactions.
 Moreover, recent years showed a rapid development of Big Data frameworks such as Spark and Flink that provide more and efficient data processing capabilities.
@@
 \par The processing models of Big Data are Batch and Stream processing, Batch processing is a way to handle large volumes of data (e.g., processing of historical transactions), is more concerned with throughput.
-While the Stream processing  is to manages fast and continues data (real-time) and processes it as it arrives (e.g., fraud detection and network monitoring applications), where the low latency is required.
+While the Stream processing  is to manages fast and continuously incoming data (real-time), processes it as it arrives (e.g., credit card fraud detection and network monitoring applications), where the low latency is required.
 

### 2017-02-19 paper/paper.tex
added spark section

@@
 \par The processing models of Big Data are Batch and Stream processing, Batch processing is a way to handle large volumes of data (e.g., processing of historical transactions), is more concerned with throughput.
-While the Stream processing  is to manages fast and continuously incoming data (real-time), processes it as it arrives (e.g., credit card fraud detection and network monitoring applications), where the low latency is required.
+While the Stream processing  is a method to manage fast and continuously incoming data (real-time), process it as it arrives (e.g., credit card fraud detection and network monitoring applications), where the low latency is required.
+The variety of data (structured and semi-structured) is included in the two processing models.
 
 \par The purpose of this paper is to provide a comparative experimental evaluation of throughput, latency (i.e., performance) of stream processing in Apache Spark and Apache Flink.
-We use a dataset of airplanes trajectories provided in the context of Datacron project\footnote{\url{http://www.datacron-project.eu/}}, in order to simulate real-world use cases.
+We use a datasets of airplanes trajectories provided in the context of Datacron project\footnote{\url{http://www.datacron-project.eu/}}, in order to simulate real-world use cases.
 Using multiple streaming data workloads we try to cover the main programming API differences between the both frameworks.
@@
 \section{Technical Background}
-This section presents the main characteristics, APIs, and main data abstraction of Spark and Flink.
+ In this section, we present the architecture, key characteristics, programming APIs  of Apache Spark, Apache Flink. The Apache Kafka platform is also discussed briefly.
 
+\subsection{Apache Spark}
+
+Apache Spark is an open source project that provide general framework for large-scale data processing \cite{spark}.
+It offers programming API in Java, Scala, Python and R.
+Its stack includes set of built-in modules including Spark SQL, MLlib for machine learning, GraphX for graph processing, and Spark Streaming to process real-time data streams.
+It can access different data sources including Hadoop Distributed File System (HDFS), Apache Cassandra database, Apache HBase database etc.
+
+The main data abstraction in Spark is the Resilient Distributed Datasets (RDDs), which is a in-memory collection of elements partitioned across cluster of computers that can be processed in parallel \cite{rdd}.
+RDDs supports two categories of operations: transformations that drive new RDDs from the operated ones (e.g., map).
+And actions which return a value (e.g., count).
 \section{Experiment Setup and Implementation}
@@
 
+\bibitem[2]{spark} 
+Apache Spark. Available: https://spark.apache.org/ [Accessed February 2017]
+
+\bibitem[3]{rdd}
+Zaharia, Matei, et al.
+"Resilient distributed datasets: A fault-tolerant abstraction for in-memory cluster computing." Proceedings of the 9th USENIX conference on Networked Systems Design and Implementation.
+USENIX Association, 2012.
 %

### 2017-02-19 paper/paper.tex
added flink section

@@
 
-\par The processing models of Big Data are Batch and Stream processing, Batch processing is a way to handle large volumes of data (e.g., processing of historical transactions), is more concerned with throughput.
+\par The processing models of Big Data are Batch and Stream processing, Batch processing is a way to handle large, finite volumes of data (e.g., processing of historical transaction records), is more concerned with throughput.
 While the Stream processing  is a method to manage fast and continuously incoming data (real-time), process it as it arrives (e.g., credit card fraud detection and network monitoring applications), where the low latency is required.
@@
 
-Apache Spark is an open source project that provide general framework for large-scale data processing \cite{spark}.
+\par Apache Spark is an open source project that provide general framework for large-scale data processing \cite{spark}.
 It offers programming API in Java, Scala, Python and R.
@@
 
-The main data abstraction in Spark is the Resilient Distributed Datasets (RDDs), which is a in-memory collection of elements partitioned across cluster of computers that can be processed in parallel \cite{rdd}.
+\par The main data abstraction in Spark is the Resilient Distributed Datasets (RDDs), which is a in-memory collection of elements partitioned across cluster of computers that can be processed in parallel \cite{rdd}.
 RDDs supports two categories of operations: transformations that drive new RDDs from the operated ones (e.g., map).
 And actions which return a value (e.g., count).
+
+//TODO: DStreams
+
+\subsection{Apache Flink}
+
+Apache Flink is an open source project that provide large-scale, distributed stream processing platform, while the batch processing treated as special case of streaming (as finite data stream) \cite{flink}.
+It offers programming API in Java, Scala.
+Its stack includes the core DataStream and DataSet APIs with additional libraries such as Complex event processing for Flink (FlinkCEP), Machine Learning for Flink (FlinkML), and Flink Graph API (Gelly).
+
+The main data abstraction of Flink are DataStream and DataSet that represents read-only collection of elements of the same type.
+The list of elements is bounded (i.e., finite) in DataSet, while it is unbounded (i.e., infinite) in case of DataStream.
+
 \section{Experiment Setup and Implementation}
@@
 \bibitem[2]{spark} 
-Apache Spark. Available: https://spark.apache.org/ [Accessed February 2017]
+Apache Spark. Available: https://spark.apache.org/ [Accessed February 2017].
 
@@
 USENIX Association, 2012.
+
+\bibitem[4]{flink} 
+Apache Flink. Available: https://flink.apache.org/ [Accessed February 2017].
+
 %

### 2017-02-19 paper/paper.tex
added apache kafka

@@
 
-Apache Flink is an open source project that provide large-scale, distributed stream processing platform, while the batch processing treated as special case of streaming (as finite data stream) \cite{flink}.
+Apache Flink is an open source project that provide large-scale, distributed stream processing platform, while the batch processing treated as special case of streaming applications \cite{flink}.
 It offers programming API in Java, Scala.
@@
 The list of elements is bounded (i.e., finite) in DataSet, while it is unbounded (i.e., infinite) in case of DataStream.
+
+\subsection{Apache Kafka}
+Apache Kafka is scalable, fault-tolerant distributed publish/subscribe streaming framework \cite{kafka}.
+It manges the stream records in different categorizes (i.e., topics) that portioned and distributed over the servers in the Kafka cluster.
+It allows the data producers to publish stream of records to certain Kafka topic.
+While the consumer applications can subscribe to one or more topic to read the data stream.
 
@@
 
+\bibitem[5]{kafka} 
+Apache Kafka. Available: https://kafka.apache.org/ [Accessed February 2017].
+
 %

### 2017-02-20 paper/paper.tex
added data stream setup section

@@
 \usepackage[hidelinks]{hyperref}
+\usepackage{graphicx}
 %opening
@@
 While the consumer applications can subscribe to one or more topic to read the data stream.
+It's widely  adopted, for example,
+Spark and Flink can receive data stream from Kafka.
 
 \section{Experiment Setup and Implementation}
-This section describes the experiment workloads and their implementation in each framework.
+This section describes the data stream setup, design of evaluation streaming workloads, and implementation details in Spark and Flink.
+
+\subsection{Data Stream Setup}
+Our streaming workloads read input data stream from Kafka.
+In order to simulate real-world uses cases we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircrafts position.
+In our experiment a data stream producer component reads the ADS-B datasets, then publish it to Kafka cluster to be consumed by the workloads in Spark and Flink.
+Figure 1 illustrates the data producer and the Kafka cluster setup.
+The data stream is portioned into four portions over two server, while the data producer publish the stream records randomly to Kafka partitions.
+
+\begin{figure}[h]
+
+  \centering
+    \includegraphics[width=.9\textwidth, height=.2\textheight]{kafka.png}
+     \caption{Data Stream Producer \& Kafka Cluster Setup.}
+\end{figure}
+
 \section{Performance Results}

### 2017-02-20 paper/paper.tex
update first page style
added fisrt workload

@@
 %opening
-\title{Comparative Evaluation of Spark and Flink Stream Processing}
 
@@
 
-\maketitle
+\begin{flushleft}
+\centering\LARGE {\bf Comparative Evaluation of Spark and Flink Stream Processing}
+\thispagestyle{empty}
+\end{flushleft}
 \newpage
@@
 Our streaming workloads read input data stream from Kafka.
-In order to simulate real-world uses cases we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircrafts position.
-In our experiment a data stream producer component reads the ADS-B datasets, then publish it to Kafka cluster to be consumed by the workloads in Spark and Flink.
+In order to simulate real-world uses cases we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircrafts position over time, comprise 22 fields of data such as AircraftID, Date message generated, longitude, latitude, and altitude.
+\par In our experiment a data stream producer component reads the ADS-B datasets, then publish it to Kafka cluster to be consumed by the workloads in Spark and Flink.
 Figure 1 illustrates the data producer and the Kafka cluster setup.
@@
 
+\subsection{Workloads Design}
+In our experimental evaluation, we developed two streaming processing workloads that read the ADS-B messages stream from Kafka to perform basic trajectories analysis methods.
+In our design choice, we try to cover key main points related to the streaming processing, to evaluate how  does each framework perform.
+The following are general key aspects of  streaming data processing were covered by our workloads design:
+
+\begin{itemize}
+\item Handling parallel input streams.
+\item How to aggregate the state of input stream records.
+\item Mange the order of stream records.
+\item How to provide and update global data model in stream processing task.
+\item Evaluate the performance by measuring  the latency and throughput.
+\end{itemize}
+We give  in Section 3.3 how Spark and Flink are handling these key aspects and related issues, then we discuss the performance evaluation in Section 4.
+
+\subsubsection{Statistics Computation per Trajectory}
+In first stream processing workload, we perform statistics calculation for each trajectory (i.e.,  messages of the same ID) in ADS-B messages stream.
+As example of computed statistics indexes speed mean, mean of location coordinates, min and max altitude, etc.
+In this workload, we try to cover
+the parallel receiving of input data stream, state-full stream aggregation, and preserve the order of stream records.
+
+\subsection{Workloads Implementation}
 \section{Performance Results}

### 2017-02-20 paper/paper.tex
added spark & flink impl sections

@@
 \usepackage{graphicx}
+\usepackage[inline]{enumitem}
+\usepackage{listings}
 %opening
@@
 
-\par Big Data is now one of the hot topics in Information Technology, it has gained increasing attention of professionals in industry and academics, since the amount of data we generate is increasing exponentially; generated from social media platforms, mobile phones, IoT and finical transactions.
+\par Big Data is now one of the most appealing topics in Information Technology, it has gained increasing attention of professionals in industry and academics, since the amount of data we generate is increasing exponentially; generated from social media platforms, mobile phones, IoT and finical transactions.
 Moreover, recent years showed a rapid development of Big Data frameworks such as Spark and Flink that provide more and efficient data processing capabilities.
@@
 Our streaming workloads read input data stream from Kafka.
-In order to simulate real-world uses cases we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircrafts position over time, comprise 22 fields of data such as AircraftID, Date message generated, longitude, latitude, and altitude.
+In order to simulate real-world uses cases we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircrafts position over time, comprise 22 fields of data such as aircraft ID, date message generated, longitude, latitude, and altitude.
 \par In our experiment a data stream producer component reads the ADS-B datasets, then publish it to Kafka cluster to be consumed by the workloads in Spark and Flink.
@@
 \end{itemize}
-We give  in Section 3.3 how Spark and Flink are handling these key aspects and related issues, then we discuss the performance evaluation in Section 4.
+We present how Spark and Flink are handling these key aspects and related issues in each workload, then we discuss the performance evaluation in Section 4.
 
@@
 the parallel receiving of input data stream, state-full stream aggregation, and preserve the order of stream records.
+The implementation details of this workload in Spark and Flink, are presented in the following: 
+\begin{itemize}
+\item {\bf{Implementation in Spark }}
 
+The statistics calculation workload is implemented in Spark by performing the following steps: \begin{enumerate*}[label={\alph*)}]
+\item read parallel Kafka input streams 
+\item parse the multiple streams records   
+\item filter only the relevant  messages 
+\item combine the all messages related to the same trajectory
+\item manage the aggregate state and compute the statistics for each trajectory.
+\end{enumerate*}
+
+\par The used Spark Streaming API operations to execute those steps are shown in Figure 2.
+First, in Spark to consume parallel Kafka stream, multiple Kafka stream receivers must be created then combine them using the stream \texttt{union} API.
+Second, we parse each stream record (i.e., ADS-B message) then map it to tuple of aircraft ID and aircraft position object through \texttt{mapToPair} transformation API.
+Third, the irrelevant messages are filtered by the \texttt{filter}.
+Fourth, the tuples with same ID are grouped together using the \texttt{groupByKey} Spark API function, after applying this transformation the input stream records are shuffled which means to preserve the order of records a sort operation is needed.
+Finally, due to the mini-batch nature of Spark streaming, the input stream is processed batch by batch, we use the  \texttt{updateStateByKey} to compute statistics per trajectory after sorting the list of positions, then update the aggregated state of the trajectories stream.
+
+\begin{figure}[h]
+
+  \centering
+
+     \caption{Spark streaming implementation of statistics computation workload.}
+\end{figure}
+
+\item {\bf{Implementation in Flink }}
+
+The statistics computation of trajectories stream is solved in Flink by performing the following steps: \begin{enumerate*}[label={\alph*)}]
+\item read Kafka input stream
+\item parse the input stream records   
+\item filter only the relevant  messages 
+\item combine the all messages related to the same trajectory
+\item compute the statistics for each trajectory.
+\end{enumerate*}
+
+\par The utilized Flink APIs to perform those steps are shown in Figure 3.
+First, create Kafka stream consumers that run in multiple parallel instances based on the configured \texttt{parallelism} factor.
+Second,  stream elements (i.e., ADS-B message) are mapped  to tuples of aircraft ID and  position object through \texttt{map} transformation API.
+Third, the irrelevant messages are filtered by the \texttt{filter}.
+Fourth, tuples with same ID are combined together using the \texttt{keyBy} API function.
+Finally, the \texttt{reduce} API function is used to calculate the statistics and aggregate it.
+
+\begin{figure}[h]
+
+  \centering
+
+     \caption{Flink implementation of statistics computation workload.}
+\end{figure}
+
+//TODO: discuss differences between flink and spark.
+
+\end{itemize}
 \subsection{Workloads Implementation}

### 2017-02-21 paper/paper.tex
added sector change detetcion workload

@@
 
-\subsection{Workloads Design}
+\subsection{Evaluation Workloads}
 In our experimental evaluation, we developed two streaming processing workloads that read the ADS-B messages stream from Kafka to perform basic trajectories analysis methods.
@@
 \end{itemize}
-\subsection{Workloads Implementation}
+
+\subsubsection{Sector of Trajectory Change Detection}
+
+The goal of sector change detection workload is to detect the  enter or leaving of aircraft from one air sector to other, by processing the trajectories stream.
+Given that the dataset of sectors (i.e., polygons) available as reference to  check against it, also it used to assign the corresponding sector for certain aircraft position.
+In this workload, providing the  global model of data (sectors dataset) in streaming processing workload is tested.Moreover, preserving the stream records order and parallel streams consuming are covered.
+
 \section{Performance Results}

### 2017-02-21 paper/paper.tex
added spark impl of sector change detection

@@
 
-\subsection{Evaluation Workloads}
+\subsection{Evaluation Streaming Workloads}
 In our experimental evaluation, we developed two streaming processing workloads that read the ADS-B messages stream from Kafka to perform basic trajectories analysis methods.
@@
 the parallel receiving of input data stream, state-full stream aggregation, and preserve the order of stream records.
-The implementation details of this workload in Spark and Flink, are presented in the following: 
+The implementation details of this workload in Spark and Flink are presented in the following: 
 \begin{itemize}
@@
 
-//TODO: discuss differences between flink and spark.
+//TODO: discuss differences between Flink and Spark.
 
@@
 
+ The implementation details of this workload in Spark and Flink are presented in the following: 
+ \begin{itemize}
+ \item {\bf{Implementation in Spark }}
+
+ we implement this workload, based on the following operation stages: 
+ \begin{enumerate*}[label={\alph*)}]
+ \item receive and build the trajectories stream 
+ \item manage the aggregate state and assign the sector for each trajectory.
+ \item identify the  change of sector per trajectory if exist
+ \end{enumerate*}
+
+ figure 4 shows the Spark streaming APIs used to implement trajectory's sector change detection workload.
+First, the trajectories stream is built using \texttt{mapToPair}, \texttt{filter}, and \texttt{groupByKey} APIs as was discussed in statistics computation workload.
+Second,  \texttt{updateStateByKey} API is used to assign corresponding sector for all new aircraft positions in  micro batch  and aggregate last position in previous batch.
+Finally, the detection of sector change is done by extra filtering transformation using \texttt{filter} API, by identifying the change of sectors in each trajectory.
+ \begin{figure}[h]
+
+   \centering
+
+      \caption{Spark streaming implementation of sector change detection workload.}
+ \end{figure}
+
+ \item {\bf{Implementation in Flink }}
+ \end{itemize}
 \section{Performance Results}

### 2017-02-21 paper/paper.tex
added flink impl of sector change workload

@@
 
-//TODO: DStreams
+//TODO: DStreams, streaming model
 
@@
 While the consumer applications can subscribe to one or more topic to read the data stream.
-It's widely  adopted, for example,
+It has been widely  adopted, for example,
 Spark and Flink can receive data stream from Kafka.
@@
 
-\subsubsection{Sector of Trajectory Change Detection}
+\subsubsection{Air Sector Change Detection}
 
@@
 Finally, the detection of sector change is done by extra filtering transformation using \texttt{filter} API, by identifying the change of sectors in each trajectory.
+While the sector data set is provided using the \texttt{Broadcast} feature in Spark.
  \begin{figure}[h]
@@
  \item {\bf{Implementation in Flink }}
+
+ The Flink implementation of sector change detection workload, uses the \texttt{FlinkKafkaConsumer09} that read the ADS-B messages stream from Kafka, which used to construct \texttt{DataStream} instance.
+The \texttt{DataStream} records are parsed to tuples of aircraft ID and aircraft position, by a configured \texttt{map} function.
+The irrelevant tuples are filtered based on the message type using a \texttt{filter} operation.
+Afterward, the tuples with common IDs are grouped together by the \texttt{keyBy} operator.
+Since the Flink streaming processes the stream records as it come, a defined \texttt{reduce} operator assign corresponding sector of new tuple, and the old tuple with same ID is used to retrieve the previous sector to attach it to the new tuple, while the old tuple is discarded.
+Finally, the tuples with different previous and current sector that represent change in sector case, they are filtered using defined \texttt{filter} transformation.
  \end{itemize}

### 2017-02-21 paper/paper.tex
add TODO
added idc reference

@@
 
-\par Big Data is now one of the most appealing topics in Information Technology, it has gained increasing attention of professionals in industry and academics, since the amount of data we generate is increasing exponentially; generated from social media platforms, mobile phones, IoT and finical transactions.
+\par Big Data is now one of the most appealing topics in Information Technology, it has gained increasing attention of professionals in industry and academics, since the amount of data we generate is increasing exponentially \cite{idc} that generated from social media platforms, mobile phones, IoT and finical transactions, etc.
 Moreover, recent years showed a rapid development of Big Data frameworks such as Spark and Flink that provide more and efficient data processing capabilities.
@@
 
+\par TODO: add paragraph about Yahoo benchmark and why do we still need evaluation 1- continuous dev of frameworks 2- explore more complex operation like groupby that shuffle the data.
 \par The purpose of this paper is to provide a comparative experimental evaluation of throughput, latency (i.e., performance) of stream processing in Apache Spark and Apache Flink.
@@
 %
-\bibitem[1]{svs} 
+\bibitem[1]{idc} 
+Gantz, John, and David Reinsel. "The digital universe in 2020: Big data, bigger digital shadows, and biggest growth in the far east." IDC iView: IDC Analyze the future 2007.2012 (2012): 1-16.
+
+\bibitem[2]{svs} 
 Laney, D 2001 3D Data Management: Controlling Data Volume, Velocity, and Variety. META Group.
 
-\bibitem[2]{spark} 
+\bibitem[3]{spark} 
 Apache Spark. Available: https://spark.apache.org/ [Accessed February 2017].
 
-\bibitem[3]{rdd}
+\bibitem[4]{rdd}
 Zaharia, Matei, et al.
@@
 
-\bibitem[4]{flink} 
+\bibitem[5]{flink} 
 Apache Flink. Available: https://flink.apache.org/ [Accessed February 2017].
 
-\bibitem[5]{kafka} 
+\bibitem[6]{kafka} 
 Apache Kafka. Available: https://kafka.apache.org/ [Accessed February 2017].

### 2017-02-22 paper/paper.tex
fix name convention of trajectory to position message

@@
 Our streaming workloads read input data stream from Kafka.
-In order to simulate real-world uses cases we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircrafts position over time, comprise 22 fields of data such as aircraft ID, date message generated, longitude, latitude, and altitude.
+In order to simulate real-world uses cases we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircraft's position over time, comprise 22 fields of data such as aircraft ID, date message generated, longitude, latitude, and altitude.
 \par In our experiment a data stream producer component reads the ADS-B datasets, then publish it to Kafka cluster to be consumed by the workloads in Spark and Flink.
@@
 First, in Spark to consume parallel Kafka stream, multiple Kafka stream receivers must be created then combine them using the stream \texttt{union} API.
-Second, we parse each stream record (i.e., ADS-B message) then map it to tuple of aircraft ID and aircraft position object through \texttt{mapToPair} transformation API.
+Second, we parse each stream record (i.e., ADS-B message) then map it to tuple of aircraft ID and a position of aircraft object through \texttt{mapToPair} transformation API.
 Third, the irrelevant messages are filtered by the \texttt{filter}.
@@
 The goal of sector change detection workload is to detect the  enter or leaving of aircraft from one air sector to other, by processing the trajectories stream.
-Given that the dataset of sectors (i.e., polygons) available as reference to check against it, also it used to assign the corresponding sector for certain aircraft position.
+Given that the dataset of sectors (i.e., polygons) available as reference to check against it, also it used to assign the corresponding sector for certain aircraft's position.
 In this workload, providing the  global model of data (sectors dataset) in streaming processing workload is tested.Moreover, preserving the stream records order and parallel streams consuming are covered.
@@
  The Flink implementation of sector change detection workload, uses the \texttt{FlinkKafkaConsumer09} that read the ADS-B messages stream from Kafka, which used to construct \texttt{DataStream} instance.
-The \texttt{DataStream} records are parsed to tuples of aircraft ID and aircraft position, by a configured \texttt{map} function.
+The \texttt{DataStream} records are parsed to tuples of aircraft ID and aircraft's position, by a configured \texttt{map} function.
 The irrelevant tuples are filtered based on the message type using a \texttt{filter} operation.

### 2017-02-22 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/SectorUtils.java

@@ -3,7 +3,7 @@ package de.kdml.bigdatalab.spark_and_flink.common_utils;
 import java.util.List;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Trajectory;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 
 /**
  * 
@@ -20,7 +20,7 @@ public class SectorUtils {
 	 * @param sectors
 	 * @return
 	 */
-	public static Sector getSectorForTrajectory(Trajectory trajectory, List<Sector> sectors) {
+	public static Sector getSectorForTrajectory(PositionMessage trajectory, List<Sector> sectors) {
 
 		for (Sector sector : sectors) {
 

### 2017-02-22 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/StatisticsUtils.java

@@ -2,7 +2,7 @@ package de.kdml.bigdatalab.spark_and_flink.common_utils;
 
 import java.util.List;
 
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Trajectory;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.TrajectoryStatisticsWrapper;
 
 /**
@@ -19,15 +19,15 @@ public class StatisticsUtils {
 	 * @param trajectories
 	 * @return
 	 */
-	public static List<Trajectory> computeStatistics(List<Trajectory> trajectories) {
+	public static List<PositionMessage> computeStatistics(List<PositionMessage> trajectories) {
 
 		double minLongtitude = Double.MAX_VALUE, minLatitude = Double.MAX_VALUE, minAltitude = Double.MAX_VALUE,
 				minSpeed = Double.MAX_VALUE, minAcceleration = Double.MAX_VALUE;
 		double maxLongtitude = Double.MIN_VALUE, maxLatitude = Double.MIN_VALUE, maxAltitude = Double.MIN_VALUE,
 				maxSpeed = Double.MIN_VALUE, maxAcceleration = Double.MIN_VALUE;
 
-		Trajectory prevTrajectory = null;
-		for (Trajectory trajectory : trajectories) {
+		PositionMessage prevTrajectory = null;
+		for (PositionMessage trajectory : trajectories) {
 
 			TrajectoriesUtils.calculateDistanceAndSpeedOfTrajectory(prevTrajectory, trajectory);
 
@@ -65,7 +65,7 @@ public class StatisticsUtils {
 		statistics.setMaxSpeed(maxSpeed);
 		statistics.setMinSpeed(minSpeed);
 
-		for (Trajectory trajectory : trajectories) {
+		for (PositionMessage trajectory : trajectories) {
 
 			// just update new trajectories
 			if (trajectory.getStatistics() == null) {
@@ -84,7 +84,7 @@ public class StatisticsUtils {
 	 * @param oldTrajectory
 	 * @param newTrajectory
 	 */
-	public static void computeStatistics(Trajectory oldTrajectory, Trajectory newTrajectory) {
+	public static void computeStatistics(PositionMessage oldTrajectory, PositionMessage newTrajectory) {
 
 		if (oldTrajectory == null) {
[…]

### 2017-02-22 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/TrajectoriesUtils.java

@@ -8,7 +8,7 @@ import javax.swing.text.html.HTMLDocument.HTMLReader.SpecialAction;
 
 import com.google.common.collect.Lists;
 
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Trajectory;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 
 /***
  * 
@@ -27,11 +27,11 @@ public class TrajectoriesUtils {
 	 * @param line
 	 * @return
 	 */
-	public static Trajectory parseDataInput(String line) {
+	public static PositionMessage parseDataInput(String line) {
 
 		String[] attributes = line.split(",");
 
-		Trajectory trajectory = new Trajectory();
+		PositionMessage trajectory = new PositionMessage();
 		if (attributes.length > 5) {
 			trajectory.setType(attributes[0] + attributes[1]);
 			trajectory.setID(attributes[4]);
@@ -56,9 +56,9 @@ public class TrajectoriesUtils {
 	 * @param trajectories
 	 * @return
 	 */
-	public static List<Trajectory> sortTrajectories(List<Trajectory> trajectories) {
+	public static List<PositionMessage> sortTrajectories(List<PositionMessage> trajectories) {
 
-		Iterator<Trajectory> sortedTrajectories = trajectories.stream().sorted((trajectory1, trajector2) -> {
+		Iterator<PositionMessage> sortedTrajectories = trajectories.stream().sorted((trajectory1, trajector2) -> {
 
 			return -1 * trajectory1.getCreatedDateTime().compareTo(trajector2.getCreatedDateTime());
 		}).iterator();
@@ -73,7 +73,7 @@ public class TrajectoriesUtils {
 	 * @param prevTrajectory
 	 * @param trajectory
 	 */
-	public static void calculateDistanceAndSpeedOfTrajectory(Trajectory prevTrajectory, Trajectory trajectory) {
+	public static void calculateDistanceAndSpeedOfTrajectory(PositionMessage prevTrajectory, PositionMessage trajectory) {
 
 		if (trajectory.getSpeed() != null && trajectory.getDistance() != null) {
 			return;

### 2017-02-22 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/data/PositionMessage.java

@@ -3,7 +3,17 @@ package de.kdml.bigdatalab.spark_and_flink.common_utils.data;
 import java.io.Serializable;
 import java.time.LocalDateTime;
 
-public class Trajectory implements Serializable {
+/**
+ * Position of Aircraft follows the Automatic Dependent Surveillance–Broadcast
+ * 
+ * (ADS-B) messages schema, with additional fields.
+ * 
+ * @author Ehab Qadah
+ * 
+ *         Feb 22, 2017
+ */
+
+public class PositionMessage implements Serializable {
 
 	private static final long serialVersionUID = 1013765199466780042L;
 	private String ID;
@@ -21,10 +31,10 @@ public class Trajectory implements Serializable {
 	private Double speed = null;
 	private Double acceleration;
 
-	public Trajectory() {
+	public PositionMessage() {
 	}
 
-	public Trajectory(String id, String type, double latitude, double longtitude, double altitude) {
+	public PositionMessage(String id, String type, double latitude, double longtitude, double altitude) {
 
 		this.ID = id;
 		this.setLatitude(latitude);

### 2017-02-22 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesSectorChangeDetector.java

@@ -10,7 +10,7 @@ import org.apache.flink.streaming.api.environment.StreamExecutionEnvironment;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Trajectory;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 import de.kdml.bigdatalab.spark_and_flink.flink.utils.FlinkUtils;
 
 /**
@@ -26,7 +26,7 @@ public class TrajectoriesSectorChangeDetector {
 	public static void main(String[] args) throws Exception {
 
 		StreamExecutionEnvironment env = FlinkUtils.getInitializedEnv();
-		KeyedStream<Tuple2<String, Trajectory>, Tuple> trajectoriesStream = TrajectoriesStreamUtils
+		KeyedStream<Tuple2<String, PositionMessage>, Tuple> trajectoriesStream = TrajectoriesStreamUtils
 				.getTrajectoriesStream(env);
 
 		DataSet<Sector> sectorsDataSet = getSectorDataSet();
@@ -35,7 +35,7 @@ public class TrajectoriesSectorChangeDetector {
 		// the sector between two consecutive trajectories
 		DataStream<String> trajectoriesStreamWithSectors = trajectoriesStream
 				.reduce(new TrajectoriesSectorsReducer(sectorsDataSet.collect())).filter(tuple -> {
-					Trajectory trajectory = tuple.f1;
+					PositionMessage trajectory = tuple.f1;
 					if (trajectory.getPrevSector() != null
 							&& !trajectory.getSector().equals(trajectory.getPrevSector())) {
 						return true;
@@ -44,7 +44,7 @@ public class TrajectoriesSectorChangeDetector {
 				}).map(tuple -> {
 
 					// print the sector transition
-					Trajectory trajectory = tuple.f1;
+					PositionMessage trajectory = tuple.f1;
 					return tuple.f0 + ": " + trajectory.getPrevSector().getNameAndAirBlock() + " --> "
 							+ trajectory.getSector().getNameAndAirBlock();
 

### 2017-02-22 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesSectorsReducer.java

@@ -7,9 +7,9 @@ import org.apache.flink.api.java.tuple.Tuple2;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.SectorUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Trajectory;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 
-public class TrajectoriesSectorsReducer implements ReduceFunction<Tuple2<String, Trajectory>> {
+public class TrajectoriesSectorsReducer implements ReduceFunction<Tuple2<String, PositionMessage>> {
 
 	private static final long serialVersionUID = 644511065257204161L;
 	public static List<Sector> sectors = null;
@@ -20,10 +20,10 @@ public class TrajectoriesSectorsReducer implements ReduceFunction<Tuple2<String,
 	}
 
 	@Override
-	public Tuple2<String, Trajectory> reduce(Tuple2<String, Trajectory> tuple1, Tuple2<String, Trajectory> tuple2) {
+	public Tuple2<String, PositionMessage> reduce(Tuple2<String, PositionMessage> tuple1, Tuple2<String, PositionMessage> tuple2) {
 
-		Tuple2<String, Trajectory> oldTuple = tuple2.f1.isNew() ? tuple1 : tuple2;
-		Tuple2<String, Trajectory> newTuple = tuple2.f1.isNew() ? tuple2 : tuple1;
+		Tuple2<String, PositionMessage> oldTuple = tuple2.f1.isNew() ? tuple1 : tuple2;
+		Tuple2<String, PositionMessage> newTuple = tuple2.f1.isNew() ? tuple2 : tuple1;
 
 		assignSector(oldTuple.f1);
 		assignSector(newTuple.f1);
@@ -35,7 +35,7 @@ public class TrajectoriesSectorsReducer implements ReduceFunction<Tuple2<String,
 		return newTuple;
 	}
 
-	private void assignSector(Trajectory trajectory) {
+	private void assignSector(PositionMessage trajectory) {
 		if (trajectory.getSector() == null) {
 			Sector sector = SectorUtils.getSectorForTrajectory(trajectory, sectors);
 			trajectory.setSector(sector);

### 2017-02-22 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesStatistics.java

@@ -5,7 +5,7 @@ import org.apache.flink.streaming.api.datastream.DataStream;
 import org.apache.flink.streaming.api.environment.StreamExecutionEnvironment;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.StatisticsUtils;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Trajectory;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 import de.kdml.bigdatalab.spark_and_flink.flink.utils.FlinkUtils;
 
 /**
@@ -28,13 +28,13 @@ public class TrajectoriesStatistics {
 		 * 
 		 **/
 
-		DataStream<Tuple2<String, Trajectory>> trajectoriesStream = TrajectoriesStreamUtils.getTrajectoriesStream(env)
-				.reduce((Tuple2<String, Trajectory> tuple1, Tuple2<String, Trajectory> tuple2) -> {
+		DataStream<Tuple2<String, PositionMessage>> trajectoriesStream = TrajectoriesStreamUtils.getTrajectoriesStream(env)
+				.reduce((Tuple2<String, PositionMessage> tuple1, Tuple2<String, PositionMessage> tuple2) -> {
 					// compute & aggregate statistics for the new trajectory
 					// based on the statistics of old trajectory
 					// discard the old trajectory and just keep the new one
-					Tuple2<String, Trajectory> oldTuple = tuple2.f1.isNew() ? tuple1 : tuple2;
-					Tuple2<String, Trajectory> newTuple = tuple2.f1.isNew() ? tuple2 : tuple1;
+					Tuple2<String, PositionMessage> oldTuple = tuple2.f1.isNew() ? tuple1 : tuple2;
+					Tuple2<String, PositionMessage> newTuple = tuple2.f1.isNew() ? tuple2 : tuple1;
 					StatisticsUtils.computeStatistics(oldTuple.f1, newTuple.f1);
 					newTuple.f1.setNew(false);
 					return newTuple;
@@ -52,12 +52,12 @@ public class TrajectoriesStatistics {
 	 * 
 	 * @param trajectoriesStream
 	 */
-	private static void showLatecies(DataStream<Tuple2<String, Trajectory>> trajectoriesStream) {
+	private static void showLatecies(DataStream<Tuple2<String, PositionMessage>> trajectoriesStream) {
 		DataStream<Long> latencies = trajectoriesStream.map(tuple -> {
 
[…]

### 2017-02-22 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesStreamUtils.java

@@ -13,7 +13,7 @@ import org.apache.flink.streaming.util.serialization.SimpleStringSchema;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.TrajectoriesUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.StreamRecord;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Trajectory;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 
 /**
  * Initialize the trajectories stream
@@ -32,7 +32,7 @@ public class TrajectoriesStreamUtils {
 	 * @param env
 	 * @return
 	 */
-	public static KeyedStream<Tuple2<String, Trajectory>, Tuple> getTrajectoriesStream(StreamExecutionEnvironment env) {
+	public static KeyedStream<Tuple2<String, PositionMessage>, Tuple> getTrajectoriesStream(StreamExecutionEnvironment env) {
 
 		// configure Kafka consumer
 		Properties props = new Properties();
@@ -67,10 +67,10 @@ public class TrajectoriesStreamUtils {
 		 * 
 		 **/
 
-		KeyedStream<Tuple2<String, Trajectory>, Tuple> trajectoriesStream = dataLines.map(line -> {
+		KeyedStream<Tuple2<String, PositionMessage>, Tuple> trajectoriesStream = dataLines.map(line -> {
 
 			StreamRecord streamRecord = StreamRecord.parseData(line);
-			Trajectory trajectory = TrajectoriesUtils.parseDataInput(streamRecord.getValue());
+			PositionMessage trajectory = TrajectoriesUtils.parseDataInput(streamRecord.getValue());
 			trajectory.setStreamedTime(streamRecord.getStreamedTime());
 			trajectory.setNew(true);
 			return new Tuple2<>(trajectory.getID(), trajectory);

### 2017-02-22 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectorySectorMapper.java

@@ -8,7 +8,7 @@ import org.apache.flink.api.java.tuple.Tuple2;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.SectorUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.TrajectoriesUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Trajectory;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 
 /**
  * Trajectory and Sector maper
@@ -18,7 +18,7 @@ import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Trajectory;
  *         Jan 5, 2017
  */
 final class TrajectorySectorMapper
-		implements MapFunction<Tuple2<String, List<Trajectory>>, Tuple2<String, List<Trajectory>>> {
+		implements MapFunction<Tuple2<String, List<PositionMessage>>, Tuple2<String, List<PositionMessage>>> {
 
 	private static final long serialVersionUID = -4643101889705396118L;
 	public static List<Sector> sectors = null;
@@ -38,12 +38,12 @@ final class TrajectorySectorMapper
 	 * 
 	 */
 	@Override
-	public Tuple2<String, List<Trajectory>> map(Tuple2<String, List<Trajectory>> tuple) throws Exception {
+	public Tuple2<String, List<PositionMessage>> map(Tuple2<String, List<PositionMessage>> tuple) throws Exception {
 
 		// sort trajectories
-		List<Trajectory> trajectories = TrajectoriesUtils.sortTrajectories(tuple.f1);
+		List<PositionMessage> trajectories = TrajectoriesUtils.sortTrajectories(tuple.f1);
 		// assign sector for all trajectories
-		for (Trajectory trajectory : trajectories) {
+		for (PositionMessage trajectory : trajectories) {
 
 			if (trajectory.getSector() == null) {
 				Sector sector = SectorUtils.getSectorForTrajectory(trajectory, sectors);
@@ -51,7 +51,7 @@ final class TrajectorySectorMapper
 			}
 
 		}
-		return new Tuple2<String, List<Trajectory>>(tuple.f0, trajectories);
+		return new Tuple2<String, List<PositionMessage>>(tuple.f0, trajectories);
 
 	}
 }

### 2017-02-22 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesSectorChangeDetector.java

@@ -16,7 +16,7 @@ import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.SectorUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.TrajectoriesUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Trajectory;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 import de.kdml.bigdatalab.spark_and_flink.spark_project.SparkConfigsUtils;
 
 /**
@@ -44,10 +44,10 @@ public class TrajectoriesSectorChangeDetector {
 		sectors = sc.broadcast(getSectorsDataSet(sc));
 		// Start the computation
 
-		JavaPairDStream<String, Iterable<Trajectory>> trajectories = TrajectoriesStreamUtils
+		JavaPairDStream<String, Iterable<PositionMessage>> trajectories = TrajectoriesStreamUtils
 				.getTrajectoriesStream(jssc);
 		// update the stream state
-		JavaPairDStream<String, Iterable<Trajectory>> runningTrajectories = trajectories
+		JavaPairDStream<String, Iterable<PositionMessage>> runningTrajectories = trajectories
 				.updateStateByKey(updateTrajectoriesAndAssignSectors);
 
 		// find trajectories with change in sector
@@ -55,10 +55,10 @@ public class TrajectoriesSectorChangeDetector {
 
 			// filter the trajectories that contain change in sector between two
 			// consecutive trajectories
-			Iterable<Trajectory> trajectoriesList = trajectoryTuple._2;
+			Iterable<PositionMessage> trajectoriesList = trajectoryTuple._2;
 
-			Trajectory prevTrajectory = null;
-			for (Trajectory trajectory : trajectoriesList) {
+			PositionMessage prevTrajectory = null;
+			for (PositionMessage trajectory : trajectoriesList) {
 				if (prevTrajectory != null) {
 					if ((prevTrajectory.isNew() || trajectory.isNew())
 							&& (!trajectory.getSector().equals(prevTrajectory.getSector()))) {
@@ -74,10 +74,10 @@ public class TrajectoriesSectorChangeDetector {
 			// print the sector change transitions
 			StringBuilder output = new StringBuilder(trajectoryTuple._1);
 			output.append(": ");
[…]

### 2017-02-22 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesStatistics.java

@@ -14,7 +14,7 @@ import org.apache.spark.streaming.api.java.JavaStreamingContext;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.StatisticsUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.TrajectoriesUtils;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Trajectory;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 import de.kdml.bigdatalab.spark_and_flink.spark_project.SparkConfigsUtils;
 
 /**
@@ -40,10 +40,10 @@ public class TrajectoriesStatistics {
 
 		// Start the computation
 
-		JavaPairDStream<String, Iterable<Trajectory>> trajectories = TrajectoriesStreamUtils
+		JavaPairDStream<String, Iterable<PositionMessage>> trajectories = TrajectoriesStreamUtils
 				.getTrajectoriesStream(jssc);
 		// update the stream state
-		JavaPairDStream<String, Iterable<Trajectory>> runningTrajectories = trajectories
+		JavaPairDStream<String, Iterable<PositionMessage>> runningTrajectories = trajectories
 				.updateStateByKey(updateTrajectoriesAndComputeStatistics);
 
 		printLatencies(runningTrajectories);
@@ -58,13 +58,13 @@ public class TrajectoriesStatistics {
 		}
 	}
 
-	private static void printLatencies(JavaPairDStream<String, Iterable<Trajectory>> runningTrajectories) {
+	private static void printLatencies(JavaPairDStream<String, Iterable<PositionMessage>> runningTrajectories) {
 		JavaDStream<Long> latencies = runningTrajectories.map(trajectoryTuple -> {
 
 			long currentTime = System.currentTimeMillis(), sumLatency = 0, count = 0;
 
-			Iterable<Trajectory> trajectoriesList = trajectoryTuple._2;
-			for (Trajectory trajectory : trajectoriesList) {
+			Iterable<PositionMessage> trajectoriesList = trajectoryTuple._2;
+			for (PositionMessage trajectory : trajectoriesList) {
 				if (trajectory.isNew()) {
 					sumLatency += currentTime - trajectory.getStreamedTime();
 					count++;
@@ -85,26 +85,26 @@ public class TrajectoriesStatistics {
 	 * values for the key
 	 * 
 	 */
[…]

### 2017-02-22 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesStreamUtils.java

@@ -20,7 +20,7 @@ import org.apache.spark.streaming.kafka010.LocationStrategies;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.TrajectoriesUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.StreamRecord;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Trajectory;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 import scala.Tuple2;
 
 /**
@@ -34,7 +34,7 @@ public class TrajectoriesStreamUtils {
 
 	private static Configs configs = Configs.getInstance();
 
-	public static JavaPairDStream<String, Iterable<Trajectory>> getTrajectoriesStream(JavaStreamingContext jssc) {
+	public static JavaPairDStream<String, Iterable<PositionMessage>> getTrajectoriesStream(JavaStreamingContext jssc) {
 
 		Set<String> topicsSet = new HashSet<String>();
 		topicsSet.add(configs.getStringProp("topicId"));
@@ -81,10 +81,10 @@ public class TrajectoriesStreamUtils {
 		 * RDDs of the new DStream. org.apache.spark.Partitioner is used to
 		 * control the partitioning of each RDD)
 		 */
-		JavaPairDStream<String, Iterable<Trajectory>> trajectories = dataStream.mapToPair(record -> {
+		JavaPairDStream<String, Iterable<PositionMessage>> trajectories = dataStream.mapToPair(record -> {
 
 			StreamRecord streamRecord = StreamRecord.parseData(record.value());
-			Trajectory trajectory = TrajectoriesUtils.parseDataInput(streamRecord.getValue());
+			PositionMessage trajectory = TrajectoriesUtils.parseDataInput(streamRecord.getValue());
 			trajectory.setStreamedTime(streamRecord.getStreamedTime());
 			trajectory.setNew(true);
 			

### 2017-02-22 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/data/StreamRecord.java
Added more documentation for the Spark implementation of Statistics computation

@@ -3,7 +3,8 @@ package de.kdml.bigdatalab.spark_and_flink.common_utils.data;
 import java.io.Serializable;
 
 /***
- * Stream record of value and streamed time
+ * Stream record of value (i.e., ADB-S message) {@link PositionMessage} and
+ * streamed time
  * 
  * @author Ehab Qadah
  * 
@@ -53,10 +54,7 @@ public class StreamRecord implements Serializable {
 	}
 
 	/**
-	 * Parse a stream record
-	 * 
-	 * @param line
-	 * @return
+	 * Parse a stream record from string line 
 	 */
 	public static StreamRecord parseData(String line) {
 

### 2017-02-22 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesStatistics.java

@@ -19,7 +19,7 @@ import de.kdml.bigdatalab.spark_and_flink.spark_project.SparkConfigsUtils;
 
 /**
  * 
- * Compute statistics per trajectory
+ * Compute statistics per trajectory in Spark Streaming
  * 
  * @author Ehab Qadah
  * 
@@ -38,11 +38,14 @@ public class TrajectoriesStatistics {
 		JavaStreamingContext jssc = new JavaStreamingContext(sc,
 				Durations.seconds(configs.getIntProp("batchDuration")));
 
-		// Start the computation
-
+		// create the trajectories stream
 		JavaPairDStream<String, Iterable<PositionMessage>> trajectories = TrajectoriesStreamUtils
 				.getTrajectoriesStream(jssc);
-		// update the stream state
+		/**
+		 * Manage the state of the trajectories stream by calculating the
+		 * statistics in every batch from aggregated old statistics, for more
+		 * details check {updateTrajectoriesAndComputeStatistics}
+		 */
 		JavaPairDStream<String, Iterable<PositionMessage>> runningTrajectories = trajectories
 				.updateStateByKey(updateTrajectoriesAndComputeStatistics);
 
@@ -58,6 +61,11 @@ public class TrajectoriesStatistics {
 		}
 	}
 
+	/**
+	 * Print latencies measurements
+	 * 
+	 * @param runningTrajectories
+	 */
 	private static void printLatencies(JavaPairDStream<String, Iterable<PositionMessage>> runningTrajectories) {
 		JavaDStream<Long> latencies = runningTrajectories.map(trajectoryTuple -> {
 
@@ -80,9 +88,10 @@ public class TrajectoriesStatistics {
 	}
 
 	/**
-	 * Return a new "state" DStream where the state for each key is updated by
-	 * applying the given function on the previous state of the key and the new
-	 * values for the key
+	 * Return a new "state" trajectories DStream where the state for each
+	 * trajectory key is updated by applying statistics calculation and
+	 * aggregation on the previous state of the key and the new values for the
+	 * key
 	 * 
 	 */
 	public static Function2<List<Iterable<PositionMessage>>, Optional<Iterable<PositionMessage>>, Optional<Iterable<PositionMessage>>> updateTrajectoriesAndComputeStatistics = new Function2<List<Iterable<PositionMessage>>, Optional<Iterable<PositionMessage>>, Optional<Iterable<PositionMessage>>>() {
@@ -90,83 +99,61 @@ public class TrajectoriesStatistics {
[…]

### 2017-02-22 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesStreamUtils.java

@@ -36,25 +36,26 @@ public class TrajectoriesStreamUtils {
 
 	public static JavaPairDStream<String, Iterable<PositionMessage>> getTrajectoriesStream(JavaStreamingContext jssc) {
 
+		//setup Kafka parameters 
 		Set<String> topicsSet = new HashSet<String>();
 		topicsSet.add(configs.getStringProp("topicId"));
-
 		Map<String, Object> kafkaParams = new HashMap<>();
 		kafkaParams.put("bootstrap.servers", configs.getStringProp("bootstrap.servers"));
 		kafkaParams.put("key.deserializer", StringDeserializer.class);
 		kafkaParams.put("value.deserializer", StringDeserializer.class);
 		kafkaParams.put("group.id", configs.getStringProp("kafkaGroupId"));
 		kafkaParams.put("auto.offset.reset", "latest");
-		//kafkaParams.put("enable.auto.commit", false);
+		kafkaParams.put("enable.auto.commit", false);
 
 		List<JavaDStream<ConsumerRecord<String, String>>> kafkaStreams = new ArrayList<>();
 
-		// to create parallel receivers for the same topic with same group to
-		// scale with kafka partitions
+		/**
+		 * create parallel receivers for the same topic within same group, in order to
+		 * scale with Kafka partitions
+		 **/
 		int numParallelKafkaStream = configs.getIntProp("numberOfKafkParallelStreams");
 		for (int i = 0; i < numParallelKafkaStream; i++) {
 			// Create direct kafka stream
-			// Create multiple stream and union them
 			JavaDStream<ConsumerRecord<String, String>> dataStreami = KafkaUtils.createDirectStream(jssc,
 					LocationStrategies.PreferConsistent(),
 					ConsumerStrategies.<String, String>Subscribe(topicsSet, kafkaParams));
@@ -62,19 +63,15 @@ public class TrajectoriesStreamUtils {
 			kafkaStreams.add(dataStreami);
 		}
 
-		// union all kafka streams
+		// union all parallel Kafka streams
 		JavaDStream<ConsumerRecord<String, String>> dataStream = numParallelKafkaStream == 1 ? kafkaStreams.get(0)
 				: jssc.union(kafkaStreams.get(0), kafkaStreams.subList(1, kafkaStreams.size()));
 
 		// Start the computation
-
 		/**
-		 * 1- parse data lines to create tuple<id,trajectory>
-		 * 
-		 * 2-filter messages
-		 * 
[…]

### 2017-02-22 paper/paper.tex
updated the introduction paragraph of section 3.2

@@
 \subsection{Evaluation Streaming Workloads}
-In our experimental evaluation, we developed two streaming processing workloads that read the ADS-B messages stream from Kafka to perform basic trajectories analysis methods.
-In our design choice, we try to cover key main points related to the streaming processing, to evaluate how  does each framework perform.
-The following are general key aspects of  streaming data processing were covered by our workloads design:
+In our experimental evaluation, we developed two real-time stream processing workloads that read the ADS-B messages stream from Kafka in order to perform basic trajectories analysis methods.
+In our design choice, we try to cover some key points that related to the streaming processing tasks, and evaluate the corresponding solutions in Spark and Flink.
+The following are general key aspects of  streaming data processing were covered by our workloads's design:
 
 \begin{itemize}
-\item Handling parallel input streams.
-\item How to aggregate the state of input stream records.
-\item Mange the order of stream records.
+\item Handling parallel input streams (e.g., Kafka Stream).
+\item How to aggregate the state of input stream.
+\item Manage the order of stream records.
 \item How to provide and update global data model in stream processing task.
@@
 \end{itemize}
-We present how Spark and Flink are handling these key aspects and related issues in each workload, then we discuss the performance evaluation in Section 4.
+We present the relation of workloads and these aspects, and th implementation details of Spark and Flink solutions in Section 3.2.1 \& 3.2.1.
+Afterward, In Section 4 we discuss the performance evaluation and analysis of workloads.
 

### 2017-02-22 paper/paper.tex
updated statistics computation workload.

@@
 \end{itemize}
-We present the relation of workloads and these aspects, and th implementation details of Spark and Flink solutions in Section 3.2.1 \& 3.2.1.
+We present the description of workloads and the relation with these aspects, and th implementation details of Spark and Flink solutions in Section 3.2.1 \& 3.2.1.
 Afterward, In Section 4 we discuss the performance evaluation and analysis of workloads.
@@
 \subsubsection{Statistics Computation per Trajectory}
-In first stream processing workload, we perform statistics calculation for each trajectory (i.e.,  messages of the same ID) in ADS-B messages stream.
+In first stream processing workload, we construct stream of trajectories by considering  the aircraft's positions (i.e., ADS-B messages) that belong to the same aircraft as trajectory.
+Moreover, we continuously compute and aggregate statistics for each arriving position in a  trajectory .
 As example of computed statistics indexes speed mean, mean of location coordinates, min and max altitude, etc.
-In this workload, we try to cover
-the parallel receiving of input data stream, state-full stream aggregation, and preserve the order of stream records.
+In this workload, we  cover the parallel receiving of input data stream, state-full aggregation on input data stream, and preserving the right order of the stream's records.
 The implementation details of this workload in Spark and Flink are presented in the following: 
@@
 
-The statistics calculation workload is implemented in Spark by performing the following steps: \begin{enumerate*}[label={\alph*)}]
-\item read parallel Kafka input streams 
-\item parse the multiple streams records   
-\item filter only the relevant  messages 
-\item combine the all messages related to the same trajectory
-\item manage the aggregate state and compute the statistics for each trajectory.
-\end{enumerate*}
-
-\par The used Spark Streaming API operations to execute those steps are shown in Figure 2.
-First, in Spark to consume parallel Kafka stream, multiple Kafka stream receivers must be created then combine them using the stream \texttt{union} API.
-Second, we parse each stream record (i.e., ADS-B message) then map it to tuple of aircraft ID and a position of aircraft object through \texttt{mapToPair} transformation API.
-Third, the irrelevant messages are filtered by the \texttt{filter}.
-Fourth, the tuples with same ID are grouped together using the \texttt{groupByKey} Spark API function, after applying this transformation the input stream records are shuffled which means to preserve the order of records a sort operation is needed.
-Finally, due to the mini-batch nature of Spark streaming, the input stream is processed batch by batch, we use the  \texttt{updateStateByKey} to compute statistics per trajectory after sorting the list of positions, then update the aggregated state of the trajectories stream.
-
-\begin{figure}[h]
-
-  \centering
-
-     \caption{Spark streaming implementation of statistics computation workload.}
-\end{figure}
+\par The Spark Streaming implementation of trajectory statistics computation read the Kafka stream using \texttt{KafkaUtils.createDirectStream} that creates \texttt{DStream} instance, in order to scale up and have parallel Kafka stream receivers, a multiple instances of \texttt{DStream} are created and combined together using \texttt{union} operation.
+The output stream of the \texttt{union} is processed using \texttt{mapToPair}, to map each position message to tuple of aircraft ID and position message.
+Then, the irrelevant tuples are filtered based on the position message type using  a \texttt{filter} transformation.
+Afterward, the tuple with same aircraft ID are groped together to construct trajectories (tuples of Id and list of positions), by applying the \texttt{groupByKey} operation.
+Since streaming model of Spark is micro-batch based and stateless, \texttt{updateStateByKey} operation is must be used to manage the state between the batches.
+In context of this workload, the statistics computation is preformed within the custom \texttt{updateStateByKey} function, which uses the last position of trajectory from previous batch to aggregate and compute statistics values to each trajectory's position in the new arriving batch.
+The list of new positions must be sorted to preserve the correct order, since the tuples after \texttt{groupByKey} operation are shuffled across the cluster's  nodes.
 

### 2017-02-22 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesStreamUtils.java

@@ -10,7 +10,6 @@ import java.util.Set;
 import org.apache.kafka.clients.consumer.ConsumerRecord;
 import org.apache.kafka.common.serialization.StringDeserializer;
 import org.apache.spark.streaming.api.java.JavaDStream;
-import org.apache.spark.streaming.api.java.JavaInputDStream;
 import org.apache.spark.streaming.api.java.JavaPairDStream;
 import org.apache.spark.streaming.api.java.JavaStreamingContext;
 import org.apache.spark.streaming.kafka010.ConsumerStrategies;
@@ -19,8 +18,8 @@ import org.apache.spark.streaming.kafka010.LocationStrategies;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.TrajectoriesUtils;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.StreamRecord;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.StreamRecord;
 import scala.Tuple2;
 
 /**
@@ -36,7 +35,7 @@ public class TrajectoriesStreamUtils {
 
 	public static JavaPairDStream<String, Iterable<PositionMessage>> getTrajectoriesStream(JavaStreamingContext jssc) {
 
-		//setup Kafka parameters 
+		// setup Kafka parameters
 		Set<String> topicsSet = new HashSet<String>();
 		topicsSet.add(configs.getStringProp("topicId"));
 		Map<String, Object> kafkaParams = new HashMap<>();
@@ -50,8 +49,8 @@ public class TrajectoriesStreamUtils {
 		List<JavaDStream<ConsumerRecord<String, String>>> kafkaStreams = new ArrayList<>();
 
 		/**
-		 * create parallel receivers for the same topic within same group, in order to
-		 * scale with Kafka partitions
+		 * create parallel receivers for the same topic within same group, in
+		 * order to scale with Kafka partitions
 		 **/
 		int numParallelKafkaStream = configs.getIntProp("numberOfKafkParallelStreams");
 		for (int i = 0; i < numParallelKafkaStream; i++) {
@@ -69,14 +68,15 @@ public class TrajectoriesStreamUtils {
 
 		// Start the computation
 		/**
-		 * 1- parse the stream record to create tuple<trajectory_id,position_message>
-		 * 2-filter irrelevant tuples based on position message type
[…]

### 2017-02-25 paper/paper.tex
small refinements

@@
 \centering\LARGE {\bf Comparative Evaluation of Spark and Flink Stream Processing}
+
+Supervisor: PD Dr. Michael Mock
 \thispagestyle{empty}
@@
 
-\par The rest of this paper is organized as follows: Section 2
-presents the main characteristics, APIs, and main data abstraction of Spark and Flink. Section 3  describes the experiment workloads and their implementation in each framework. Section 4
+\par The remainder of this paper is organized as follows.
+In Section 2, we present the main characteristics, APIs, and main data abstraction of Spark and Flink.
+Section 3  describes the experiment workloads and their implementation in each framework.
+Section 4
 provides and compares the performance results of the different workloads. Finally, Section 5 gives the overall conclusion.
@@
 In our experimental evaluation, we developed two real-time stream processing workloads that read the ADS-B messages stream from Kafka in order to perform basic trajectories analysis methods.
-In our design choice, we try to cover some key points that related to the streaming processing tasks, and evaluate the corresponding solutions in Spark and Flink.
-The following are general key aspects of  streaming data processing were covered by our workloads's design:
+The workloads are designed in such way that cover general key aspects of real-time streaming processing tasks, and evaluate the corresponding solutions in Spark and Flink.
+The following are the aspects of streaming data processing were covered by our workloads's design:
 
@@
 
+The detail of Flink's implementation of statistics computation workload as follows. First, \texttt{map}
+
 The statistics computation of trajectories stream is solved in Flink by performing the following steps: \begin{enumerate*}[label={\alph*)}]
@@
 Finally, the \texttt{reduce} API function is used to calculate the statistics and aggregate it.
-
-\begin{figure}[h]
-
-  \centering
-
-     \caption{Flink implementation of statistics computation workload.}
-\end{figure}
 

### 2017-02-25 paper/paper.tex
added Flink's implementation of statistics computaion

@@
 
-The detail of Flink's implementation of statistics computation workload as follows. First, \texttt{map}
-
-The statistics computation of trajectories stream is solved in Flink by performing the following steps: \begin{enumerate*}[label={\alph*)}]
-\item read Kafka input stream
-\item parse the input stream records   
-\item filter only the relevant  messages 
-\item combine the all messages related to the same trajectory
-\item compute the statistics for each trajectory.
-\end{enumerate*}
-
-\par The utilized Flink APIs to perform those steps are shown in Figure 3.
-First, create Kafka stream consumers that run in multiple parallel instances based on the configured \texttt{parallelism} factor.
-Second,  stream elements (i.e., ADS-B message) are mapped  to tuples of aircraft ID and  position object through \texttt{map} transformation API.
-Third, the irrelevant messages are filtered by the \texttt{filter}.
-Fourth, tuples with same ID are combined together using the \texttt{keyBy} API function.
-Finally, the \texttt{reduce} API function is used to calculate the statistics and aggregate it.
+The detail of Flink's implementation of statistics computation workload as follows.
+First, a \texttt{FlinkKafkaConsumer} is used to read the Kafka stream that constructs \texttt{DataStream} instance, moreover, it implicitly runs multiple parallel Kafka stream consumers, which makes the Kafka partitions balanced between the instances.
+Second, the stream records (i.e., ADS-B messages) are parsed and transformed to tuples of ID and position message using a configured \texttt{map} transformation.
+Third, a \texttt{filter} function is used to filter irrelevant tuples based on the message type.
+Fourth, the result tuples after the filtration steps are grouped with similar ID using \texttt{keyBy} operator, to constructs the stream of trajectories.
+Finally,
+a \texttt{reduce} transformation on the trajectories tuples is used to calculate the statistics for new received trajectory's position, utilizing the aggregated statistics of previous position, while old position is discarded and the new record is used to aggregate the statistics.
 

### 2017-02-26 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesStatistics.java
added more docs

@@ -25,10 +25,10 @@ public class TrajectoriesStatistics {
 		 * Reduce the trajectories by aggregate the statics from the previous
 		 * trajectory record and keep the new record, the further aggregation
 		 * and discard old record.
-		 * 
 		 **/
 
-		DataStream<Tuple2<String, PositionMessage>> trajectoriesStream = TrajectoriesStreamUtils.getTrajectoriesStream(env)
+		DataStream<Tuple2<String, PositionMessage>> trajectoriesStream = TrajectoriesStreamUtils
+				.getTrajectoriesStream(env)
 				.reduce((Tuple2<String, PositionMessage> tuple1, Tuple2<String, PositionMessage> tuple2) -> {
 					// compute & aggregate statistics for the new trajectory
 					// based on the statistics of old trajectory

### 2017-02-26 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesStreamUtils.java

@@ -12,8 +12,8 @@ import org.apache.flink.streaming.util.serialization.SimpleStringSchema;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.TrajectoriesUtils;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.StreamRecord;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.StreamRecord;
 
 /**
  * Initialize the trajectories stream
@@ -32,43 +32,37 @@ public class TrajectoriesStreamUtils {
 	 * @param env
 	 * @return
 	 */
-	public static KeyedStream<Tuple2<String, PositionMessage>, Tuple> getTrajectoriesStream(StreamExecutionEnvironment env) {
+	public static KeyedStream<Tuple2<String, PositionMessage>, Tuple> getTrajectoriesStream(
+			StreamExecutionEnvironment env) {
 
 		// configure Kafka consumer
 		Properties props = new Properties();
 		props.setProperty("zookeeper.connect", configs.getStringProp("zookeeper"));
 		props.setProperty("bootstrap.servers", configs.getStringProp("bootstrap.servers"));
 		props.setProperty("group.id", configs.getStringProp("kafkaGroupId"));
-
 		// Always read topicfrom start
 		props.setProperty("auto.offset.reset", "earliest");
 
 		// create a Kafka consumer
 		FlinkKafkaConsumer09<String> kafkaConsumer = new FlinkKafkaConsumer09<>(configs.getStringProp("topicId"),
 				new SimpleStringSchema(), props);
-
+		// constructs the data stream
 		DataStream<String> dataLines = env.addSource(kafkaConsumer);
 
 		/**
-		 * 1- map each data line to Tuple2<id,list of Trajectories>
+		 * 1- Map each data line to Tuple2<id,list of Trajectories>
 		 * 
 		 * 2-Filter messages based on type 3 &2
 		 * 
-		 * 3- keyBy- Logically partitions a stream into disjoint partitions,
-		 * each partition containing elements of the same key (0-> trajectory
-		 * ID) (keyBy Internally is implemented using Partitioner based on the
-		 * hash value of a key)
-		 * 
-		 * 4- Group the same trajectories using the reduce function on
[…]

### 2017-02-26 paper/paper.tex
added spark implementation details of sector change detection workload

@@
 
-The goal of sector change detection workload is to detect the  enter or leaving of aircraft from one air sector to other, by processing the trajectories stream.
-Given that the dataset of sectors (i.e., polygons) available as reference to check against it, also it used to assign the corresponding sector for certain aircraft's position.
-In this workload, providing the  global model of data (sectors dataset) in streaming processing workload is tested.Moreover, preserving the stream records order and parallel streams consuming are covered.
+The goal of sector change detection workload is to detect the  enter or leaving of aircraft from one air sector to other, by processing the real-time stream of trajectories.
+Given that the dataset of sectors (i.e., polygons) is available as reference to assign the corresponding sector for certain aircraft's position.
+In this workload, we test the aspect of providing the  global model of data (sectors dataset) in streaming processing workload.
+Moreover, the preserving od stream's records order and consuming of parallel Kafka streams points were covered.
 
  The implementation details of this workload in Spark and Flink are presented in the following:
+
  \begin{itemize}
@@
 
- we implement this workload, based on the following operation stages: 
- \begin{enumerate*}[label={\alph*)}]
- \item receive and build the trajectories stream 
- \item manage the aggregate state and assign the sector for each trajectory.
- \item identify the  change of sector per trajectory if exist
- \end{enumerate*}
-
- figure 4 shows the Spark streaming APIs used to implement trajectory's sector change detection workload.
-First, the trajectories stream is built using \texttt{mapToPair}, \texttt{filter}, and \texttt{groupByKey} APIs as was discussed in statistics computation workload.
-Second,  \texttt{updateStateByKey} API is used to assign corresponding sector for all new aircraft positions in  micro batch  and aggregate last position in previous batch.
-Finally, the detection of sector change is done by extra filtering transformation using \texttt{filter} API, by identifying the change of sectors in each trajectory.
-While the sector data set is provided using the \texttt{Broadcast} feature in Spark.
- \begin{figure}[h]
-
-   \centering
-
-      \caption{Spark streaming implementation of sector change detection workload.}
- \end{figure}
+ We implemented the sector change detection workload in Spark by constructing the trajectories stream and assigning the sectors for each position in all trajectories, then a change between the assigned sectors of two consecutive aircraft's positions is marked as alert case.
+First, a multiple instances of \texttt{DStream} are created using \texttt{KafkaUtils.createDirectStream} to read in  parallel from Kafka, after  all input data streams are combined using \texttt{union} function.
+Second, the records (i.e., ADS-B messages) of the resulted union stream are parsed and mapped to tuples of (aircraft ID, position message) using defined \texttt{mapToPair} transformation.
+Third, the unrelated  tuples are filtered based on the message type through \texttt{filter} operation.
+Fourth, the trajectories stream is constructed by combining all tuples with the same ID using the \texttt{groupByKey} function.
+Then, to manage and aggregate the state of trajectories stream between the mini-batches a configured \texttt{updateStateByKey} function is applied,  the corresponding sector are assigned, the old positions of certain trajectory from previous batch are discarded, but the last position of old batch is kept to be used alongside with the new positions.
+ The sector data set is provided using the \texttt{Broadcast} shared value feature in Spark that distributes the  data set to all cluster nodes efficiently.
+Finally, the detection of sector change is done by extra filtering transformation using \texttt{filter}, by checking the sector of the positions list in each tuple.
 

### 2017-02-26 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesSectorChangeDetector.java

@@ -15,8 +15,8 @@ import org.apache.spark.streaming.api.java.JavaStreamingContext;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.SectorUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.TrajectoriesUtils;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
 import de.kdml.bigdatalab.spark_and_flink.spark_project.SparkConfigsUtils;
 
 /**
@@ -54,7 +54,7 @@ public class TrajectoriesSectorChangeDetector {
 		JavaDStream<String> changedTrajectories = runningTrajectories.filter(trajectoryTuple -> {
 
 			// filter the trajectories that contain change in sector between two
-			// consecutive trajectories
+			// consecutive positions
 			Iterable<PositionMessage> trajectoriesList = trajectoryTuple._2;
 
 			PositionMessage prevTrajectory = null;
@@ -187,7 +187,8 @@ public class TrajectoriesSectorChangeDetector {
 		 * @param aggregatedTrajectories
 		 * @param trajectory
 		 */
-		private void appendTrajectoryAndAssignSector(List<PositionMessage> aggregatedTrajectories, PositionMessage trajectory) {
+		private void appendTrajectoryAndAssignSector(List<PositionMessage> aggregatedTrajectories,
+				PositionMessage trajectory) {
 
 			// assign sector for trajectory
 			if (trajectory.getSector() == null) {

### 2017-03-01 paper/paper.tex
updated  spark section

@@
 \section{Technical Background}
- In this section, we present the architecture, key characteristics, programming APIs  of Apache Spark, Apache Flink. The Apache Kafka platform is also discussed briefly.
+ In this section, we present the architecture, key characteristics, programming APIs  of Apache Spark and Apache Flink. Moreover, a brief overview of the Apache Kafka platform is  provided.
 
@@
 
-\par The main data abstraction in Spark is the Resilient Distributed Datasets (RDDs), which is a in-memory collection of elements partitioned across cluster of computers that can be processed in parallel \cite{rdd}.
-RDDs supports two categories of operations: transformations that drive new RDDs from the operated ones (e.g., map).
-And actions which return a value (e.g., count).
+\par The main data abstraction in Spark core is the Resilient Distributed Datasets (RDDs), which is in-memory collection of elements partitioned across cluster of computers that can be processed in parallel \cite{rdd}.
+RDDs supports two categories of operations: transformations that drive new RDDs from the operated ones (e.g., map), and actions which return a certain value (e.g., count).
+On the other hand, the Discretized Stream (DStream) is the data abstraction in Spark Streaming, which is series of RDDs that represent an input stream of data \cite{spark_streaming}.
+In other words, the Spark Streaming is batch-based processing model, which process the continuous stream of data by divided into micro-batches to be processed by the Spark engine.
+Figure 1 illustrates the general process flow of Spark Streaming.
 
-//TODO: DStreams, streaming model
+\begin{figure}[h]
+
+  \centering
+    \includegraphics[width=.9\textwidth, height=.2\textheight]{streaming-flow.png}
+     \caption{ Process flow of Spark Streaming \cite{spark_streaming}.}
+\end{figure}
 
@@
 
-\bibitem[5]{flink} 
+\bibitem[5]{spark_streaming}
+Spark Streaming. Available: http://spark.apache.org/docs/latest/streaming-programming-guide.html [Accessed March 2017].
+
+\bibitem[6]{flink} 
 Apache Flink. Available: https://flink.apache.org/ [Accessed February 2017].
 
-\bibitem[6]{kafka} 
+\bibitem[7]{kafka} 
 Apache Kafka. Available: https://kafka.apache.org/ [Accessed February 2017].

### 2017-03-01 paper/paper.tex
updated flink section

@@
 
-Apache Flink is an open source project that provide large-scale, distributed stream processing platform, while the batch processing treated as special case of streaming applications \cite{flink}.
-It offers programming API in Java, Scala.
-Its stack includes the core DataStream and DataSet APIs with additional libraries such as Complex event processing for Flink (FlinkCEP), Machine Learning for Flink (FlinkML), and Flink Graph API (Gelly).
+\par Apache Flink is an open source project that provide large-scale, distributed stream processing platform, while the batch processing treated as special case of streaming applications (i.e., finite stream)  \cite{flink}.
+It offers programming API in Java and Scala.
+Its software stack includes the core DataStream and DataSet APIs with additional libraries such as Complex event processing for Flink (FlinkCEP), Machine Learning for Flink (FlinkML), and Flink Graph API (Gelly).
 
-The main data abstraction of Flink are DataStream and DataSet that represents read-only collection of elements of the same type.
+\par The main data abstraction of Flink are DataStream and DataSet that represent read-only collection of data elements.
 The list of elements is bounded (i.e., finite) in DataSet, while it is unbounded (i.e., infinite) in case of DataStream.
 
+\par The Flink's core is a distributed streaming dataflow engine,  each Flink program is represented by a data flow graph (i.e., directed acyclic graph - DAG) that executed by the Flink's engine \cite{flink_paper}.
+The data flow graphs are composed of stateful, parallel operations and intermediate data stream partitions.
+Figure 2 shows a data flow graph of Flink program.
+
+\begin{figure}[h]
+
+  \centering
+    \includegraphics[width=\textwidth, height=.4\textheight]{flink_engine.png}
+     \caption{ An example of data flow graph \cite{flink}.}
+\end{figure} 
 \subsection{Apache Kafka}
-Apache Kafka is scalable, fault-tolerant distributed publish/subscribe streaming framework \cite{kafka}.
+\par Apache Kafka is scalable, fault-tolerant distributed publish/subscribe streaming framework \cite{kafka}.
 It manges the stream records in different categorizes (i.e., topics) that portioned and distributed over the servers in the Kafka cluster.
@@
 
-\bibitem[7]{kafka} 
+\bibitem[7]{flink_paper} 
+Carbone, P., Katsifodimos, A., Ewen, S., Markl, V., Haridi, S. et al. (2015)
+Apache flink: Stream and batch processing in a single engine.
+Bulletin of the IEEE Computer Society Technical Committee on Data Engineering, 36(4).
+
+\bibitem[8]{kafka} 
 Apache Kafka. Available: https://kafka.apache.org/ [Accessed February 2017].

### 2017-03-01 paper/paper.tex
updated kafka section

@@
 \par Apache Kafka is scalable, fault-tolerant distributed publish/subscribe streaming framework \cite{kafka}.
-It manges the stream records in different categorizes (i.e., topics) that portioned and distributed over the servers in the Kafka cluster.
+It manges the stream records in different categorizes (i.e., topics) that portioned and distributed over the servers of the Kafka cluster.
 It allows the data producers to publish stream of records to certain Kafka topic.
 While the consumer applications can subscribe to one or more topic to read the data stream.
-It has been widely  adopted, for example,
+The consumers are grouped into groups based on the group name that allows Kafka to distribute and balance the stream partitions among the  members of a certain group for the sake of scalability.
+Figure 3 shows how 4 partitions of stream is distributed differentially between two groups of consumers based on the number of members.
+Apache Kafka has been widely  adopted, for example, 
 Spark and Flink can receive data stream from Kafka.
+
+\begin{figure}[h]
+
+  \centering
+    \includegraphics[width=\textwidth, height=.3\textheight]{kafka_groups.png}
+     \caption{ Distribution of stream partitions for consumer groups in Kafka \cite{kafka}.}
+\end{figure}
 
@@
 \bibitem[8]{kafka} 
-Apache Kafka. Available: https://kafka.apache.org/ [Accessed February 2017].
+Apache Kafka. Available: https://kafka.apache.org/intro.html [Accessed February 2017].
 

### 2017-03-02 paper/paper.tex
updated introduction section

@@
 streaming processing. While which platform is the best still an open question. 
-In this paper, we provide a performance comparison of streaming data processing in Spark and Flink, by measuring different performance metrics, namely latency and throughput.
+In this paper, we provide a performance comparison of streaming data processing in Apache Spark and Apache Flink, by measuring different performance metrics, namely latency and throughput.
+In addition, we cover some key aspects of streaming data applications and how they are handled in both frameworks.
 The experiment is done using multiple streaming workloads over real-world datasets.
@@
 
-\par Big Data is now one of the most appealing topics in Information Technology, it has gained increasing attention of professionals in industry and academics, since the amount of data we generate is increasing exponentially \cite{idc} that generated from social media platforms, mobile phones, IoT and finical transactions, etc.
-Moreover, recent years showed a rapid development of Big Data frameworks such as Spark and Flink that provide more and efficient data processing capabilities.
- \par The Big Data is a method to process not just only just large datasets (volume), but also that arrives in high rate (velocity), and it comes in all kinds of data format (variety) (i.e., 3Vs of Big Data (Volume, Velocity, Variety))\cite{svs}.
+\par In recent years, Big Data is considered as one of the most appealing topics in Information Technology, it has gained increasing attention of professionals in industry and academics, as a result of the fact that the amount of data we generate is increasing exponentially \cite{idc} that generated from social media platforms, mobile phones, IoT and finical transactions, etc.
+Moreover, recent years showed a rapid development of Big Data frameworks such as Apache Spark and Apache Flink that provide more and efficient data processing capabilities.
+ \par The Big Data is a method to process not just only large datasets (volume), but also the data that arrives in high rate (velocity), and it comes in all kinds of data format (variety) (i.e., 3Vs of Big Data (Volume, Velocity, Variety)) \cite{svs}.
 
-\par The processing models of Big Data are Batch and Stream processing, Batch processing is a way to handle large, finite volumes of data (e.g., processing of historical transaction records), is more concerned with throughput.
+\par The processing models of Big Data are Batch processing and Stream processing, Batch processing is a way to handle large, finite volumes of data (e.g., processing of historical transaction records), is more concerned with throughput.
 While the Stream processing  is a method to manage fast and continuously incoming data (real-time), process it as it arrives (e.g., credit card fraud detection and network monitoring applications), where the low latency is required.
@@
 
-\par TODO: add paragraph about Yahoo benchmark and why do we still need evaluation 1- continuous dev of frameworks 2- explore more complex operation like groupby that shuffle the data.
+\par The are some performance evaluation of Apache Spark and Apache Flink and other framework such as Yahoo Streaming Benchmark \cite{yahoo}.
+We believe that we still need to carry our evaluation for two reasons: \begin{enumerate*}[label=(\roman*)]
+\item both frameworks are continuously improving in short time 
+\item we aim to study the performance of some complex operations such as \texttt{groupByKey} \& \texttt{keyBy} in Apache Spark and Apache Flink, respectively.
+\end{enumerate*}
+
 \par The purpose of this paper is to provide a comparative experimental evaluation of throughput, latency (i.e., performance) of stream processing in Apache Spark and Apache Flink.
 We use a datasets of airplanes trajectories provided in the context of Datacron project\footnote{\url{http://www.datacron-project.eu/}}, in order to simulate real-world use cases.
-Using multiple streaming data workloads we try to cover the main programming API differences between the both frameworks.
-Informally,  this work aims to help developer to determine which platform to choose in production.
+Using multiple streaming data workloads.
+Also we try to cover general aspects of streaming processing tasks (later details in Section 3) and how they can be solved in each platform.
+Informally,  this work aims to help the developers to determine which platform to choose in production for different use cases.
 
 \par The remainder of this paper is organized as follows.
-In Section 2, we present the main characteristics, APIs, and main data abstraction of Spark and Flink.
-Section 3  describes the experiment workloads and their implementation in each framework.
+In Section 2, we present the main characteristics, APIs, and main data abstraction of Apache Spark and Flink.
+Section 3  presents the general aspects of streaming processing, the design and implementation of experimental workloads.
 Section 4
@@
 
-\bibitem[3]{spark} 
+\bibitem[3]{yahoo} 
+Chintapalli, Sanket, et al.
+"Benchmarking streaming computation engines: Storm, Flink and Spark streaming." Parallel and Distributed Processing Symposium Workshops, 2016 IEEE International.
+IEEE, 2016.
+
+\bibitem[4]{spark} 
 Apache Spark. Available: https://spark.apache.org/ [Accessed February 2017].
 
-\bibitem[4]{rdd}
+\bibitem[5]{rdd}
 Zaharia, Matei, et al.
@@
 
-\bibitem[5]{spark_streaming}
+\bibitem[6]{spark_streaming}
 Spark Streaming. Available: http://spark.apache.org/docs/latest/streaming-programming-guide.html [Accessed March 2017].
 
-\bibitem[6]{flink} 
+\bibitem[7]{flink} 
 Apache Flink. Available: https://flink.apache.org/ [Accessed February 2017].
 
-\bibitem[7]{flink_paper} 
+\bibitem[8]{flink_paper} 
 Carbone, P., Katsifodimos, A., Ewen, S., Markl, V., Haridi, S. et al. (2015)
@@
 
-\bibitem[8]{kafka} 
+\bibitem[9]{kafka} 
 Apache Kafka. Available: https://kafka.apache.org/intro.html [Accessed February 2017].

### 2017-03-02 paper/paper.tex
Updated read me.

@@
 
-\par In recent years, Big Data is considered as one of the most appealing topics in Information Technology, it has gained increasing attention of professionals in industry and academics, as a result of the fact that the amount of data we generate is increasing exponentially \cite{idc} that generated from social media platforms, mobile phones, IoT and finical transactions, etc.
+\par In recent years, Big Data is considered as one of the most appealing topics in Information Technology, it has gained increasing attention of professionals in industry and academia, as a result of the fact that the amount of data we generate is increasing exponentially \cite{idc} that generated from social media platforms, mobile phones, IoT and finical transactions, etc.
 Moreover, recent years showed a rapid development of Big Data frameworks such as Apache Spark and Apache Flink that provide more and efficient data processing capabilities.
@@
 \subsubsection{Statistics Computation per Trajectory}
-In first stream processing workload, we construct stream of trajectories by considering  the aircraft's positions (i.e., ADS-B messages) that belong to the same aircraft as trajectory.
+In the first stream processing workload, we construct stream of trajectories by considering  the aircraft's positions (i.e., ADS-B messages) that belong to the same aircraft as trajectory.
 Moreover, we continuously compute and aggregate statistics for each arriving position in a trajectory .
-As example of computed statistics indexes speed mean, mean of location coordinates, min and max altitude, etc.
-In this workload, we  cover the parallel receiving of input data stream, state-full aggregation on input data stream, and preserving the right order of the stream's records.
+As example of computed statistics quantities speed mean, mean of location coordinates, min and max altitude, etc.
+In this workload, we  cover the parallel receiving of input data stream, statefull aggregation over the input data stream, and preserving the correct order of the stream's records.
 The implementation details of this workload in Spark and Flink are presented in the following: 

### 2017-03-02 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/StatisticsUtils.java
add avg speed calculation

@@ -81,24 +81,26 @@ public class StatisticsUtils {
 	/**
 	 * Compute statistics by aggregating from old trajectory
 	 * 
-	 * @param oldTrajectory
-	 * @param newTrajectory
+	 * @param oldPosition
+	 * @param newPosition
 	 */
-	public static void computeStatistics(PositionMessage oldTrajectory, PositionMessage newTrajectory) {
+	public static void computeStatistics(PositionMessage oldPosition, PositionMessage newPosition) {
 
-		if (oldTrajectory == null) {
-			intitStaticsForOldTrajectory(newTrajectory);
+		if (oldPosition == null) {
+			intitStaticsForIntitialPositionOfTrajectory(newPosition);
 			return;
 		}
-		intitStaticsForOldTrajectory(oldTrajectory);// initialize statics for
-													// the first trajectory only
+		intitStaticsForIntitialPositionOfTrajectory(oldPosition);// initialize
+																	// statics
+																	// for
+		// the first trajectory only
 		double minLongtitude, minLatitude, minAltitude, minSpeed, minAcceleration;
 		double maxLongtitude, maxLatitude, maxAltitude, maxSpeed, maxAcceleration;
-		TrajectoriesUtils.calculateDistanceAndSpeedOfTrajectory(oldTrajectory, newTrajectory);
-		double longtitude = newTrajectory.getLongitude(), lat = newTrajectory.getLatitude(),
-				altit = newTrajectory.getAltitude(), speed = newTrajectory.getSpeed(),
-				acceleration = newTrajectory.getAcceleration();
-		TrajectoryStatisticsWrapper oldStatistics = oldTrajectory.getStatistics();
+		TrajectoriesUtils.calculateDistanceAndSpeedOfTrajectory(oldPosition, newPosition);
+		double longtitude = newPosition.getLongitude(), lat = newPosition.getLatitude(),
+				altit = newPosition.getAltitude(), speed = newPosition.getSpeed(),
+				acceleration = newPosition.getAcceleration(), newSpeed = newPosition.getSpeed();
+		TrajectoryStatisticsWrapper oldStatistics = oldPosition.getStatistics();
 		// update min longitude
 		minLongtitude = Math.min(oldStatistics.getMinLong(), longtitude);
 		// update min latitude
@@ -114,6 +116,10 @@ public class StatisticsUtils {
[…]

### 2017-03-02 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/TrajectoriesUtils.java
code cleaning and refactoring

@@ -4,8 +4,6 @@ import java.time.Duration;
 import java.util.Iterator;
 import java.util.List;
 
-import javax.swing.text.html.HTMLDocument.HTMLReader.SpecialAction;
-
 import com.google.common.collect.Lists;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
@@ -51,14 +49,14 @@ public class TrajectoriesUtils {
 	}
 
 	/**
-	 * Sort the trajectories based on created time
+	 * Sort the positions based on created time
 	 * 
-	 * @param trajectories
+	 * @param positions
 	 * @return
 	 */
-	public static List<PositionMessage> sortTrajectories(List<PositionMessage> trajectories) {
+	public static List<PositionMessage> sortPositionsOfTrajectory(List<PositionMessage> positions) {
 
-		Iterator<PositionMessage> sortedTrajectories = trajectories.stream().sorted((trajectory1, trajector2) -> {
+		Iterator<PositionMessage> sortedTrajectories = positions.stream().sorted((trajectory1, trajector2) -> {
 
 			return -1 * trajectory1.getCreatedDateTime().compareTo(trajector2.getCreatedDateTime());
 		}).iterator();
@@ -70,35 +68,36 @@ public class TrajectoriesUtils {
 	/**
 	 * Calculate Speed and distance of trajectory
 	 * 
-	 * @param prevTrajectory
-	 * @param trajectory
+	 * @param prevPosition
+	 * @param currentPosition
 	 */
-	public static void calculateDistanceAndSpeedOfTrajectory(PositionMessage prevTrajectory, PositionMessage trajectory) {
+	public static void calculateDistanceAndSpeedOfTrajectory(PositionMessage prevPosition,
+			PositionMessage currentPosition) {
 
-		if (trajectory.getSpeed() != null && trajectory.getDistance() != null) {
+		if (currentPosition.getSpeed() != null && currentPosition.getDistance() != null) {
 			return;
 		}
-		double distance = prevTrajectory == null ? 0
-				: GeoUtils.greatCircleDistance(prevTrajectory.getLatitude(), prevTrajectory.getLongitude(),
-						trajectory.getLatitude(), trajectory.getLongitude()),
+		double distance = prevPosition == null ? 0
+				: GeoUtils.greatCircleDistance(prevPosition.getLatitude(), prevPosition.getLongitude(),
[…]

### 2017-03-02 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/data/TrajectoryStatisticsWrapper.java

@@ -14,72 +14,36 @@ public class TrajectoryStatisticsWrapper implements Serializable {
 	public static final long serialVersionUID = 686131645808716203L;
 
 	private long numPoints;
-
 	private double minSpeed;
-
 	private double maxSpeed;
-
 	private double avgSpeed;
-
 	private double medianSpeed;
-
 	private double minAcceleration;
-
 	private double maxAcceleration;
-
 	private double avgAcceleration;
-
 	private double medianAcceleration;
-
 	private double minDifftime;
-
 	private double maxDifftime;
-
 	private double avgDifftime;
-
 	private double median_difftime;
-
 	private double minDate;
-
 	private double maxDate;
-
 	private double minLong;
-
 	private double maxLong;
-
 	private double maxAltitude;
-
 	private double minAltitude;
-
 	private double minLat;
-
 	private double maxLat;
+	private double aggergatedSpeedSum;
 
 	public TrajectoryStatisticsWrapper() {
 	}
 
-	public TrajectoryStatisticsWrapper(//
-			String id, //
-			int id_c, //
-			long nr_points, //
-			double min_speed, //
-			double max_speed, //
-			double avg_speed, //
-			double median_speed, //
-			double min_acceleration, //
-			double max_acceleration, //
-			double avg_acceleration, //
-			double median_acceleration, //
-			double min_difftime, //
-			double max_difftime, //
-			double avg_difftime, //
-			double median_difftime, //
-			double min_date, //
-			double max_date, //
-			double min_X, //
-			double max_X, //
-			double min_Y, //
-			double max_Y) {
+	public TrajectoryStatisticsWrapper(String id, int id_c, long nr_points, double min_speed, double max_speed,
+			double avg_speed, double median_speed, double min_acceleration, double max_acceleration,
+			double avg_acceleration, double median_acceleration, double min_difftime, double max_difftime,
+			double avg_difftime, double median_difftime, double min_date, double max_date, double min_X, double max_X,
+			double min_Y, double max_Y) {
 
 		super();
 
@@ -108,37 +72,8 @@ public class TrajectoryStatisticsWrapper implements Serializable {
 	public String toString() {
 		StringBuilder builder = new StringBuilder();
 		builder.append("Statistics ");
-
 		builder.append("[ nr_points=");
 		builder.append(getNumPoints());
[…]

### 2017-03-02 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectorySectorMapper.java

@@ -41,7 +41,7 @@ final class TrajectorySectorMapper
 	public Tuple2<String, List<PositionMessage>> map(Tuple2<String, List<PositionMessage>> tuple) throws Exception {
 
 		// sort trajectories
-		List<PositionMessage> trajectories = TrajectoriesUtils.sortTrajectories(tuple.f1);
+		List<PositionMessage> trajectories = TrajectoriesUtils.sortPositionsOfTrajectory(tuple.f1);
 		// assign sector for all trajectories
 		for (PositionMessage trajectory : trajectories) {
 

### 2017-03-02 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesSectorChangeDetector.java

@@ -178,7 +178,7 @@ public class TrajectoriesSectorChangeDetector {
 
 			}
 			// update state
-			return Optional.of(TrajectoriesUtils.sortTrajectories(aggregatedTrajectories));
+			return Optional.of(TrajectoriesUtils.sortPositionsOfTrajectory(aggregatedTrajectories));
 		}
 
 		/**

### 2017-03-02 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesStatistics.java

@@ -130,7 +130,7 @@ public class TrajectoriesStatistics {
 			// calculating the statistics quantities, and the sort is needed
 			// since we need to preserve the right order of the position
 			// messages based on the streamed time
-			aggregatedPositions = TrajectoriesUtils.sortTrajectories(aggregatedPositions);
+			aggregatedPositions = TrajectoriesUtils.sortPositionsOfTrajectory(aggregatedPositions);
 
 			for (PositionMessage trajectory : aggregatedPositions) {
 				// Compute statistics for each position message based on

### 2017-03-03 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/StatisticsUtils.java
updated data stream producer & speed calculation

@@ -86,14 +86,12 @@ public class StatisticsUtils {
 	 */
 	public static void computeStatistics(PositionMessage oldPosition, PositionMessage newPosition) {
 
+		// for the first received position
 		if (oldPosition == null) {
 			intitStaticsForIntitialPositionOfTrajectory(newPosition);
 			return;
 		}
-		intitStaticsForIntitialPositionOfTrajectory(oldPosition);// initialize
-																	// statics
-																	// for
-		// the first trajectory only
+		intitStaticsForIntitialPositionOfTrajectory(oldPosition);
 		double minLongtitude, minLatitude, minAltitude, minSpeed, minAcceleration;
 		double maxLongtitude, maxLatitude, maxAltitude, maxSpeed, maxAcceleration;
 		TrajectoriesUtils.calculateDistanceAndSpeedOfTrajectory(oldPosition, newPosition);

### 2017-03-03 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/TrajectoriesUtils.java

@@ -80,13 +80,13 @@ public class TrajectoriesUtils {
 		double distance = prevPosition == null ? 0
 				: GeoUtils.greatCircleDistance(prevPosition.getLatitude(), prevPosition.getLongitude(),
 						currentPosition.getLatitude(), currentPosition.getLongitude()),
-				speed = 0.0, diffTime, acceleration = 0.0;
+				speed = 0.0, diffTime = 0.0, acceleration = 0.0;
 
 		if (prevPosition != null) {
 
 			long diff = Duration.between(prevPosition.getCreatedDateTime(), currentPosition.getCreatedDateTime())
 					.toMillis();
-			diffTime = ((double) (diff)) / (1000.0 * 60.0 * 60.0);
+			diffTime = ((double) (diff)) / (1000.0 * 60.0 * 60.0); // to hours
 
 			if (diffTime != 0.0) {
 				speed = distance / diffTime;
@@ -95,7 +95,7 @@ public class TrajectoriesUtils {
 
 		}
 
-		currentPosition.setSpeed(speed);
+		currentPosition.setSpeed(Math.abs(speed));
 		currentPosition.setDistance(distance);
 		currentPosition.setAcceleration(acceleration);
 

### 2017-03-03 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/data/PositionMessage.java

@@ -89,7 +89,7 @@ public class PositionMessage implements Serializable {
 
 		// TODO: use string builder
 		return "|" + isNew() + getCreatedDateTime() + ":(" + getLatitude() + "," + getLongitude() + "," + getAltitude()
-				+ ") |" + " speed=" + getSpeed() + " Distance" + getDistance() + "|"
+				+ ") |" + " speed=" + getSpeed() + " Distance=" + getDistance() + "|"
 				+ (getStatistics() != null ? getStatistics() : "  ") + (getSector() != null ? getSector() : " ") + "\n";
 	}
 

### 2017-03-03 flink_project/src/main/java/kafka/datacorn/DatacornKafkaStreamProducer.java

@@ -14,7 +14,7 @@ import org.apache.flink.streaming.connectors.kafka.FlinkKafkaProducer09;
 import org.apache.flink.streaming.connectors.kafka.partitioner.KafkaPartitioner;
 import org.apache.flink.streaming.util.serialization.SimpleStringSchema;
 
-import de.kdml.bigdatalab.spark_and_flink.common_utils.*;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.StreamRecord;
 
 /**
@@ -82,7 +82,7 @@ public class DatacornKafkaStreamProducer {
 		/**
 		 * Generates a new random line every LINE_SLIDE_TIME_MS
 		 */
-		private static final int LINE_SLIDE_TIME_MS = 2*1000;
+		private static final int LINE_SLIDE_TIME_MS = 2 * 1000;
 
 		private static final long serialVersionUID = 2174904787118597072L;
 
@@ -100,13 +100,6 @@ public class DatacornKafkaStreamProducer {
 						line = new StreamRecord(line).toString();
 						ctx.collect(line);
 						Thread.sleep(LINE_SLIDE_TIME_MS);
-						ctx.collect(line);
-
-						Thread.sleep(LINE_SLIDE_TIME_MS);
-
-						ctx.collect(line);
-						Thread.sleep(LINE_SLIDE_TIME_MS);
-						
 					}
 				} catch (Exception e) {
 

### 2017-03-03 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesStatistics.java
Added more documentation for the statistics computation workload

@@ -15,6 +15,7 @@ import de.kdml.bigdatalab.spark_and_flink.flink.utils.FlinkUtils;
  * 
  *         Jan 2, 2017
  */
+
 public class TrajectoriesStatistics {
 
 	public static void main(String[] args) throws Exception {
@@ -52,6 +53,7 @@ public class TrajectoriesStatistics {
 	 * 
 	 * @param trajectoriesStream
 	 */
+
 	private static void showLatecies(DataStream<Tuple2<String, PositionMessage>> trajectoriesStream) {
 		DataStream<Long> latencies = trajectoriesStream.map(tuple -> {
 

### 2017-03-03 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesStreamUtils.java

@@ -22,16 +22,15 @@ import de.kdml.bigdatalab.spark_and_flink.common_utils.data.StreamRecord;
  * 
  *         Jan 5, 2017
  */
+
 public class TrajectoriesStreamUtils {
 
 	private static Configs configs = Configs.getInstance();
 
 	/**
-	 * Get trajectories stream
-	 * 
-	 * @param env
-	 * @return
+	 * Get the trajectories stream
 	 */
+
 	public static KeyedStream<Tuple2<String, PositionMessage>, Tuple> getTrajectoriesStream(
 			StreamExecutionEnvironment env) {
 
@@ -47,7 +46,7 @@ public class TrajectoriesStreamUtils {
 		FlinkKafkaConsumer09<String> kafkaConsumer = new FlinkKafkaConsumer09<>(configs.getStringProp("topicId"),
 				new SimpleStringSchema(), props);
 		// constructs the data stream
-		DataStream<String> dataLines = env.addSource(kafkaConsumer);
+		DataStream<String> postitionMessagesStream = env.addSource(kafkaConsumer);
 
 		/**
 		 * 1- Map each data line to Tuple2<id,list of Trajectories>
@@ -61,8 +60,9 @@ public class TrajectoriesStreamUtils {
 		 * of a key)
 		 **/
 
-		KeyedStream<Tuple2<String, PositionMessage>, Tuple> trajectoriesStream = dataLines.map(line -> {
-			// parse ADS-B messages and construct tuple for each message
+		KeyedStream<Tuple2<String, PositionMessage>, Tuple> trajectoriesStream = postitionMessagesStream.map(line -> {
+			// parse ADS-B messages and construct tuple of ID & position
+			// message for each message
 			StreamRecord streamRecord = StreamRecord.parseData(line);
 			PositionMessage trajectory = TrajectoriesUtils.parseDataInput(streamRecord.getValue());
 			trajectory.setStreamedTime(streamRecord.getStreamedTime());

### 2017-03-03 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesStatistics.java

@@ -1,6 +1,7 @@
 package de.kdml.bigdatalab.spark_and_flink.spark_project.datacorn;
 
 import java.util.ArrayList;
+import java.util.Arrays;
 import java.util.List;
 
 import org.apache.spark.api.java.JavaSparkContext;
@@ -13,7 +14,6 @@ import org.apache.spark.streaming.api.java.JavaStreamingContext;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.StatisticsUtils;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.TrajectoriesUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 import de.kdml.bigdatalab.spark_and_flink.spark_project.SparkConfigsUtils;
 
@@ -35,6 +35,7 @@ public class TrajectoriesStatistics {
 		// configure spark streaming context
 		JavaSparkContext sc = SparkConfigsUtils.getSparkContext("Trajectories Statistics Computation ");
 
+		// configure the stream batch interval
 		JavaStreamingContext jssc = new JavaStreamingContext(sc,
 				Durations.seconds(configs.getIntProp("batchDuration")));
 
@@ -102,35 +103,27 @@ public class TrajectoriesStatistics {
 		public Optional<Iterable<PositionMessage>> call(List<Iterable<PositionMessage>> newPostitions,
 				Optional<Iterable<PositionMessage>> statePositions) {
 
-			// Get last old trajectory to be used in statistics computation for
-			// new batch positions
+			// Get the last old trajectory from the previous batch to be used in
+			// statistics computation for the new batch positions
 
 			Iterable<PositionMessage> prevBatchPositions = statePositions.orElse(new ArrayList<>());
 
 			// Get last position in previous batch
-			PositionMessage lastOldPosition = getLastPosition(prevBatchPositions);
+			PositionMessage lastOldPosition = TrajectoriesStreamUtils.getLastPositionInBatch(prevBatchPositions);
 
-			// Combine all new batch positions
-			List<PositionMessage> aggregatedPositions = new ArrayList<>();
-			// aggregate new values
-			for (Iterable<PositionMessage> val : newPostitions) {
-				for (PositionMessage position : val) {
-					aggregatedPositions.add(position);
-				}
-
-			}
[…]

### 2017-03-03 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesStreamUtils.java

@@ -49,12 +49,12 @@ public class TrajectoriesStreamUtils {
 		List<JavaDStream<ConsumerRecord<String, String>>> kafkaStreams = new ArrayList<>();
 
 		/**
-		 * create parallel receivers for the same topic within same group, in
-		 * order to scale with Kafka partitions
+		 * Create parallel receivers for the same topic within the same group,
+		 * in order to scale with Kafka partitions
 		 **/
 		int numParallelKafkaStream = configs.getIntProp("numberOfKafkParallelStreams");
 		for (int i = 0; i < numParallelKafkaStream; i++) {
-			// Create direct kafka stream
+			// Create direct Kafka stream
 			JavaDStream<ConsumerRecord<String, String>> dataStreami = KafkaUtils.createDirectStream(jssc,
 					LocationStrategies.PreferConsistent(),
 					ConsumerStrategies.<String, String>Subscribe(topicsSet, kafkaParams));
@@ -62,7 +62,8 @@ public class TrajectoriesStreamUtils {
 			kafkaStreams.add(dataStreami);
 		}
 
-		// union all parallel Kafka streams
+		// union all parallel Kafka streams, and use the reslut stream as the
+		// main stream
 		JavaDStream<ConsumerRecord<String, String>> dataStream = numParallelKafkaStream == 1 ? kafkaStreams.get(0)
 				: jssc.union(kafkaStreams.get(0), kafkaStreams.subList(1, kafkaStreams.size()));
 
@@ -80,18 +81,51 @@ public class TrajectoriesStreamUtils {
 		 */
 		JavaPairDStream<String, Iterable<PositionMessage>> trajectories = dataStream.mapToPair(record -> {
 
+			// parse the ADBS-B messages from the Kafka stream
 			StreamRecord streamRecord = StreamRecord.parseData(record.value());
 			PositionMessage trajectory = TrajectoriesUtils.parseDataInput(streamRecord.getValue());
 			trajectory.setStreamedTime(streamRecord.getStreamedTime());
 			trajectory.setNew(true);
 
+			// create a tuple of Id & position message for each stream line
 			return new Tuple2<>(trajectory.getID(), trajectory);
 		}).filter(tuple -> {
-			// get only msg2 & msg3 types
+			// filter only msg2 & msg3 types
 			return ("MSG2".equals(tuple._2.getType()) || "MSG3".equals(tuple._2.getType()));
 		}).groupByKey();
-
+		// Group all message belong to same aircraft as trajectory
[…]

### 2017-03-03 paper/paper.tex
updated the spark implementaion of statistcs computation

@@
 \subsubsection{Statistics Computation per Trajectory}
-In the first stream processing workload, we construct stream of trajectories by considering  the aircraft's positions (i.e., ADS-B messages) that belong to the same aircraft as trajectory.
-Moreover, we continuously compute and aggregate statistics for each arriving position in a trajectory .
+In the first stream processing workload, we construct a stream of trajectories by considering  the position messages (i.e., ADS-B messages) that belong to the same aircraft as trajectory.
+Moreover, we continuously compute and aggregate statistics for each new position in a trajectory .
 As example of computed statistics quantities speed mean, mean of location coordinates, min and max altitude, etc.
@@
 
-\par The Spark Streaming implementation of trajectory statistics computation read the Kafka stream using \texttt{KafkaUtils.createDirectStream} that creates \texttt{DStream} instance, in order to scale up and have parallel Kafka stream receivers, a multiple instances of \texttt{DStream} are created and combined together using \texttt{union} operation.
-The output stream of the \texttt{union} is processed using \texttt{mapToPair}, to map each position message to tuple of aircraft ID and position message.
+\par The Spark Streaming implementation of trajectory statistics computation reads the Kafka stream using \texttt{KafkaUtils.createDirectStream} that creates \texttt{DStream} that represents the stream of position messages, in order to scale up and have parallel Kafka stream receivers, a multiple instances of \texttt{DStream} must be created and combined together using the \texttt{union} operation.
+The output stream of the \texttt{union} is further processed using \texttt{mapToPair}, to map each position message to tuple of aircraft ID and position message.
 Then, the irrelevant tuples are filtered based on the position message type using  a \texttt{filter} transformation.
-Afterward, the tuple with same aircraft ID are groped together to construct trajectories (tuples of Id and list of positions), by applying the \texttt{groupByKey} operation.
-Since streaming model of Spark is micro-batch based and stateless, \texttt{updateStateByKey} operation is must be used to manage the state between the batches.
-In context of this workload, the statistics computation is preformed within the custom \texttt{updateStateByKey} function, which uses the last position of trajectory from previous batch to aggregate and compute statistics values to each trajectory's position in the new arriving batch.
-The list of new positions must be sorted to preserve the correct order, since the tuples after \texttt{groupByKey} operation are shuffled across the cluster's  nodes.
+Afterward, the tuples that share the same aircraft ID are groped together to construct the stream of trajectories (tuples of Id and list of positions), by applying the \texttt{groupByKey} operation.
+Since the processing  model of Spark Streaming is micro-batch based and stateless, \texttt{updateStateByKey} operation must be used to maintain the state of the trajectories stream.
+In the context of this workload, the state is the aggregated statistics of trajectory, so the statistics computation is preformed within the custom \texttt{updateStateByKey} function that continuously applied for every batch.
+ The state update function uses the last position of trajectory from previous batch to aggregate and compute statistics quantities to each trajectory's position in the new data batch, and the new positions list is kept as the new state for the trajectory.
+Inside the state update function, the new positions must be sorted to preserve the correct order, since the tuples after \texttt{groupByKey} operation are shuffled across the cluster's  nodes and sent to the state update function.
 

### 2017-03-03 paper/paper.tex
[paper] updated sector change detection in Spark

@@
 
-\par The Spark Streaming implementation of trajectory statistics computation reads the Kafka stream using \texttt{KafkaUtils.createDirectStream} that creates \texttt{DStream} that represents the stream of position messages, in order to scale up and have parallel Kafka stream receivers, a multiple instances of \texttt{DStream} must be created and combined together using the \texttt{union} operation.
+\par The implementation of trajectory statistics computation in Spark Streaming, first, it reads the Kafka stream using \texttt{KafkaUtils.createDirectStream} that creates \texttt{DStream} that represents the stream of position messages, in order to scale up and have parallel Kafka stream receivers, a multiple instances of \texttt{DStream} must be created and combined together using the \texttt{union} operation.
 The output stream of the \texttt{union} is further processed using \texttt{mapToPair}, to map each position message to tuple of aircraft ID and position message.
@@
 
-The detail of Flink's implementation of statistics computation workload as follows.
-First, a \texttt{FlinkKafkaConsumer} is used to read the Kafka stream that constructs \texttt{DataStream} instance, moreover, it implicitly runs multiple parallel Kafka stream consumers, which makes the Kafka partitions balanced between the instances.
+The details of Flink's implementation of statistics computation workload as follows.
+First, a \texttt{FlinkKafkaConsumer} is used to  construct a \texttt{DataStream} instance that represent the Kafka stream of aircrafts positions, it implicitly runs multiple parallel Kafka stream consumers, which makes the Kafka partitions balanced between the instances.
 Second, the stream records (i.e., ADS-B messages) are parsed and transformed to tuples of ID and position message using a configured \texttt{map} transformation.
 Third, a \texttt{filter} function is used to filter irrelevant tuples based on the message type.
-Fourth, the result tuples after the filtration steps are grouped with similar ID using \texttt{keyBy} operator, to constructs the stream of trajectories.
+Fourth, the resulted tuples after the filtration transformation are grouped based on the tuple's ID using \texttt{keyBy} operator that constructs the \texttt{KeyedStream} of trajectories.
 Finally,
-a \texttt{reduce} transformation on the trajectories tuples is used to calculate the statistics for new received trajectory's position, utilizing the aggregated statistics of previous position, while old position is discarded and the new record is used to aggregate the statistics.
+a \texttt{reduce} transformation on the  \texttt{KeyedStream} is used to calculate the statistics for each new coming trajectory's position, by utilizing the aggregated statistics of previous position, while old position is discarded and the new position with attached computed statistics is used to aggregate the statistics of a certain trajectory.
 
@@
 
-The goal of sector change detection workload is to detect the  enter or leaving of aircraft from one air sector to other, by processing the real-time stream of trajectories.
+The goal of sector change detection workload is to detect the  enter or leaving of aircraft from one air sector to other, by processing the real-time stream of aircraft positions.
 Given that the dataset of sectors (i.e., polygons) is available as reference to assign the corresponding sector for certain aircraft's position.
 In this workload, we test the aspect of providing the  global model of data (sectors dataset) in streaming processing workload.
-Moreover, the preserving od stream's records order and consuming of parallel Kafka streams points were covered.
+Moreover, the preserving of the stream's records order and consuming of parallel Kafka streams points are also covered.
 
@@
 
- We implemented the sector change detection workload in Spark by constructing the trajectories stream and assigning the sectors for each position in all trajectories, then a change between the assigned sectors of two consecutive aircraft's positions is marked as alert case.
+ We implemented the sector change detection workload in Spark Streaming by constructing the trajectories stream and assigning the sectors for each position of a trajectory, then a change between the assigned sectors of two consecutive aircraft's positions is marked  as  an transition of sector for certain trajectory.
 First, a multiple instances of \texttt{DStream} are created using \texttt{KafkaUtils.createDirectStream} to read in  parallel from Kafka, after  all input data streams are combined using \texttt{union} function.
 Second, the records (i.e., ADS-B messages) of the resulted union stream are parsed and mapped to tuples of (aircraft ID, position message) using defined \texttt{mapToPair} transformation.
-Third, the unrelated  tuples are filtered based on the message type through \texttt{filter} operation.
+Third, the unrelated  tuples are filtered based on the message type through a \texttt{filter} operation.
 Fourth, the trajectories stream is constructed by combining all tuples with the same ID using the \texttt{groupByKey} function.
-Then, to manage and aggregate the state of trajectories stream between the mini-batches a configured \texttt{updateStateByKey} function is applied,  the corresponding sector are assigned, the old positions of certain trajectory from previous batch are discarded, but the last position of old batch is kept to be used alongside with the new positions.
- The sector data set is provided using the \texttt{Broadcast} shared value feature in Spark that distributes the  data set to all cluster nodes efficiently.
-Finally, the detection of sector change is done by extra filtering transformation using \texttt{filter}, by checking the sector of the positions list in each tuple.
+Then, to manage and aggregate the state of trajectories stream between the mini-batches a configured \texttt{updateStateByKey} function is applied.
+The state update function assign the corresponding sector for each new position, the old positions of a certain trajectory from previous batch are discarded, but only the last position of old batch is kept to be used alongside with the new positions.
+In addition, the new positions are sorted to maintain the correct order.
+ The sector data set is provided using the built-in \texttt{Broadcast} feature that distributes a shared value for  all cluster nodes in efficient matter.
+Finally, the detection of sector change is done by an extra filtering transformation using a \texttt{filter} operator, by checking the sectors of consecutive positions in a trajectory to find the ones with a change in the sector assignment.
 

### 2017-03-03 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesSectorChangeDetector.java
code clean up

@@ -55,17 +55,19 @@ public class TrajectoriesSectorChangeDetector {
 
 			// filter the trajectories that contain change in sector between two
 			// consecutive positions
-			Iterable<PositionMessage> trajectoriesList = trajectoryTuple._2;
+			Iterable<PositionMessage> positionsOfTrajectory = trajectoryTuple._2;
 
-			PositionMessage prevTrajectory = null;
-			for (PositionMessage trajectory : trajectoriesList) {
-				if (prevTrajectory != null) {
-					if ((prevTrajectory.isNew() || trajectory.isNew())
-							&& (!trajectory.getSector().equals(prevTrajectory.getSector()))) {
+			PositionMessage prevPosition = null;
+			for (PositionMessage currentPosition : positionsOfTrajectory) {
+				if (prevPosition != null) {
+					if ((prevPosition.isNew() || currentPosition.isNew())
+							&& (!currentPosition.getSector().equals(prevPosition.getSector()))) {
+						// filter a trajectory that contains change in sector
+						// between two consecutive positions
 						return true;
 					}
 				}
-				prevTrajectory = trajectory;
+				prevPosition = currentPosition;
 			}
 
 			return false;
@@ -74,23 +76,23 @@ public class TrajectoriesSectorChangeDetector {
 			// print the sector change transitions
 			StringBuilder output = new StringBuilder(trajectoryTuple._1);
 			output.append(": ");
-			Iterable<PositionMessage> trajectoriesList = trajectoryTuple._2;
+			Iterable<PositionMessage> positionsOfTrajetory = trajectoryTuple._2;
 
-			PositionMessage prevTrajectory = null;
-			for (PositionMessage trajectory : trajectoriesList) {
+			PositionMessage prevPosition = null;
+			for (PositionMessage currentPosition : positionsOfTrajetory) {
 
-				if (prevTrajectory != null) {
+				if (prevPosition != null) {
 
-					if ((prevTrajectory.isNew() || trajectory.isNew())
-							&& (!trajectory.getSector().equals(prevTrajectory.getSector()))) {
+					if ((prevPosition.isNew() || currentPosition.isNew())
+							&& (!currentPosition.getSector().equals(prevPosition.getSector()))) {
 
 						output.append("\n");
-						output.append(prevTrajectory.getSector().getNameAndAirBlock());
+						output.append(prevPosition.getSector().getNameAndAirBlock());
 						output.append("--->");
[…]

### 2017-03-03 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesSectorChangeDetector.java
added more documentation

@@ -9,8 +9,8 @@ import org.apache.flink.streaming.api.datastream.KeyedStream;
 import org.apache.flink.streaming.api.environment.StreamExecutionEnvironment;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
 import de.kdml.bigdatalab.spark_and_flink.flink.utils.FlinkUtils;
 
 /**
@@ -35,9 +35,9 @@ public class TrajectoriesSectorChangeDetector {
 		// the sector between two consecutive trajectories
 		DataStream<String> trajectoriesStreamWithSectors = trajectoriesStream
 				.reduce(new TrajectoriesSectorsReducer(sectorsDataSet.collect())).filter(tuple -> {
-					PositionMessage trajectory = tuple.f1;
-					if (trajectory.getPrevSector() != null
-							&& !trajectory.getSector().equals(trajectory.getPrevSector())) {
+					PositionMessage position = tuple.f1;
+					// check for change in previous and current sector
+					if (position.getPrevSector() != null && !position.getSector().equals(position.getPrevSector())) {
 						return true;
 					}
 					return false;

### 2017-03-03 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesSectorsReducer.java

@@ -6,8 +6,17 @@ import org.apache.flink.api.common.functions.ReduceFunction;
 import org.apache.flink.api.java.tuple.Tuple2;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.SectorUtils;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
+
+/**
+ * This reducer assign previous and current sector for each new tuple of
+ * (id,position message) and discard the old tuple and keeps the new one
+ * 
+ * @author Ehab Qadah
+ * 
+ *         Mar 3, 2017
+ */
 
 public class TrajectoriesSectorsReducer implements ReduceFunction<Tuple2<String, PositionMessage>> {
 
@@ -20,7 +29,8 @@ public class TrajectoriesSectorsReducer implements ReduceFunction<Tuple2<String,
 	}
 
 	@Override
-	public Tuple2<String, PositionMessage> reduce(Tuple2<String, PositionMessage> tuple1, Tuple2<String, PositionMessage> tuple2) {
+	public Tuple2<String, PositionMessage> reduce(Tuple2<String, PositionMessage> tuple1,
+			Tuple2<String, PositionMessage> tuple2) {
 
 		Tuple2<String, PositionMessage> oldTuple = tuple2.f1.isNew() ? tuple1 : tuple2;
 		Tuple2<String, PositionMessage> newTuple = tuple2.f1.isNew() ? tuple2 : tuple1;

### 2017-03-03 paper/paper.tex
[paper] updated Flink's implementation of sector change detection.

@@
 
- The Flink implementation of sector change detection workload, uses the \texttt{FlinkKafkaConsumer09} that read the ADS-B messages stream from Kafka, which used to construct \texttt{DataStream} instance.
-The \texttt{DataStream} records are parsed to tuples of aircraft ID and aircraft's position, by a configured \texttt{map} function.
+ The Flink's implementation of sector change detection workload, uses the \texttt{FlinkKafkaConsumer09} that reads the position messages (i.e., ADS-B) stream from Kafka, which used to construct \texttt{DataStream} instance.
+The \texttt{DataStream}'s records are mapped to tuples of (aircraft ID and aircraft's position), by a configured \texttt{map} transformation.
 The irrelevant tuples are filtered based on the message type using a \texttt{filter} operation.
-Afterward, the tuples with common IDs are grouped together by the \texttt{keyBy} operator.
-Since the Flink streaming processes the stream records as it come, a defined \texttt{reduce} operator assign corresponding sector of new tuple, and the old tuple with same ID is used to retrieve the previous sector to attach it to the new tuple, while the old tuple is discarded.
-Finally, the tuples with different previous and current sector that represent change in sector case, they are filtered using defined \texttt{filter} transformation.
+Afterward, the tuples that shares a common IDs are grouped together by the \texttt{keyBy} operator.
+Since the Flink streaming processes the stream records as it come, a defined \texttt{reduce} operator assign corresponding sector of position on new tuple, while the old tuple with same ID is used to retrieve the previous sector to attach it to the new tuple, while the old tuple is discarded.
+Finally, the tuples (i.e., tuple of (aircraft ID, position message)) with different previous and current sector are filtered using defined \texttt{filter} transformation.
+
+ //TODO: add comparison between spark and Flink approaches 
  \end{itemize}

### 2017-03-04 paper/paper.tex
[paper] added the introducation of preformance results

@@
 \end{abstract}
-Recent years have witnessed a big development improvements of Big Data frameworks such as Apache Spark and Apache Flink. For example, both frameworks support real-time
+Recent years have witnessed a advances in the  development of Big Data frameworks like Apache Spark and Apache Flink. For example, both frameworks support real-time
 streaming processing. While which platform is the best still an open question. 
@@
 Section 4
-provides and compares the performance results of the different workloads. Finally, Section 5 gives the overall conclusion.
+provides and compares the performance results of the different workloads. Last, Section 5 gives the overall conclusion.
 
@@
 \par Apache Spark is an open source project that provide general framework for large-scale data processing \cite{spark}.
-It offers programming API in Java, Scala, Python and R.
+It offers programming APIs in Java, Scala, Python and R.
 Its stack includes set of built-in modules including Spark SQL, MLlib for machine learning, GraphX for graph processing, and Spark Streaming to process real-time data streams.
@@
 \par Apache Flink is an open source project that provide large-scale, distributed stream processing platform, while the batch processing treated as special case of streaming applications (i.e., finite stream)  \cite{flink}.
-It offers programming API in Java and Scala.
+It offers programming APIs in Java and Scala.
 Its software stack includes the core DataStream and DataSet APIs with additional libraries such as Complex event processing for Flink (FlinkCEP), Machine Learning for Flink (FlinkML), and Flink Graph API (Gelly).
@@
 
-This section provides and compares the performance results of the different workloads.
+\par This section presents the results of performance evaluation in Spark (2.0.2) and Flink (1.1.3) for the streaming workloads were described in Section 3.2.
+Specifically, we provide  the measurements of latency and throughput of the benchmarking workloads execution in Spark and Flink.
+We carried the experiment by setup the Kafka stream as described in Section 3.1 on local machine\footnote{ The experiments have been carried  on single machine with Ubuntu 16.04 LTS os, Intel® Core™ i5-6200U CPU @ 2.30GHz × 4  processor, and 16 GB of memory.
+}, and run the different workloads on the same machine (i.e., local mode) to obtain the performance results.
+
+\par We consider the time delay between the streaming time of record in Kafka and the 
+finish processing time in a workload as the platform's processing latency (measured in milliseconds).
+On the other hand, the number of processed stream's records (aircrafts position messages) per second within each streaming workload is the throughput measure.
+Since the Spark Streaming is micro-batch based,  We examine the effect of different batch intervals
+of the Spark Streaming on the performance measurements.
+\subsection{Latency Measurements}
+\subsection{Throughput Measurements}
 

### 2017-03-04 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/App.java
Added logger utils

@@ -9,7 +9,7 @@ import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
 public class App {
 	public static void main(String[] args) {
 		System.out.println("Hello World!");
-
+		Configs.getInstance();
 		DateTimeUtils.parseDateTime("2015/08/13 15:41:37.181", "yyyy/MM/dd HH:mm:ss.SSS");
 
 		Sector sector = Sector.parseSectorData(
@@ -18,5 +18,6 @@ public class App {
 
 		System.out.println(GeoUtils.isPointInPolygon(sector.getPolygon(), -2.8564453125, 54.97761367069628));// true
 		System.out.println(GeoUtils.isPointInPolygon(sector.getPolygon(), 15.3369140625, 62.79493487887006));// false
+		LoggerUtils.logMessage("ehab");
 	}
 }

### 2017-03-04 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/Configs.java

@@ -13,7 +13,6 @@ import java.util.Properties;
 
 public class Configs {
 
-	
 	private Configs(Properties props) {
 		this.props = props;
 	}

### 2017-03-04 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/LoggerUtils.java

@@ -1,0 +1,43 @@
+package de.kdml.bigdatalab.spark_and_flink.common_utils;
+
+import java.io.IOException;
+import java.util.logging.FileHandler;
+import java.util.logging.Formatter;
+import java.util.logging.LogRecord;
+import java.util.logging.Logger;
+
+public class LoggerUtils {
+
+	static {
+
+		warmUp();
+	}
+
+	private static Logger logger;
+
+	public static void logMessage(String message) {
+		logger.info(message);
+
+	}
+
+	private static void warmUp() {
+
+		logger = Logger.getLogger("app-logger");
+
+		// configure the logger with handler and formatter
+		try {
+			FileHandler fileHandler = new FileHandler(Configs.getInstance().getStringProp("logFilePath"));
+			logger.addHandler(fileHandler);
+			fileHandler.setFormatter(new Formatter() {
+
+				@Override
+				public String format(LogRecord record) {
+					// only log the message
+					return record.getMessage() + "\n";
+				}
+			});
+		} catch (SecurityException | IOException e) {
+			System.out.println(e);
+		}
+	}
+}

### 2017-03-04 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesStatistics.java
Added latency measuring

@@ -4,6 +4,7 @@ import org.apache.flink.api.java.tuple.Tuple2;
 import org.apache.flink.streaming.api.datastream.DataStream;
 import org.apache.flink.streaming.api.environment.StreamExecutionEnvironment;
 
+import de.kdml.bigdatalab.spark_and_flink.common_utils.LoggerUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.StatisticsUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 import de.kdml.bigdatalab.spark_and_flink.flink.utils.FlinkUtils;
@@ -41,15 +42,15 @@ public class TrajectoriesStatistics {
 					return newTuple;
 				});
 
-		// showLatecies(trajectoriesStream);
+		showLatecies(trajectoriesStream);
 		trajectoriesStream.print().setParallelism(1);
 
 		env.execute(" Flink Trajectories Statistics Computation");
 	}
 
 	/**
-	 * Show latencies of processed trajectories based on the time deference
-	 * between streamed time and finished time
+	 * Show latencies of processed trajectories based on the time delay between
+	 * streamed time and finished time
 	 * 
 	 * @param trajectoriesStream
 	 */
@@ -58,11 +59,12 @@ public class TrajectoriesStatistics {
 		DataStream<Long> latencies = trajectoriesStream.map(tuple -> {
 
 			long currentTime = System.currentTimeMillis();
-			// get last entered item
-			PositionMessage trajectory = tuple.f1;
-
-			return new Long(currentTime - trajectory.getStreamedTime());
-
+			// Calculate latency
+			PositionMessage position = tuple.f1;
+			Long latency = new Long(currentTime - position.getStreamedTime());
+			// log the latency
+			LoggerUtils.logMessage(latency.toString());
+			return latency;
 		});
 
 		latencies.print().setParallelism(1);

### 2017-03-04 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesStatistics.java

@@ -13,6 +13,7 @@ import org.apache.spark.streaming.api.java.JavaPairDStream;
 import org.apache.spark.streaming.api.java.JavaStreamingContext;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.LoggerUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.StatisticsUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 import de.kdml.bigdatalab.spark_and_flink.spark_project.SparkConfigsUtils;
@@ -37,7 +38,7 @@ public class TrajectoriesStatistics {
 
 		// configure the stream batch interval
 		JavaStreamingContext jssc = new JavaStreamingContext(sc,
-				Durations.seconds(configs.getIntProp("batchDuration")));
+				Durations.milliseconds(configs.getIntProp("batchDuration")));
 
 		// create the trajectories stream
 		JavaPairDStream<String, Iterable<PositionMessage>> trajectories = TrajectoriesStreamUtils
@@ -70,12 +71,16 @@ public class TrajectoriesStatistics {
 	private static void printLatencies(JavaPairDStream<String, Iterable<PositionMessage>> runningTrajectories) {
 		JavaDStream<Long> latencies = runningTrajectories.map(trajectoryTuple -> {
 
+			// find the average latency for all positions in a trajectory
 			long currentTime = System.currentTimeMillis(), sumLatency = 0, count = 0;
 
-			Iterable<PositionMessage> trajectoriesList = trajectoryTuple._2;
-			for (PositionMessage trajectory : trajectoriesList) {
-				if (trajectory.isNew()) {
-					sumLatency += currentTime - trajectory.getStreamedTime();
+			Iterable<PositionMessage> positionsInTrajectory = trajectoryTuple._2;
+			for (PositionMessage position : positionsInTrajectory) {
+				if (position.isNew()) {
+
+					long latency = currentTime - position.getStreamedTime();
+					LoggerUtils.logMessage(String.valueOf(latency));
+					sumLatency += latency;
 					count++;
 				}
 			}
@@ -85,7 +90,8 @@ public class TrajectoriesStatistics {
 			return latency > 0;
 		});
 
-		// latencies.print();
+		// show latencies
+		latencies.print();
 	}
 
 	/**

### 2017-03-04 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/Configs.java
added throughput measure in Flink

@@ -59,4 +59,9 @@ public class Configs {
 		return Integer.parseInt(this.props.getProperty(propName));
 	}
 
+	public double getDoubleProp(String propName) {
+
+		return Double.parseDouble(this.props.getProperty(propName));
+	}
+
 }

### 2017-03-04 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesStatistics.java

@@ -1,8 +1,13 @@
 package de.kdml.bigdatalab.spark_and_flink.flink.datacorn;
 
+import org.apache.flink.api.common.functions.FoldFunction;
 import org.apache.flink.api.java.tuple.Tuple2;
 import org.apache.flink.streaming.api.datastream.DataStream;
 import org.apache.flink.streaming.api.environment.StreamExecutionEnvironment;
+import org.apache.flink.streaming.api.functions.AssignerWithPeriodicWatermarks;
+import org.apache.flink.streaming.api.functions.sink.SinkFunction;
+import org.apache.flink.streaming.api.watermark.Watermark;
+import org.apache.flink.streaming.api.windowing.time.Time;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.LoggerUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.StatisticsUtils;
@@ -42,10 +47,64 @@ public class TrajectoriesStatistics {
 					return newTuple;
 				});
 
-		showLatecies(trajectoriesStream);
+		// showLatecies(trajectoriesStream);
 		trajectoriesStream.print().setParallelism(1);
 
+		showThroughput(trajectoriesStream);
 		env.execute(" Flink Trajectories Statistics Computation");
+	}
+
+	private static void showThroughput(DataStream<Tuple2<String, PositionMessage>> trajectoriesStream) {
+		DataStream<Tuple2<String, PositionMessage>> trajectoriesStreamWithTime = trajectoriesStream
+
+				.assignTimestampsAndWatermarks(new AssignerWithPeriodicWatermarks<Tuple2<String, PositionMessage>>() {
+
+					/**
+					 *
+					 * @link{https://ci.apache.org/projects/flink/flink-docs-release-1.2/dev/event_timestamps_watermarks.html}
+					 */
+					private static final long serialVersionUID = 1L;
+					private final long maxTimeLag = 1000; // 5 seconds
+
+					@Override
+					public long extractTimestamp(Tuple2<String, PositionMessage> element,
+							long previousElementTimestamp) {
+						return element.f1.getStreamedTime();
+					}
+
+					@Override
+					public Watermark getCurrentWatermark() {
+						// return the watermark as current time minus the
+						// maximum time lag
+						return new Watermark(System.currentTimeMillis() - maxTimeLag);
+					}
+				});
+
[…]

### 2017-03-04 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesStatistics.java

@@ -12,11 +12,14 @@ import org.apache.spark.streaming.api.java.JavaDStream;
 import org.apache.spark.streaming.api.java.JavaPairDStream;
 import org.apache.spark.streaming.api.java.JavaStreamingContext;
 
+import com.google.common.collect.Iterators;
+
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.LoggerUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.StatisticsUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 import de.kdml.bigdatalab.spark_and_flink.spark_project.SparkConfigsUtils;
+import scala.Tuple2;
 
 /**
  * 
@@ -36,9 +39,9 @@ public class TrajectoriesStatistics {
 		// configure spark streaming context
 		JavaSparkContext sc = SparkConfigsUtils.getSparkContext("Trajectories Statistics Computation ");
 
+		long batchTime = configs.getIntProp("batchDuration");
 		// configure the stream batch interval
-		JavaStreamingContext jssc = new JavaStreamingContext(sc,
-				Durations.milliseconds(configs.getIntProp("batchDuration")));
+		JavaStreamingContext jssc = new JavaStreamingContext(sc, Durations.milliseconds(batchTime));
 
 		// create the trajectories stream
 		JavaPairDStream<String, Iterable<PositionMessage>> trajectories = TrajectoriesStreamUtils
@@ -51,16 +54,37 @@ public class TrajectoriesStatistics {
 		JavaPairDStream<String, Iterable<PositionMessage>> runningTrajectories = trajectories
 				.updateStateByKey(updateTrajectoriesAndComputeStatistics);
 
-		printLatencies(runningTrajectories);
+		// printLatencies(runningTrajectories);
 
 		runningTrajectories.print(1000);
 
+		showThroughput(batchTime, runningTrajectories);
 		jssc.start();
 		try {
 			jssc.awaitTermination();
 		} catch (InterruptedException e) {
 			System.out.println("Error:" + e.getMessage());
 		}
+	}
+
+	private static void showThroughput(long batchTime,
+			JavaPairDStream<String, Iterable<PositionMessage>> runningTrajectories) {
+		runningTrajectories.foreachRDD((rdd, time) -> {
+
[…]

### 2017-03-05 paper/paper.tex
[paper] update stream setup section

@@
 \begin{flushleft}
+
+ \thispagestyle{empty}
 \centering\LARGE {\bf Comparative Evaluation of Spark and Flink Stream Processing}
 
+\vspace{14pt}
+
+\centering
+ Lab: Data Science and Big Data MA-INF 4306,
+ Winter Term 2016/2017, University of Bonn
+
+\vspace{14pt}
+\centering
+ Ehab Qadah\\
  Supervisor: PD Dr. Michael Mock
-\thispagestyle{empty}
+\rule{\textwidth}{1pt}
 \end{flushleft}
@@
 \section{Experiment Setup and Implementation}
-This section describes the data stream setup, design of evaluation streaming workloads, and implementation details in Spark and Flink.
+This section describes the data stream setup, the design of evaluation streaming workloads, and the implementation details in Spark and Flink.
 
 \subsection{Data Stream Setup}
-Our streaming workloads read input data stream from Kafka.
-In order to simulate real-world uses cases we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircraft's position over time, comprise 22 fields of data such as aircraft ID, date message generated, longitude, latitude, and altitude.
-\par In our experiment a data stream producer component reads the ADS-B datasets, then publish it to Kafka cluster to be consumed by the workloads in Spark and Flink.
-Figure 1 illustrates the data producer and the Kafka cluster setup.
-The data stream is portioned into four portions over two server, while the data producer publish the stream records randomly to Kafka partitions.
+The evaluation streaming workloads read an input data stream from Kafka, in order to simulate real-world use case we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircraft's position over time, comprise 22 fields of data such as aircraft ID, date message generated, longitude, latitude, and altitude etc.
+\par In our experiment a data stream producer component reads the ADS-B message datasets, then publish them to Kafka cluster to be consumed by the workloads in Apache Spark and Apache Flink.
+Figure 4 illustrates the data producer and the Kafka cluster setup in all experiments.
+The data stream is portioned into four portions over two server, while the data producer publish the stream records randomly to Kafka partitions, and all ADS-B messages are published to same topic, which is \texttt{datacorn}.
+Moreover, the data producer attach the streaming time at the ADS-B message line before publishing it to the Kafka topic, to be used in the latency calculation.
 
@@
 \end{itemize}
-We present the description of workloads and the relation with these aspects, and th implementation details of Spark and Flink solutions in Section 3.2.1 \& 3.2.1.
-Afterward, In Section 4 we discuss the performance evaluation and analysis of workloads.
+We present the description of workloads and the relation with these aspects, also the implementation details of Spark and Flink solutions in Section 3.2.1 \& 3.2.1.
+Afterward, In Section 4 we discuss the performance evaluation and analysis of these workloads in both frameworks.
 

### 2017-03-05 paper/paper.tex
[paper] added discussion  points regarding the implementation details

@@
 
-//TODO: discuss differences between Flink and Spark.
+\end{itemize}
 
-\end{itemize}
+\par In summary, Flink  handles  the parallel consumers of Kafka stream implicitly, while a multiple \texttt{DStream} must be created in Spark Streaming and union them to have parallel Kafka receivers.
+The Flink's streaming processing is stateful that means there is no need to manage the stream state as the case in Spark streaming.
+A sort action is required in Spark Streaming to preserve the correct  order of the position messages inside the state update function.
 
@@
 
- //TODO: add comparison between spark and Flink approaches 
  \end{itemize}
+
+ \par To sum up, the same differences between Flink and Spark Streaming regarding the state management and parallel  consumers of Kafka stream as was discussed in Section 3.2.1.
+Another key difference between them, is how  to provide the global data model (e.g., sectors) in a streaming workload.
+Spark Streaming offers the \texttt{Broadcast} feature to handle this issue in efficient way, while the global data model is manually provided  to the stream operations in Flink.
 \section{Performance Results}

### 2017-03-05 paper/paper.tex
updated header page

@@
 
- \thispagestyle{empty}
-\centering\LARGE {\bf Comparative Evaluation of Spark and Flink Stream Processing}
-
-\vspace{14pt}
+\centering { \large \bf Comparative Evaluation of Spark and Flink Stream Processing}
+\vspace{4pt}
 
 \centering
- Lab: Data Science and Big Data MA-INF 4306,
+ Lab Report: Data Science and Big Data MA-INF 4306,
  Winter Term 2016/2017, University of Bonn
 
-\vspace{14pt}
+\vspace{4pt}
 \centering
  Ehab Qadah\\
- Supervisor: PD Dr. Michael Mock
-\rule{\textwidth}{1pt}
+ Supervisor: PD Dr. Michael Mock.
+ \vspace{4pt}
+
+ \today
 \end{flushleft}
-\newpage
-\tableofcontents
-
-\newpage
 
@@
 streaming processing. While which platform is the best still an open question. 
-In this paper, we provide a performance comparison of streaming data processing in Apache Spark and Apache Flink, by measuring different performance metrics, namely latency and throughput.
+In this report, we provide a performance comparison of streaming data processing in Apache Spark and Apache Flink, by measuring different performance metrics, namely latency and throughput.
 In addition, we cover some key aspects of streaming data applications and how they are handled in both frameworks.
@@
 
-\par The purpose of this paper is to provide a comparative experimental evaluation of throughput, latency (i.e., performance) of stream processing in Apache Spark and Apache Flink.
+\par The purpose of this report is to provide a comparative experimental evaluation of throughput, latency (i.e., performance) of stream processing in Apache Spark and Apache Flink.
 We use a datasets of airplanes trajectories provided in the context of Datacron project\footnote{\url{http://www.datacron-project.eu/}}, in order to simulate real-world use cases.
@@
 
-\par The remainder of this paper is organized as follows.
+\par The remainder of this report is organized as follows.
 In Section 2, we present the main characteristics, APIs, and main data abstraction of Apache Spark and Flink.
@@
   \centering
-    \includegraphics[width=\textwidth, height=.4\textheight]{flink_engine.png}
-     \caption{ An example of data flow graph \cite{flink}.}
+    \includegraphics[width=\textwidth, height=.3\textheight]{flink_engine.png}
+     \caption{ An example of data flow graph in Flink \cite{flink}.}
 \end{figure} 

### 2017-03-05 flink_project/src/main/java/kafka/datacorn/AircraftPositionsStreamGenerator.java
updated ADS-B messages producer

@@ -1,0 +1,61 @@
+package kafka.datacorn;
+
+import java.io.BufferedReader;
+import java.io.FileReader;
+
+import org.apache.flink.streaming.api.functions.source.SourceFunction;
+
+import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.StreamRecord;
+
+/**
+ * ADS-B messages stream source
+ * 
+ * @author Ehab Qadah
+ * 
+ *         Mar 5, 2017
+ */
+public class AircraftPositionsStreamGenerator implements SourceFunction<String> {
+
+	private static final int LINE_SLIDE_TIME_MS = 0;
+
+	private static final long serialVersionUID = 2174904787118597072L;
+	private static Configs configs = Configs.getInstance();
+	boolean running = true;
+	int i = 0;
+
+	@Override
+	public void run(SourceContext<String> ctx) throws Exception {
+
+		while (running) {
+
+			String[] adsbFiles = configs.getStringProp("adsbFiles").split(",");
+			for (String file : adsbFiles) {
+
+				System.out.println("streaming of " + file);
+				i++;
+				try (BufferedReader br = new BufferedReader(new FileReader(file))) {
+					String messageLine;
+					while ((messageLine = br.readLine()) != null) {
+
+						// append streaming time
+						messageLine = new StreamRecord(messageLine).toString();
+						ctx.collect(messageLine);
+						// Thread.sleep(LINE_SLIDE_TIME_MS);
+					}
+				} catch (Exception e) {
+
+					System.out.println("exp" + e);
+				}
+			}
+			// running = false;
+
+		}
+	}
+
+	@Override
+	public void cancel() {
+		running = false;
+	}
+
+}

### 2017-03-05 flink_project/src/main/java/kafka/datacorn/DatacornKafkaStreamProducer.java

@@ -1,7 +1,5 @@
 package kafka.datacorn;
 
-import java.io.BufferedReader;
-import java.io.FileReader;
 import java.io.Serializable;
 import java.util.Properties;
 
@@ -9,17 +7,15 @@ import org.apache.flink.api.common.restartstrategy.RestartStrategies;
 import org.apache.flink.streaming.api.TimeCharacteristic;
 import org.apache.flink.streaming.api.datastream.DataStream;
 import org.apache.flink.streaming.api.environment.StreamExecutionEnvironment;
-import org.apache.flink.streaming.api.functions.source.SourceFunction;
 import org.apache.flink.streaming.connectors.kafka.FlinkKafkaProducer09;
 import org.apache.flink.streaming.connectors.kafka.partitioner.KafkaPartitioner;
 import org.apache.flink.streaming.util.serialization.SimpleStringSchema;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.StreamRecord;
 
 /**
- * This kafka's stream of lines producer using FlinkKafkaProducer09 it writes a
- * new random line every 1 second
+ * This kafka's stream of lines producer using FlinkKafkaProducer09 it writes
+ * the ADS-B messages
  * 
  * @author Ehab Qadah
  * 
@@ -39,18 +35,18 @@ public class DatacornKafkaStreamProducer {
 		env.setParallelism(4);
 
 		// add a simple source which is writing some strings
-		DataStream<String> messageStream = env.addSource(new SimpleStringGenerator());
+		DataStream<String> messageStream = env.addSource(new AircraftPositionsStreamGenerator());
 
 		Properties producerConfig = new Properties();
 		producerConfig.put("bootstrap.servers", configs.getStringProp("bootstrap.servers"));
 		// // write stream to Kafka
 		messageStream.addSink(new FlinkKafkaProducer09<String>(configs.getStringProp("topicId"),
-				new SimpleStringSchema(), producerConfig, new TestPartitioner()));
+				new SimpleStringSchema(), producerConfig, new RandomPartitioner()));
 
 		env.execute("kafka stream ");
 	}
 
-	public static class TestPartitioner extends KafkaPartitioner<String> implements Serializable {
+	public static class RandomPartitioner extends KafkaPartitioner<String> implements Serializable {
 		private static final long serialVersionUID = 1627268846962918126L;
 
[…]

### 2017-03-06 paper/paper.tex
[paper] minor updates
added conclusion

@@
 \usepackage{listings}
+\usepackage{subcaption}
 %opening
@@
 \end{abstract}
-Recent years have witnessed a advances in the  development of Big Data frameworks like Apache Spark and Apache Flink. For example, both frameworks support real-time
+Recent years have witnessed a advances in the  development of Big Data frameworks like Apache Spark and Apache Flink. For instance, both frameworks support real-time
 streaming processing. While which platform is the best still an open question. 
@@
 
-\par The processing models of Big Data are Batch processing and Stream processing, Batch processing is a way to handle large, finite volumes of data (e.g., processing of historical transaction records), is more concerned with throughput.
-While the Stream processing  is a method to manage fast and continuously incoming data (real-time), process it as it arrives (e.g., credit card fraud detection and network monitoring applications), where the low latency is required.
+\par The processing models of Big Data are Batch processing and Stream processing, Batch processing involves handling  large, finite volumes of data (e.g., processing of historical transaction records), is more concerned with throughput.
+While the Stream processing  is a method to manage fast and continuously incoming data (real-time), process it as it arrives to the system (e.g., credit card fraud detection and network monitoring applications), where the low latency is required.
 The variety of data (structured and semi-structured) is included in the two processing models.
@@
 \subsection{Latency Measurements}
+
+\begin{figure}[h]
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{latency1.png}
+  \caption{Latency of Statistics Computation}
+
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{latency1.png}
+\caption{Latency of Sector Change Detection }
+
+\end{subfigure}
+\caption{Latency Results}
+\label{fig:fig}
+\end{figure}
 \subsection{Throughput Measurements}
@@
 \section{Conclusion}
-This section gives the overall conclusion.
+
+In this report, we present an experimental evaluation of the streaming processing performance and capabilities of Apache Spark and Apache Flink.
+We carried our experiment over aircraft potions provided in the context of EU Datacorn  project, we developed two streaming workloads to cover general aspects of real-time stream processing tasks, and we measured the latency and throughput of those workloads within the solution of Spark and Flink.
+Results show that Flink outperforms Spark Streaming in term of processing latency.
+Furthermore, Flink's processing model is well-suited to  the stream processing tasks, there is no need to manage the Stream's state and batch interval as it is in Spark streaming.
+Additionally, Flink operates over the individual record of the data stream, not like Spark Streaming that processes the input data stream as micro-batches.
+In Summary, Flink is true stream processing framework.
+
+In future work, we plan to carry our experiments over cluster of computers not just only on single computer.
+In addition, we will cover more aspects of streaming processing and developing more evaluation workloads.
 

### 2017-03-06 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesStatistics.java
measure the latency in Flink

@@ -40,17 +40,19 @@ public class TrajectoriesStatistics {
 					// compute & aggregate statistics for the new trajectory
 					// based on the statistics of old trajectory
 					// discard the old trajectory and just keep the new one
-					Tuple2<String, PositionMessage> oldTuple = tuple2.f1.isNew() ? tuple1 : tuple2;
-					Tuple2<String, PositionMessage> newTuple = tuple2.f1.isNew() ? tuple2 : tuple1;
+					Tuple2<String, PositionMessage> oldTuple = tuple2.f1.getStreamedTime() < tuple1.f1.getStreamedTime()
+							? tuple2 : tuple1;
+					Tuple2<String, PositionMessage> newTuple = tuple2.f1.getStreamedTime() < tuple1.f1.getStreamedTime()
+							? tuple1 : tuple2;
 					StatisticsUtils.computeStatistics(oldTuple.f1, newTuple.f1);
 					newTuple.f1.setNew(false);
 					return newTuple;
 				});
 
-		// showLatecies(trajectoriesStream);
+		showLatecies(trajectoriesStream);
 		trajectoriesStream.print().setParallelism(1);
 
-		showThroughput(trajectoriesStream);
+		// showThroughput(trajectoriesStream);
 		env.execute(" Flink Trajectories Statistics Computation");
 	}
 
@@ -117,10 +119,9 @@ public class TrajectoriesStatistics {
 	private static void showLatecies(DataStream<Tuple2<String, PositionMessage>> trajectoriesStream) {
 		DataStream<Long> latencies = trajectoriesStream.map(tuple -> {
 
-			long currentTime = System.currentTimeMillis();
 			// Calculate latency
 			PositionMessage position = tuple.f1;
-			Long latency = new Long(currentTime - position.getStreamedTime());
+			Long latency = new Long(System.currentTimeMillis() - position.getStreamedTime());
 			// log the latency
 			LoggerUtils.logMessage(latency.toString());
 			return latency;

### 2017-03-06 flink_project/src/main/java/kafka/datacorn/AircraftPositionsStreamGenerator.java

@@ -17,7 +17,7 @@ import de.kdml.bigdatalab.spark_and_flink.common_utils.data.StreamRecord;
  */
 public class AircraftPositionsStreamGenerator implements SourceFunction<String> {
 
-	private static final int LINE_SLIDE_TIME_MS = 0;
+	private static final int LINE_SLIDE_TIME_MS = 10;
 
 	private static final long serialVersionUID = 2174904787118597072L;
 	private static Configs configs = Configs.getInstance();
@@ -41,7 +41,7 @@ public class AircraftPositionsStreamGenerator implements SourceFunction<String> 
 						// append streaming time
 						messageLine = new StreamRecord(messageLine).toString();
 						ctx.collect(messageLine);
-						// Thread.sleep(LINE_SLIDE_TIME_MS);
+						Thread.sleep(LINE_SLIDE_TIME_MS);
 					}
 				} catch (Exception e) {
 

### 2017-03-06 paper/paper.tex
[paper] added latency results

@@
 
-\par The are some performance evaluation of Apache Spark and Apache Flink and other framework such as Yahoo Streaming Benchmark \cite{yahoo}.
+\par The are some performance evaluation of Apache Spark and Apache Flink  such as Yahoo Streaming Benchmark \cite{yahoo}.
 We believe that we still need to carry our evaluation for two reasons: \begin{enumerate*}[label=(\roman*)]
@@
 of the Spark Streaming on the performance measurements.
-\subsection{Latency Measurements}
+\subsection{Latency Results}
+
+Figure 5 shows the results of latency measurements of the streaming workloads in Spark and Flink.
+The observed values present the mean value of multiple trails and measurements.
+Measurements for different batch intervals are given in Spark Streaming,  if we decrease  will not get better latency but it is even getting  worse.
+Results show that Flink outperform Spark streaming in term of processing latency in the both evaluation workloads.
 
@@
 \end{subfigure}
-\caption{Latency Results}
+\caption{Latency Measurements}
 \label{fig:fig}
 \end{figure}
-\subsection{Throughput Measurements}
+\subsection{Throughput Results}
 

### 2017-03-06 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/data/PositionMessage.java
measure latency in Spark

@@ -26,7 +26,7 @@ public class PositionMessage implements Serializable {
 	private Sector sector;
 	private Sector prevSector;
 	private Boolean isNew = null;
-	private long streamedTime;
+	private long streamedTime, finishProcessingTime;
 	private Double ditance = null;
 	private Double speed = null;
 	private Double acceleration;
@@ -163,6 +163,14 @@ public class PositionMessage implements Serializable {
 
 	public void setPrevSector(Sector prevSector) {
 		this.prevSector = prevSector;
+	}
+
+	public long getFinishProcessingTime() {
+		return finishProcessingTime;
+	}
+
+	public void setFinishProcessingTime(long finishProcessingTime) {
+		this.finishProcessingTime = finishProcessingTime;
 	}
 
 }

### 2017-03-06 flink_project/src/main/java/kafka/datacorn/AircraftPositionsStreamGenerator.java

@@ -17,7 +17,7 @@ import de.kdml.bigdatalab.spark_and_flink.common_utils.data.StreamRecord;
  */
 public class AircraftPositionsStreamGenerator implements SourceFunction<String> {
 
-	private static final int LINE_SLIDE_TIME_MS = 10;
+	private static final int LINE_SLIDE_TIME_MS = 100;
 
 	private static final long serialVersionUID = 2174904787118597072L;
 	private static Configs configs = Configs.getInstance();

### 2017-03-06 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesStatistics.java

@@ -54,11 +54,11 @@ public class TrajectoriesStatistics {
 		JavaPairDStream<String, Iterable<PositionMessage>> runningTrajectories = trajectories
 				.updateStateByKey(updateTrajectoriesAndComputeStatistics);
 
-		// printLatencies(runningTrajectories);
+		printLatencies(runningTrajectories);
 
 		runningTrajectories.print(1000);
 
-		showThroughput(batchTime, runningTrajectories);
+		// showThroughput(batchTime, runningTrajectories);
 		jssc.start();
 		try {
 			jssc.awaitTermination();
@@ -102,7 +102,7 @@ public class TrajectoriesStatistics {
 			for (PositionMessage position : positionsInTrajectory) {
 				if (position.isNew()) {
 
-					long latency = currentTime - position.getStreamedTime();
+					long latency = position.getFinishProcessingTime() - position.getStreamedTime();
 					LoggerUtils.logMessage(String.valueOf(latency));
 					sumLatency += latency;
 					count++;
@@ -159,6 +159,7 @@ public class TrajectoriesStatistics {
 				// Compute statistics for each position message based on
 				// the aggregated values of previous position statistics
 				StatisticsUtils.computeStatistics(lastOldPosition, trajectory);
+				trajectory.setFinishProcessingTime(System.currentTimeMillis());
 				lastOldPosition = trajectory;
 
 				/**

### 2017-03-07 paper/paper.tex
[paper] updated the abstract and introduction

@@
 \end{abstract}
-Recent years have witnessed a advances in the  development of Big Data frameworks like Apache Spark and Apache Flink. For instance, both frameworks support real-time
-streaming processing. While which platform is the best still an open question. 
-In this report, we provide a performance comparison of streaming data processing in Apache Spark and Apache Flink, by measuring different performance metrics, namely latency and throughput.
-In addition, we cover some key aspects of streaming data applications and how they are handled in both frameworks.
+Recent years have witnessed advances in the  development of Big Data frameworks like Apache Spark and Apache Flink. For instance, both frameworks support real-time data 
+stream processing. The question of which platform is superior to the other still open. 
+In this report, we provide a performance comparison of streaming data processing in Apache Spark and Apache Flink by measuring different performance metrics (latency and throughput).
+In addition, we cover some key aspects of streaming data applications and how they are handled in the two frameworks.
 The experiment is done using multiple streaming workloads over real-world datasets.
@@
 
-\par In recent years, Big Data is considered as one of the most appealing topics in Information Technology, it has gained increasing attention of professionals in industry and academia, as a result of the fact that the amount of data we generate is increasing exponentially \cite{idc} that generated from social media platforms, mobile phones, IoT and finical transactions, etc.
-Moreover, recent years showed a rapid development of Big Data frameworks such as Apache Spark and Apache Flink that provide more and efficient data processing capabilities.
- \par The Big Data is a method to process not just only large datasets (volume), but also the data that arrives in high rate (velocity), and it comes in all kinds of data format (variety) (i.e., 3Vs of Big Data (Volume, Velocity, Variety)) \cite{svs}.
+\par In recent years, Big Data is considered as one of the most appealing topics in Information Technology, it has gained increasing attention of professionals in industry and academia.
+As result of the fact that the amount of data we generate is increasing exponentially \cite{idc}.
+Different sources such as  social media platforms, mobile phones, IoT and finical transactions, etc.
+create large volumes of data.
+Moreover, recent years have shown a rapid development of Big Data frameworks such as Apache Spark and Apache Flink that provide efficient data processing functionalities.
+ \par Big Data is a term that does not only describe large datasets (volume), but also the data that arrives in high rate (velocity), which comes in all kinds of data format (variety) (i.e., 3Vs of Big Data (Volume, Velocity, Variety)) \cite{svs}.
 
-\par The processing models of Big Data are Batch processing and Stream processing, Batch processing involves handling  large, finite volumes of data (e.g., processing of historical transaction records), is more concerned with throughput.
-While the Stream processing  is a method to manage fast and continuously incoming data (real-time), process it as it arrives to the system (e.g., credit card fraud detection and network monitoring applications), where the low latency is required.
-The variety of data (structured and semi-structured) is included in the two processing models.
+\par The processing models of Big Data are Batch processing and Stream processing.
+Batch processing involves handling  large, finite volumes of data (e.g., processing of historical transaction records), and is more concerned with throughput.
+While Stream processing  is a method to manage fast and continuously incoming data (real-time), as it arrives to the system (e.g., credit card fraud detection and network monitoring applications).
+Where low latency is crucial for stream processing systems.
+Both processing models are capable of operating with structured or semi-structured data (i.e., variety).
 
-\par The are some performance evaluation of Apache Spark and Apache Flink  such as Yahoo Streaming Benchmark \cite{yahoo}.
+\par There exist some performance evaluations of Apache Spark and Apache Flink  such as Yahoo Streaming Benchmark \cite{yahoo}.
 We believe that we still need to carry our evaluation for two reasons: \begin{enumerate*}[label=(\roman*)]
@@
 
-\par The purpose of this report is to provide a comparative experimental evaluation of throughput, latency (i.e., performance) of stream processing in Apache Spark and Apache Flink.
-We use a datasets of airplanes trajectories provided in the context of Datacron project\footnote{\url{http://www.datacron-project.eu/}}, in order to simulate real-world use cases.
-Using multiple streaming data workloads.
-Also we try to cover general aspects of streaming processing tasks (later details in Section 3) and how they can be solved in each platform.
-Informally,  this work aims to help the developers to determine which platform to choose in production for different use cases.
+\par The purpose of this report is to provide a comparative experimental evaluation of throughput and latency of stream processing in Apache Spark and Apache Flink.
+In order to simulate real-world use cases, we use datasets of airplane trajectories provided by the  Datacron project\footnote{\url{http://www.datacron-project.eu/}} to develop multiple streaming data processing workloads.
+Beside the performance comparison of stream processing of the two frameworks, we try to cover general aspects of stream processing tasks (see Section 3) and how they can be solved in each platform.
+Informally,  this work aims to help the developers to determine which platform to use for different use cases.
 
 \par The remainder of this report is organized as follows.
-In Section 2, we present the main characteristics, APIs, and main data abstraction of Apache Spark and Flink.
-Section 3  presents the general aspects of streaming processing, the design and implementation of experimental workloads.
+In Section 2, we present the main characteristics, programming APIs, and main data abstraction of Apache Spark and Apache Flink.
+Section 3  presents the general aspects of streaming processing and the implementation of experimental workloads.
 Section 4
-provides and compares the performance results of the different workloads. Last, Section 5 gives the overall conclusion.
+compares the performance results of the different workloads. Section 5 gives the overall conclusion.
 

### 2017-03-07 paper/paper.tex
[paper] updated Spark and Flink sections

@@
 
-\par Apache Spark is an open source project that provide general framework for large-scale data processing \cite{spark}.
+\par Apache Spark is an open source project that provides a general framework for large-scale data processing \cite{spark}.
 It offers programming APIs in Java, Scala, Python and R.
-Its stack includes set of built-in modules including Spark SQL, MLlib for machine learning, GraphX for graph processing, and Spark Streaming to process real-time data streams.
+Its stack includes a set of built-in modules including Spark SQL, MLlib for machine learning, GraphX for graph processing, and Spark Streaming to process real-time data streams.
 It can access different data sources including Hadoop Distributed File System (HDFS), Apache Cassandra database, Apache HBase database etc.
 
-\par The main data abstraction in Spark core is the Resilient Distributed Datasets (RDDs), which is in-memory collection of elements partitioned across cluster of computers that can be processed in parallel \cite{rdd}.
-RDDs supports two categories of operations: transformations that drive new RDDs from the operated ones (e.g., map), and actions which return a certain value (e.g., count).
-On the other hand, the Discretized Stream (DStream) is the data abstraction in Spark Streaming, which is series of RDDs that represent an input stream of data \cite{spark_streaming}.
-In other words, the Spark Streaming is batch-based processing model, which process the continuous stream of data by divided into micro-batches to be processed by the Spark engine.
+\par The main data abstraction of Spark core is the Resilient Distributed Datasets (RDDs), which is in-memory collection of elements partitioned across cluster of computers that can be processed in parallel \cite{rdd}.
+RDDs support two categories of operations: transformations that drive new RDDs from the operated ones (e.g., \texttt{map}, \texttt{filter} and \texttt{groupByKey}), and actions which return a certain value (e.g., \texttt{count} and \texttt{collect}).
+Additionally, Discretized Stream (DStream) is the data abstraction in Spark Streaming, which is a series of RDDs that represent an input stream of data \cite{spark_streaming}.
+In other words, the Spark Streaming is a batch-based processing model, which processes the continuous stream of data by divided it into micro-batches that are  processed by the Spark engine.
 Figure 1 illustrates the general process flow of Spark Streaming.
@@
 
-\par Apache Flink is an open source project that provide large-scale, distributed stream processing platform, while the batch processing treated as special case of streaming applications (i.e., finite stream)  \cite{flink}.
+\par Apache Flink is an open source project that provides a large-scale, distributed stream processing platform \cite{flink}.
+Flink treats the batch processing as special case of streaming applications (i.e., finite stream).
 It offers programming APIs in Java and Scala.
-Its software stack includes the core DataStream and DataSet APIs with additional libraries such as Complex event processing for Flink (FlinkCEP), Machine Learning for Flink (FlinkML), and Flink Graph API (Gelly).
+Its software stack includes the core DataStream and DataSet APIs with additional libraries such as Complex event processing for Flink (FlinkCEP), Machine Learning for Flink (FlinkML)  and Flink Graph API (Gelly).
 
@@
 
-\par The Flink's core is a distributed streaming dataflow engine,  each Flink program is represented by a data flow graph (i.e., directed acyclic graph - DAG) that executed by the Flink's engine \cite{flink_paper}.
+\par The Flink's core is a distributed streaming dataflow engine,  with each Flink program is represented by a data flow graph (i.e., directed acyclic graph - DAG) that executed by the Flink's engine \cite{flink_paper}.
 The data flow graphs are composed of stateful, parallel operations and intermediate data stream partitions.
-Figure 2 shows a data flow graph of Flink program.
+Figure 2 shows a data flow graph of a Flink program.
 

### 2017-03-07 paper/paper.tex
[paper] updates

@@
 \subsection{Apache Kafka}
-\par Apache Kafka is scalable, fault-tolerant distributed publish/subscribe streaming framework \cite{kafka}.
-It manges the stream records in different categorizes (i.e., topics) that portioned and distributed over the servers of the Kafka cluster.
-It allows the data producers to publish stream of records to certain Kafka topic.
-While the consumer applications can subscribe to one or more topic to read the data stream.
+\par Apache Kafka is scalable, fault-tolerant and distributed streaming framework \cite{kafka}. It allows to publish and subscribe to data streams.
+Kafka manages the stream records in different categories (i.e., topics) that partitioned and distributed over the servers of the Kafka cluster.
+It allows the data producers to publish a stream of records to one or more  Kafka topic.
+Additionally, consumer applications can subscribe to one or more topic to read data streams.
 The consumers are grouped into groups based on the group name that allows Kafka to distribute and balance the stream partitions among the  members of a certain group for the sake of scalability.
-Figure 3 shows how 4 partitions of stream is distributed differentially between two groups of consumers based on the number of members.
+Figure 3 shows how 4 partitions of a stream are distributed between two groups of consumers based on the number of members.
 Apache Kafka has been widely  adopted, for example, 
-Spark and Flink can receive data stream from Kafka.
+Apache Spark and Apache Flink can ingest Kafka streams.
 
@@
     \includegraphics[width=\textwidth, height=.3\textheight]{kafka_groups.png}
-     \caption{ Distribution of stream partitions for consumer groups in Kafka \cite{kafka}.}
+     \caption{ Distribution of a stream partitions for consumer groups in Kafka \cite{kafka}.}
 \end{figure}
@@
 \section{Experiment Setup and Implementation}
-This section describes the data stream setup, the design of evaluation streaming workloads, and the implementation details in Spark and Flink.
+This section describes the data stream setup, the design of evaluation streaming workloads and the solution's implementation details of  Apache Spark and Apache Flink.
 
 \subsection{Data Stream Setup}
-The evaluation streaming workloads read an input data stream from Kafka, in order to simulate real-world use case we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircraft's position over time, comprise 22 fields of data such as aircraft ID, date message generated, longitude, latitude, and altitude etc.
-\par In our experiment a data stream producer component reads the ADS-B message datasets, then publish them to Kafka cluster to be consumed by the workloads in Apache Spark and Apache Flink.
+The evaluation streaming workloads read an input data stream from Kafka.
+In order to simulate real-world use case we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircraft's position over time.
+The message line comprises 22 fields of data such as aircraft ID, date message generated, longitude, latitude and altitude.
+\par In our experiment a data stream producer component reads the ADS-B message datasets, then publishes them to the Kafka cluster to be consumed by the workloads in Apache Spark and Apache Flink.
 Figure 4 illustrates the data producer and the Kafka cluster setup in all experiments.
-The data stream is portioned into four portions over two server, while the data producer publish the stream records randomly to Kafka partitions, and all ADS-B messages are published to same topic, which is \texttt{datacorn}.
-Moreover, the data producer attach the streaming time at the ADS-B message line before publishing it to the Kafka topic, to be used in the latency calculation.
+The data stream is divided into four portions over two server.
+The data producer publishes the stream records randomly to a Kafka partition and all ADS-B messages are published to the same topic (\texttt{datacorn}).
+Moreover, the data producer attaches the time of publishing to the ADS-B message line before publishing it to the Kafka topic.
 
@@
     \includegraphics[width=.9\textwidth, height=.2\textheight]{kafka.png}
-     \caption{Data Stream Producer \& Kafka Cluster Setup.}
+     \caption{The setuop of the Data Stream Producer \& Kafka Cluster.}
 \end{figure}
@@
 \subsection{Evaluation Streaming Workloads}
-In our experimental evaluation, we developed two real-time stream processing workloads that read the ADS-B messages stream from Kafka in order to perform basic trajectories analysis methods.
-The workloads are designed in such way that cover general key aspects of real-time streaming processing tasks, and evaluate the corresponding solutions in Spark and Flink.
+In our experimental evaluation, we developed two real-time stream processing workloads that read a ADS-B messages stream from Kafka in order to perform basic trajectories analysis methods.
+The workloads are designed to cover general key aspects of real-time stream processing tasks and evaluate the corresponding solutions in Spark and Flink.
 The following are the aspects of streaming data processing were covered by our workloads's design:
@@
 \item Manage the order of stream records.
-\item How to provide and update global data model in stream processing task.
+\item How to provide and update global data model in a stream processing task.
 \item Evaluate the performance by measuring the latency and throughput.
 \end{itemize}
-We present the description of workloads and the relation with these aspects, also the implementation details of Spark and Flink solutions in Section 3.2.1 \& 3.2.1.
-Afterward, In Section 4 we discuss the performance evaluation and analysis of these workloads in both frameworks.
+We give the description of workloads, the relation with these aspects and  the implementation details of Spark's and Flink's solutions in Section 3.2.1 \& 3.2.1.
+Afterward, In Section 4 we discuss the performance evaluation and analysis of stream processing workloads in the two frameworks.
 

### 2017-03-08 paper/paper.tex
[paper] updated results section

@@
 The experiment is done using multiple streaming workloads over real-world datasets.
+Results show that Flink outperform Spark streaming in term of latency.
 \section{Introduction}
@@
 
-Figure 5 shows the results of latency measurements of the streaming workloads in Spark and Flink.
+Figure 5 shows the results of the latency measurements of the streaming workloads in Spark and Flink.
 The observed values present the mean value of multiple trails and measurements.
-Measurements for different batch intervals are given in Spark Streaming,  if we decrease  will not get better latency but it is even getting  worse.
-Results show that Flink outperform Spark streaming in term of processing latency in the both evaluation workloads.
+Measurements for different batch intervals are given in Spark Streaming, we noticed that making the batch time too small does not give best latency results.
+Results show that Flink outperform Spark streaming in term of processing latency in the two evaluation workloads.
 
@@
   \centering
-  \includegraphics[width=\linewidth]{latency1.png}
+  \includegraphics[width=\linewidth]{latency2.png}
 \caption{Latency of Sector Change Detection }
@@
 
+The  measurements of throughput for the two streaming workloads are give in Figure 6. The observed values reflect the mean of throughput (message/minute) 
+ in Spark streaming and Flink for multiple executions. Also for Spark streaming we tested different 
+ batch times. Results show that
+
+\begin{figure}[h]
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{latency1.png}
+  \caption{Throughput of Statistics Computation}
+
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{latency1.png}
+\caption{Throughput of Sector Change Detection }
+
+\end{subfigure}
+\caption{Throughput Measurements}
+\label{fig:fig}
+\end{figure}
+
+\newpage
+
 \section{Conclusion}
 
-In this report, we present an experimental evaluation of the streaming processing performance and capabilities of Apache Spark and Apache Flink.
-We carried our experiment over aircraft potions provided in the context of EU Datacorn  project, we developed two streaming workloads to cover general aspects of real-time stream processing tasks, and we measured the latency and throughput of those workloads within the solution of Spark and Flink.
+In this report, we present an experimental evaluation of the stream processing performance and capabilities of Apache Spark and Apache Flink.
+We carried our experiment over aircraft positions datasets provided by the Datacorn project, we developed two streaming workloads to cover general aspects of real-time stream processing tasks.
+Additionally, and we measured the latency and throughput of those workloads solution of Spark and Flink.
 Results show that Flink outperforms Spark Streaming in term of processing latency.
 Furthermore, Flink's processing model is well-suited to the stream processing tasks, there is no need to manage the Stream's state and batch interval as it is in Spark streaming.
-Additionally, Flink operates over the individual record of the data stream, not like Spark Streaming that processes the input data stream as micro-batches.
+Moreover, Flink operates over the individual record of the data stream, while Spark Streaming  processes the input data stream as micro-batches.
 In Summary, Flink is true stream processing framework.
+On the other hand, we noticed that finding online resources related to Spark is easier than to Flink.
 
 In future work, we plan to carry our experiments over cluster of computers not just only on single computer.
-In addition, we will cover more aspects of streaming processing and developing more evaluation workloads.
+In addition, we will cover more aspects of real-time data stream processing and developing more evaluation workloads.
 

### 2017-03-08 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesSectorChangeDetector.java
finish latency measurements

@@ -9,6 +9,7 @@ import org.apache.flink.streaming.api.datastream.KeyedStream;
 import org.apache.flink.streaming.api.environment.StreamExecutionEnvironment;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.LoggerUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
 import de.kdml.bigdatalab.spark_and_flink.flink.utils.FlinkUtils;
@@ -33,24 +34,31 @@ public class TrajectoriesSectorChangeDetector {
 
 		// assign sector for each trajectory, then check if there is change in
 		// the sector between two consecutive trajectories
-		DataStream<String> trajectoriesStreamWithSectors = trajectoriesStream
-				.reduce(new TrajectoriesSectorsReducer(sectorsDataSet.collect())).filter(tuple -> {
-					PositionMessage position = tuple.f1;
-					// check for change in previous and current sector
-					if (position.getPrevSector() != null && !position.getSector().equals(position.getPrevSector())) {
-						return true;
-					}
-					return false;
-				}).map(tuple -> {
 
-					// print the sector transition
-					PositionMessage trajectory = tuple.f1;
-					return tuple.f0 + ": " + trajectory.getPrevSector().getNameAndAirBlock() + " --> "
-							+ trajectory.getSector().getNameAndAirBlock();
+		DataStream<Tuple2<String, PositionMessage>> runningTrajectories = trajectoriesStream
+				.reduce(new TrajectoriesSectorsReducer(sectorsDataSet.collect()));
 
-				});
+		DataStream<String> trajectoriesStreamWithSectors = runningTrajectories.filter(tuple -> {
+			PositionMessage position = tuple.f1;
+			// check for change in previous and current sector
+			if (position.getPrevSector() != null && !position.getSector().equals(position.getPrevSector())) {
+				return true;
+			}
+			return false;
+		}).map(tuple -> {
+
+			// print the sector transition
+			PositionMessage trajectory = tuple.f1;
+			String outputLine = tuple.f0 + ": " + trajectory.getPrevSector().getNameAndAirBlock() + " --> "
+					+ trajectory.getSector().getNameAndAirBlock();
+
[…]

### 2017-03-08 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesStatistics.java

@@ -1,13 +1,8 @@
 package de.kdml.bigdatalab.spark_and_flink.flink.datacorn;
 
-import org.apache.flink.api.common.functions.FoldFunction;
 import org.apache.flink.api.java.tuple.Tuple2;
 import org.apache.flink.streaming.api.datastream.DataStream;
 import org.apache.flink.streaming.api.environment.StreamExecutionEnvironment;
-import org.apache.flink.streaming.api.functions.AssignerWithPeriodicWatermarks;
-import org.apache.flink.streaming.api.functions.sink.SinkFunction;
-import org.apache.flink.streaming.api.watermark.Watermark;
-import org.apache.flink.streaming.api.windowing.time.Time;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.LoggerUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.StatisticsUtils;
@@ -27,7 +22,7 @@ public class TrajectoriesStatistics {
 	public static void main(String[] args) throws Exception {
 
 		StreamExecutionEnvironment env = FlinkUtils.getInitializedEnv();
-
+		LoggerUtils.logMessage("Start Time:" + System.currentTimeMillis());
 		/**
 		 * Reduce the trajectories by aggregate the statics from the previous
 		 * trajectory record and keep the new record, the further aggregation
@@ -44,89 +39,25 @@ public class TrajectoriesStatistics {
 							? tuple2 : tuple1;
 					Tuple2<String, PositionMessage> newTuple = tuple2.f1.getStreamedTime() < tuple1.f1.getStreamedTime()
 							? tuple1 : tuple2;
+
+					// to count first tuple in the throughput
+					// if (oldTuple.f1.getSpeed() == null) {
+
+					// LoggerUtils.logMessage("-" + System.currentTimeMillis());
+					// }
+
 					StatisticsUtils.computeStatistics(oldTuple.f1, newTuple.f1);
 					newTuple.f1.setNew(false);
+
+					// LoggerUtils.logMessage("-" + System.currentTimeMillis());
 					return newTuple;
 				});
 
-		showLatecies(trajectoriesStream);
+		// TrajectoriesStreamUtils.showLatecies(trajectoriesStream);
 		trajectoriesStream.print().setParallelism(1);
 
 		// showThroughput(trajectoriesStream);
 		env.execute(" Flink Trajectories Statistics Computation");
 	}
 
-	private static void showThroughput(DataStream<Tuple2<String, PositionMessage>> trajectoriesStream) {
[…]

### 2017-03-08 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesStreamUtils.java

@@ -2,15 +2,21 @@ package de.kdml.bigdatalab.spark_and_flink.flink.datacorn;
 
 import java.util.Properties;
 
+import org.apache.flink.api.common.functions.FoldFunction;
 import org.apache.flink.api.java.tuple.Tuple;
 import org.apache.flink.api.java.tuple.Tuple2;
 import org.apache.flink.streaming.api.datastream.DataStream;
 import org.apache.flink.streaming.api.datastream.KeyedStream;
 import org.apache.flink.streaming.api.environment.StreamExecutionEnvironment;
+import org.apache.flink.streaming.api.functions.AssignerWithPeriodicWatermarks;
+import org.apache.flink.streaming.api.functions.sink.SinkFunction;
+import org.apache.flink.streaming.api.watermark.Watermark;
+import org.apache.flink.streaming.api.windowing.time.Time;
 import org.apache.flink.streaming.connectors.kafka.FlinkKafkaConsumer09;
 import org.apache.flink.streaming.util.serialization.SimpleStringSchema;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.LoggerUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.TrajectoriesUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.StreamRecord;
@@ -39,7 +45,7 @@ public class TrajectoriesStreamUtils {
 		props.setProperty("zookeeper.connect", configs.getStringProp("zookeeper"));
 		props.setProperty("bootstrap.servers", configs.getStringProp("bootstrap.servers"));
 		props.setProperty("group.id", configs.getStringProp("kafkaGroupId"));
-		// Always read topicfrom start
+		// Always read topicfrom start 
 		props.setProperty("auto.offset.reset", "earliest");
 
 		// create a Kafka consumer
@@ -77,4 +83,79 @@ public class TrajectoriesStreamUtils {
 
 		return trajectoriesStream;
 	}
+
+	/**
+	 * Show latencies of processed trajectories based on the time delay between
+	 * streamed time and finished time
+	 * 
+	 * @param trajectoriesStream
+	 */
+
[…]

### 2017-03-08 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesSectorChangeDetector.java

@@ -1,6 +1,7 @@
 package de.kdml.bigdatalab.spark_and_flink.spark_project.datacorn;
 
 import java.util.ArrayList;
+import java.util.Arrays;
 import java.util.List;
 
 import org.apache.spark.api.java.JavaSparkContext;
@@ -38,7 +39,7 @@ public class TrajectoriesSectorChangeDetector {
 		JavaSparkContext sc = SparkConfigsUtils.getSparkContext("Trajectories sector change  deetctor ");
 
 		JavaStreamingContext jssc = new JavaStreamingContext(sc,
-				Durations.seconds(configs.getIntProp("batchDuration")));
+				Durations.milliseconds(configs.getIntProp("batchDuration")));
 
 		// broadcast the sector list
 		sectors = sc.broadcast(getSectorsDataSet(sc));
@@ -50,6 +51,7 @@ public class TrajectoriesSectorChangeDetector {
 		JavaPairDStream<String, Iterable<PositionMessage>> runningTrajectories = trajectories
 				.updateStateByKey(updateTrajectoriesAndAssignSectors);
 
+		TrajectoriesStreamUtils.printLatencies(runningTrajectories);
 		// find trajectories with change in sector
 		JavaDStream<String> changedTrajectories = runningTrajectories.filter(trajectoryTuple -> {
 
@@ -94,6 +96,8 @@ public class TrajectoriesSectorChangeDetector {
 				}
 				prevPosition = currentPosition;
 			}
+
+			// LoggerUtils.logMessage(output.toString());
 			return output.toString();
 		});
 
@@ -133,9 +137,8 @@ public class TrajectoriesSectorChangeDetector {
 	}
 
 	/**
-	 * Return a new "state" DStream where the state for each key is updated by
-	 * applying the given function on the previous state of the key and the new
-	 * values for the key
+	 * Update the positions of trajectory every new batch, by assigning the
+	 * sectors for the received positions
 	 * 
 	 */
 	public static Function2<List<Iterable<PositionMessage>>, Optional<Iterable<PositionMessage>>, Optional<Iterable<PositionMessage>>> updateTrajectoriesAndAssignSectors = new Function2<List<Iterable<PositionMessage>>, Optional<Iterable<PositionMessage>>, Optional<Iterable<PositionMessage>>>() {
@@ -143,61 +146,62 @@ public class TrajectoriesSectorChangeDetector {
 		private static final long serialVersionUID = -76088662409004569L;
 
 		@Override
[…]

### 2017-03-08 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesStatistics.java

@@ -8,7 +8,6 @@ import org.apache.spark.api.java.JavaSparkContext;
 import org.apache.spark.api.java.Optional;
 import org.apache.spark.api.java.function.Function2;
 import org.apache.spark.streaming.Durations;
-import org.apache.spark.streaming.api.java.JavaDStream;
 import org.apache.spark.streaming.api.java.JavaPairDStream;
 import org.apache.spark.streaming.api.java.JavaStreamingContext;
 
@@ -42,7 +41,7 @@ public class TrajectoriesStatistics {
 		long batchTime = configs.getIntProp("batchDuration");
 		// configure the stream batch interval
 		JavaStreamingContext jssc = new JavaStreamingContext(sc, Durations.milliseconds(batchTime));
-
+		LoggerUtils.logMessage("Start Time:" + System.currentTimeMillis());
 		// create the trajectories stream
 		JavaPairDStream<String, Iterable<PositionMessage>> trajectories = TrajectoriesStreamUtils
 				.getTrajectoriesStream(jssc);
@@ -54,7 +53,7 @@ public class TrajectoriesStatistics {
 		JavaPairDStream<String, Iterable<PositionMessage>> runningTrajectories = trajectories
 				.updateStateByKey(updateTrajectoriesAndComputeStatistics);
 
-		printLatencies(runningTrajectories);
+		// printLatencies(runningTrajectories);
 
 		runningTrajectories.print(1000);
 
@@ -85,37 +84,6 @@ public class TrajectoriesStatistics {
 			LoggerUtils.logMessage(time + " count=" + count + "Current throughput = " + (count / batchTimeInSeconds)
 					+ " records / second");
 		});
-	}
-
-	/**
-	 * Print latencies measurements
-	 * 
-	 * @param runningTrajectories
-	 */
-	private static void printLatencies(JavaPairDStream<String, Iterable<PositionMessage>> runningTrajectories) {
-		JavaDStream<Long> latencies = runningTrajectories.map(trajectoryTuple -> {
-
-			// find the average latency for all positions in a trajectory
-			long currentTime = System.currentTimeMillis(), sumLatency = 0, count = 0;
-
-			Iterable<PositionMessage> positionsInTrajectory = trajectoryTuple._2;
-			for (PositionMessage position : positionsInTrajectory) {
-				if (position.isNew()) {
-
-					long latency = position.getFinishProcessingTime() - position.getStreamedTime();
[…]

### 2017-03-08 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesStreamUtils.java

@@ -17,6 +17,7 @@ import org.apache.spark.streaming.kafka010.KafkaUtils;
 import org.apache.spark.streaming.kafka010.LocationStrategies;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.LoggerUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.TrajectoriesUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.StreamRecord;
@@ -127,5 +128,36 @@ public class TrajectoriesStreamUtils {
 		// Sort the new positions based on create time to preserve the
 		// correct order of message order
 		return TrajectoriesUtils.sortPositionsOfTrajectory(aggregatedPositions);
+	}
+
+	/**
+	 * Print latencies measurements
+	 * 
+	 * @param runningTrajectories
+	 */
+	public static void printLatencies(JavaPairDStream<String, Iterable<PositionMessage>> runningTrajectories) {
+		JavaDStream<Long> latencies = runningTrajectories.map(trajectoryTuple -> {
+
+			// find the average latency for all positions in a trajectory
+			long currentTime = System.currentTimeMillis(), sumLatency = 0, count = 0;
+
+			Iterable<PositionMessage> positionsInTrajectory = trajectoryTuple._2;
+			for (PositionMessage position : positionsInTrajectory) {
+				if (position.isNew()) {
+
+					long latency = position.getFinishProcessingTime() - position.getStreamedTime();
+					LoggerUtils.logMessage(String.valueOf(latency));
+					sumLatency += latency;
+					count++;
+				}
+			}
+			return count == 0 ? 0 : new Long(sumLatency / count);
+
+		}).filter(latency -> {
+			return latency > 0;
+		});
+
+		// show latencies
+		latencies.print();
 	}
 }

### 2017-03-08 paper/paper.tex
[paper] refinements

@@
 Results show that Flink outperform Spark streaming in term of latency.
+While Spark Streaming gives higher throughput rates than Flink.
 \section{Introduction}
@@
 The message line comprises 22 fields of data such as aircraft ID, date message generated, longitude, latitude and altitude.
+And  these datasets contain around \texttt{26 millions} messages.
 \par In our experiment a data stream producer component reads the ADS-B message datasets, then publishes them to the Kafka cluster to be consumed by the workloads in Apache Spark and Apache Flink.
@@
 Afterward, In Section 4 we discuss the performance evaluation and analysis of stream processing workloads in the two frameworks.
+The source code of the two streaming workloads and the data stream producers is available in a repository on Github\footnote{\url{https://github.com/ehabqadah/Spark_vs_Flink}}.
 
@@
 \par This section presents the results of performance evaluation in Spark (2.0.2) and Flink (1.1.3) for the streaming workloads were described in Section 3.2.
-Specifically, we provide  the measurements of latency and throughput of the benchmarking workloads execution in Spark and Flink.
+Specifically, we provide  the measurements of latency and throughput of the benchmarking workloads execution in Spark Streaming and Flink.
 We carried the experiment by setup the Kafka stream as described in Section 3.1 on local machine\footnote{ The experiments have been carried  on single machine with Ubuntu 16.04 LTS os, Intel® Core™ i5-6200U CPU @ 2.30GHz × 4  processor, and 16 GB of memory.
-}, and run the different workloads on the same machine (i.e., local mode) to obtain the performance results.
+}, and execute the different workloads on the same machine (i.e., local mode) to obtain the performance results.
 
@@
 The observed values present the mean value of multiple trails and measurements.
-Measurements for different batch intervals are given in Spark Streaming, we noticed that making the batch time too small does not give best latency results.
+Measurements for different batch durations are given in Spark Streaming, we noticed that making the batch time too small does not give best latency results.
 Results show that Flink outperform Spark streaming in term of processing latency in the two evaluation workloads.
@@
  in Spark streaming and Flink for multiple executions. Also for Spark streaming we tested different 
- batch times. Results show that
+ batch times. Results show that Spark Streaming has higher throughput than Flink, with increasing the batch duration.
 
 \begin{figure}[h]
+\centering
 \begin{subfigure}{.5\textwidth}
   \centering
-  \includegraphics[width=\linewidth]{latency1.png}
-  \caption{Throughput of Statistics Computation}
+  \includegraphics[width=\linewidth]{throughput1.png}
+  \caption{Statistics Computation}
 
@@
   \centering
-  \includegraphics[width=\linewidth]{latency1.png}
-\caption{Throughput of Sector Change Detection }
+  \includegraphics[width=\linewidth]{throughput1.png}
+\caption{Sector Change Detection }
 
@@
 \end{figure}
-
-\newpage
 
@@
 Results show that Flink outperforms Spark Streaming in term of processing latency.
-Furthermore, Flink's processing model is well-suited to the stream processing tasks, there is no need to manage the Stream's state and batch interval as it is in Spark streaming.
+In contrast, Spark Streaming has showed better throughput rates than Flink.
+\par To sum up, Flink's processing model is well-suited to the stream processing tasks that require low latency, also there is no need to manage the Stream's state and batch interval as it is in Spark streaming.
 Moreover, Flink operates over the individual record of the data stream, while Spark Streaming  processes the input data stream as micro-batches.
-In Summary, Flink is true stream processing framework.
-On the other hand, we noticed that finding online resources related to Spark is easier than to Flink.
+Thus we can state that  Flink is a true stream processing framework.
+A side remark, we have  noticed that finding online resources related to Spark is easier than to Flink.
 
@@
 
+\newpage
 \begin{thebibliography}{[MT1]}

### 2017-03-08 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesStatistics.java
throughput measurements

@@ -49,7 +49,7 @@ public class TrajectoriesStatistics {
 					StatisticsUtils.computeStatistics(oldTuple.f1, newTuple.f1);
 					newTuple.f1.setNew(false);
 
-					// LoggerUtils.logMessage("-" + System.currentTimeMillis());
+					LoggerUtils.logMessage("-" + System.currentTimeMillis());
 					return newTuple;
 				});
 

### 2017-03-08 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesStatistics.java

@@ -129,7 +129,7 @@ public class TrajectoriesStatistics {
 				StatisticsUtils.computeStatistics(lastOldPosition, position);
 				long currentTime = System.currentTimeMillis();
 				position.setFinishProcessingTime(currentTime);
-				LoggerUtils.logMessage("_" + currentTime);
+				LoggerUtils.logMessage(currentTime + "-");
 				lastOldPosition = position;
 
 				/**

### 2017-03-09 paper/paper.tex
[paper] updated sections 1&2

@@
 \centering
- Lab Report: Data Science and Big Data MA-INF 4306,
- Winter Term 2016/2017, University of Bonn
+ MA-INF 4306 - Data Science and Big Data \\
+ Lab Report\\
+
+   Winter Semester 2016.17\\
+
+		 University of Bonn\\
 
@@
 \par In recent years, Big Data is considered as one of the most appealing topics in Information Technology, it has gained increasing attention of professionals in industry and academia.
-As result of the fact that the amount of data we generate is increasing exponentially \cite{idc}.
-Different sources such as  social media platforms, mobile phones, IoT and finical transactions, etc.
+This attention is a result of the fact that the amount of data we generate is increasing exponentially \cite{idc}.
+Different sources such as  social media platforms, mobile phones, Internet of things (IoT), financial transactions, etc.
 create large volumes of data.
-Moreover, recent years have shown a rapid development of Big Data frameworks such as Apache Spark and Apache Flink that provide efficient data processing functionalities.
- \par Big Data is a term that does not only describe large datasets (volume), but also the data that arrives in high rate (velocity), which comes in all kinds of data format (variety) (i.e., 3Vs of Big Data (Volume, Velocity, Variety)) \cite{svs}.
+Moreover, many Big Data frameworks have been recently developed such as Apache Spark and Apache Flink that provide efficient data processing functionalities.
+ \par Big Data is a term that does not only describe large datasets (volume), but also the data that arrives in high rate (velocity), and it comes in a mix of different data format (variety) (i.e., 3Vs of Big Data (Volume, Velocity, Variety)) \cite{svs}.
 
@@
 While Stream processing  is a method to manage fast and continuously incoming data (real-time), as it arrives to the system (e.g., credit card fraud detection and network monitoring applications).
-Where low latency is crucial for stream processing systems.
+Stream processing systems usually  provide low latency processing.
 Both processing models are capable of operating with structured or semi-structured data (i.e., variety).
 
-\par There exist some performance evaluations of Apache Spark and Apache Flink  such as Yahoo Streaming Benchmark \cite{yahoo}.
-We believe that we still need to carry our evaluation for two reasons: \begin{enumerate*}[label=(\roman*)]
+\par There exist some publications that compare Apache Spark and Apache Flink
+ such as the Yahoo Streaming Benchmark \cite{yahoo}. But we believe that we still need to carry our evaluation for two reasons: \begin{enumerate*}[label=(\roman*)]
 \item both frameworks are continuously improving in short time 
@@
 In order to simulate real-world use cases, we use datasets of airplane trajectories provided by the  Datacron project\footnote{\url{http://www.datacron-project.eu/}} to develop multiple streaming data processing workloads.
-Beside the performance comparison of stream processing of the two frameworks, we try to cover general aspects of stream processing tasks (see Section 3) and how they can be solved in each platform.
+Beside the performance comparison of the stream processing of the two frameworks, we try to cover general aspects of stream processing tasks (see Section 3) and how they can be solved in each platform.
 Informally,  this work aims to help the developers to determine which platform to use for different use cases.
@@
 In Section 2, we present the main characteristics, programming APIs, and main data abstraction of Apache Spark and Apache Flink.
-Section 3  presents the general aspects of streaming processing and the implementation of experimental workloads.
+Section 3  presents the general aspects of stream processing tasks and the implementation of experimental workloads.
 Section 4
-compares the performance results of the different workloads. Section 5 gives the overall conclusion.
+compares the performance results of the different workloads. And finally, Section 5 gives the overall conclusion.
 
@@
 Additionally, Discretized Stream (DStream) is the data abstraction in Spark Streaming, which is a series of RDDs that represent an input stream of data \cite{spark_streaming}.
-In other words, the Spark Streaming is a batch-based processing model, which processes the continuous stream of data by divided it into micro-batches that are  processed by the Spark engine.
+In other words, the Spark Streaming is a batch-based processing model, which processes the continuous stream of data by dividing it into micro-batches that are  processed by the Spark engine.
 Figure 1 illustrates the general process flow of Spark Streaming.
@@
 
-\par The Flink's core is a distributed streaming dataflow engine,  with each Flink program is represented by a data flow graph (i.e., directed acyclic graph - DAG) that executed by the Flink's engine \cite{flink_paper}.
+\par The Flink's core is a distributed streaming dataflow engine,  with each Flink program is represented by a dataflow graph (i.e., directed acyclic graph - DAG) that executed by the Flink's engine \cite{flink_paper}.
 The data flow graphs are composed of stateful, parallel operations and intermediate data stream partitions.
@@
 \par Apache Kafka is scalable, fault-tolerant and distributed streaming framework \cite{kafka}. It allows to publish and subscribe to data streams.
-Kafka manages the stream records in different categories (i.e., topics) that partitioned and distributed over the servers of the Kafka cluster.
+Kafka manages the stream records in different categories (i.e., topics) that are partitioned and distributed over the servers of the Kafka cluster.
 It allows the data producers to publish a stream of records to one or more  Kafka topic.
 Additionally, consumer applications can subscribe to one or more topic to read data streams.
-The consumers are grouped into groups based on the group name that allows Kafka to distribute and balance the stream partitions among the  members of a certain group for the sake of scalability.
+The consumers within a group use the  same group name, which allows Kafka to distribute and balance the stream partitions among the  members of a certain group for the sake of scalability.
 Figure 3 shows how 4 partitions of a stream are distributed between two groups of consumers based on the number of members.
@@
 \section{Experiment Setup and Implementation}
-This section describes the data stream setup, the design of evaluation streaming workloads and the solution's implementation details of  Apache Spark and Apache Flink.
+This section describes the data stream setup, the design of evaluation streaming workloads and the  implementation details of the proposed solution of Apache Spark and Apache Flink.
 
@@
 The evaluation streaming workloads read an input data stream from Kafka.
-In order to simulate real-world use case we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that present the aircraft's position over time.
+In order to simulate real-world use case we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that represent the aircraft's position over time.
 The message line comprises 22 fields of data such as aircraft ID, date message generated, longitude, latitude and altitude.

### 2017-03-09 paper/paper.tex
[paper] updated throughput results

@@
 
-\vspace{4pt}
+\vspace{6pt}
 \centering
@@
 In this report, we provide a performance comparison of streaming data processing in Apache Spark and Apache Flink by measuring different performance metrics (latency and throughput).
-In addition, we cover some key aspects of streaming data applications and how they are handled in the two frameworks.
+Furthermore, we cover some key aspects of streaming data applications and how they are handled in the two frameworks.
 The experiment is done using multiple streaming workloads over real-world datasets.
@@
   \centering
-    \includegraphics[width=.9\textwidth, height=.2\textheight]{streaming-flow.png}
+    \includegraphics[width=.9\textwidth, height=.3\textheight]{streaming-flow.png}
      \caption{ Process flow of Spark Streaming \cite{spark_streaming}.}
@@
   \centering
-    \includegraphics[width=\textwidth, height=.3\textheight]{flink_engine.png}
+    \includegraphics[width=\textwidth, height=.5\textheight]{flink_engine.png}
      \caption{ An example of data flow graph in Flink \cite{flink}.}
 \end{figure} 
+\newpage
 \subsection{Apache Kafka}
@@
   \centering
-    \includegraphics[width=\textwidth, height=.3\textheight]{kafka_groups.png}
+    \includegraphics[width=\textwidth, height=.4\textheight]{kafka_groups.png}
      \caption{ Distribution of a stream partitions for consumer groups in Kafka \cite{kafka}.}
@@
 
+\newpage
 \section{Experiment Setup and Implementation}
@@
   \centering
-    \includegraphics[width=.9\textwidth, height=.2\textheight]{kafka.png}
+    \includegraphics[width=.9\textwidth, height=.3\textheight]{kafka.png}
      \caption{The setuop of the Data Stream Producer \& Kafka Cluster.}
@@
 
+\newpage
 \subsection{Evaluation Streaming Workloads}
@@
 \end{itemize}
-We give the description of workloads, the relation with these aspects and  the implementation details of Spark's and Flink's solutions in Section 3.2.1 \& 3.2.1.
+We give the description of workloads, the relation with these aspects and  the implementation details of Spark's and Flink's solutions in Sections 3.2.1 \& 3.2.1.
 Afterward, In Section 4 we discuss the performance evaluation and analysis of stream processing workloads in the two frameworks.
-The source code of the two streaming workloads and the data stream producers is available in a repository on Github\footnote{\url{https://github.com/ehabqadah/Spark_vs_Flink}}.
+\par The two streaming workloads and the data stream producers are implemented in Java and the source code is available in a repository on Github\footnote{\url{https://github.com/ehabqadah/Spark_vs_Flink}}.
 
@@
 Since the processing  model of Spark Streaming is micro-batch based and stateless, \texttt{updateStateByKey} operation must be used to maintain the state of the trajectories stream.
-In the context of this workload, the state is the aggregated statistics of trajectory, so the statistics computation is preformed within the custom \texttt{updateStateByKey} function that continuously applied for every batch.
+In the context of this workload, the state is the aggregated statistics of trajectory, so the statistics computation is preformed within the custom \texttt{updateStateByKey} function that is continuously applied for every batch.
  The state update function uses the last position of trajectory from previous batch to aggregate and compute statistics quantities to each trajectory's position in the new data batch, and the new positions list is kept as the new state for the trajectory.
@@
 
-The goal of sector change detection workload is to detect the  enter or leaving of aircraft from one air sector to other, by processing the real-time stream of aircraft positions.
-Given that the dataset of sectors (i.e., polygons) is available as reference to assign the corresponding sector for certain aircraft's position.
-In this workload, we test the aspect of providing the  global model of data (sectors dataset) in streaming processing workload.
-Moreover, the preserving of the stream's records order and consuming of parallel Kafka streams points are also covered.
+The goal of sector change detection workload is to detect the  entering or leaving of an aircraft from one air sector to another one, by processing the real-time stream of aircraft positions.
+Given the fact that a dataset of sectors (i.e., polygons) is available as a reference to assign the corresponding sector for certain aircraft's position.
+In this workload, we test the aspect of providing a global data model (sectors dataset) in stream processing workload.
+Moreover, the preserving of the stream's records order and consuming of parallel Kafka streams  are covered.
 
@@
 
- We implemented the sector change detection workload in Spark Streaming by constructing the trajectories stream and assigning the sectors for each position of a trajectory, then a change between the assigned sectors of two consecutive aircraft's positions is marked  as  an transition of sector for certain trajectory.
-First, a multiple instances of \texttt{DStream} are created using \texttt{KafkaUtils.createDirectStream} to read in  parallel from Kafka, after  all input data streams are combined using \texttt{union} function.
-Second, the records (i.e., ADS-B messages) of the resulted union stream are parsed and mapped to tuples of (aircraft ID, position message) using defined \texttt{mapToPair} transformation.
-Third, the unrelated  tuples are filtered based on the message type through a \texttt{filter} operation.
-Fourth, the trajectories stream is constructed by combining all tuples with the same ID using the \texttt{groupByKey} function.
+ We implemented the sector change detection workload in Spark Streaming by constructing the stream of trajectories and assigning a sector for each position of a trajectory.
+A change between the assigned sectors of two consecutive aircraft's positions is marked  as a transition of sector for a certain trajectory.
+\par First, multiple instances of \texttt{DStream} are created using \texttt{KafkaUtils.create} \texttt{DirectStream}  are combined using \texttt{union} to read in  parallel from Kafka.
+Second, the records (i.e., ADS-B messages) of the resulting stream are parsed and mapped to tuples of (aircraft ID, position message) using \texttt{mapToPair} transformation.
+Third, the unrelated  tuples are filtered based on the message type using a \texttt{filter} operation.
+Fourth, the stream of trajectories is constructed by combining all tuples with the same ID using the \texttt{groupByKey} function.
 Then, to manage and aggregate the state of trajectories stream between the mini-batches a configured \texttt{updateStateByKey} function is applied.
-The state update function assign the corresponding sector for each new position, the old positions of a certain trajectory from previous batch are discarded, but only the last position of old batch is kept to be used alongside with the new positions.
-In addition, the new positions are sorted to maintain the correct order.
- The sector data set is provided using the built-in \texttt{Broadcast} feature that distributes a shared value for  all cluster nodes in efficient matter.
-Finally, the detection of sector change is done by an extra filtering transformation using a \texttt{filter} operator, by checking the sectors of consecutive positions in a trajectory to find the ones with a change in the sector assignment.
+The state update function assigns the corresponding sector for each new position, the old positions of a certain trajectory from the previous batch are discarded, but only the last position of the old batch is kept to be used alongside with the new positions.
+In addition to that the new positions are sorted by th streaming time to maintain the correct order.
+ \par The sector data set is provided using the built-in \texttt{Broadcast} feature that distributes a shared value for  all cluster nodes in an efficient matter.
+Finally, the detection of a sector change is done by an extra filtering transformation using a \texttt{filter} operator, which checks the sectors of consecutive positions in a trajectory to find the ones with a change in the sector assignment.
 
@@
 
- The Flink's implementation of sector change detection workload, uses the \texttt{FlinkKafkaConsumer09} that reads the position messages (i.e., ADS-B) stream from Kafka, which used to construct \texttt{DataStream} instance.
-The \texttt{DataStream}'s records are mapped to tuples of (aircraft ID and aircraft's position), by a configured \texttt{map} transformation.
-The irrelevant tuples are filtered based on the message type using a \texttt{filter} operation.
+ The Flink's implementation of sector change detection workload uses the \texttt{FlinkKafkaConsumer09} that reads the position messages (i.e., ADS-B) stream from Kafka, which is used to construct a \texttt{DataStream} instance.
+The \texttt{DataStream}'s records are mapped to tuples of (aircraft ID and aircraft's position) by a configured \texttt{map} transformation.
+Irrelevant tuples are filtered based on the message type using a \texttt{filter} operation.
 Afterward, the tuples that shares a common IDs are grouped together by the \texttt{keyBy} operator.
@@
 
-\par This section presents the results of performance evaluation in Spark (2.0.2) and Flink (1.1.3) for the streaming workloads were described in Section 3.2.
+\par This section presents the results of performance evaluation in Spark (2.0.2) and Flink (1.1.3) for the streaming workloads that were described in Section 3.2.
 Specifically, we provide  the measurements of latency and throughput of the benchmarking workloads execution in Spark Streaming and Flink.
-We carried the experiment by setup the Kafka stream as described in Section 3.1 on local machine\footnote{ The experiments have been carried  on single machine with Ubuntu 16.04 LTS os, Intel® Core™ i5-6200U CPU @ 2.30GHz × 4  processor, and 16 GB of memory.
-}, and execute the different workloads on the same machine (i.e., local mode) to obtain the performance results.
+We carried the experiment by setting up the Kafka stream as described in Section 3.1 on a local machine\footnote{ The experiments have been carried  on single machine with Ubuntu 16.04 LTS os, Intel® Core™ i5-6200U CPU @ 2.30GHz × 4  processor, and 16 GB of memory.} and execute the different workloads on the same machine (i.e., local mode) to obtain the performance results.
 
-\par We consider the time delay between the streaming time of record in Kafka and the 
-finish processing time in a workload as the platform's processing latency (measured in milliseconds).
-On the other hand, the number of processed stream's records (aircrafts position messages) per second within each streaming workload is the throughput measure.
-Since the Spark Streaming is micro-batch based,  We examine the effect of different batch intervals
+\par We consider the time delay between the streaming time of a record in Kafka and  the time when the record is processed by a workload as the platform's processing latency (measured in milliseconds).
+And the number of processed stream's records (aircrafts position messages) per second within each streaming workload is the throughput measure.
+Since the Spark Streaming is micro-batch based,  we examine the effect of different batch intervals
 of the Spark Streaming on the performance measurements.
+
 \subsection{Latency Results}
@@
 Figure 5 shows the results of the latency measurements of the streaming workloads in Spark and Flink.
-The observed values present the mean value of multiple trails and measurements.
-Measurements for different batch durations are given in Spark Streaming, we noticed that making the batch time too small does not give best latency results.
+The presented values present the average value of multiple trails and measurements.
+Measurements for different batch durations are given in Spark Streaming, we noticed that making the batch time too small does not give better latency results.
 Results show that Flink outperform Spark streaming in term of processing latency in the two evaluation workloads.
@@
 \begin{figure}[h]
-\begin{subfigure}{.5\textwidth}
+\centering
+\begin{subfigure}{.7\textwidth}
   \centering
@@
 \end{subfigure}%
-\begin{subfigure}{.5\textwidth}
+\vskip\baselineskip
+\centering
+\begin{subfigure}{.7\textwidth}
   \centering
@@
 \end{figure}
+\newpage
 \subsection{Throughput Results}
 
-The  measurements of throughput for the two streaming workloads are give in Figure 6. The observed values reflect the mean of throughput (message/minute) 
+The  measurements of throughput for the two streaming workloads are given in Figure 6. The presented values reflect the average of throughput (message/minute) 
  in Spark streaming and Flink for multiple executions. Also for Spark streaming we tested different 
- batch times. Results show that Spark Streaming has higher throughput than Flink, with increasing the batch duration.
+ batch times. Results show that Spark Streaming has higher throughput than Flink, also  increasing the batch duration of Spark Streaming gives better throughput.
 
@@
 \centering
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.7\textwidth}
   \centering
@@
 \end{subfigure}%
-\begin{subfigure}{.5\textwidth}
+
+\vskip\baselineskip
 \centering
-  \includegraphics[width=\linewidth]{throughput1.png}
+\begin{subfigure}{.7\textwidth,}
+  \centering
+  \includegraphics[width=\linewidth]{throughput2.png}
 \caption{Sector Change Detection }
@@
 In contrast, Spark Streaming has showed better throughput rates than Flink.
-\par To sum up, Flink's processing model is well-suited to the stream processing tasks that require low latency, also there is no need to manage the Stream's state and batch interval as it is in Spark streaming.
+\par To sum up, Flink's processing model is well-suited to the stream processing tasks that require low latency, also there is no need to manage the Stream's state and the batch interval in contrast to Spark streaming.
 Moreover, Flink operates over the individual record of the data stream, while Spark Streaming  processes the input data stream as micro-batches.
@@
 
-\newpage
 \begin{thebibliography}{[MT1]}

### 2017-03-09 paper/paper.tex
[paper] minor updates and added cross referencing of figures and sections

@@
 In order to simulate real-world use cases, we use datasets of airplane trajectories provided by the  Datacron project\footnote{\url{http://www.datacron-project.eu/}} to develop multiple streaming data processing workloads.
-Beside the performance comparison of the stream processing of the two frameworks, we try to cover general aspects of stream processing tasks (see Section 3) and how they can be solved in each platform.
+Beside the performance comparison of the stream processing of the two frameworks, we try to cover general aspects of stream processing tasks (see Section~\ref{sec:sec3}) and how they can be solved in each platform.
 Informally,  this work aims to help the developers to determine which platform to use for different use cases.
@@
 \par The remainder of this report is organized as follows.
-In Section 2, we present the main characteristics, programming APIs, and main data abstraction of Apache Spark and Apache Flink.
-Section 3  presents the general aspects of stream processing tasks and the implementation of experimental workloads.
-Section 4
-compares the performance results of the different workloads. And finally, Section 5 gives the overall conclusion.
+In Section~\ref{sec:sec2}, we present the main characteristics, programming APIs, and main data abstraction of Apache Spark and Apache Flink.
+Section~\ref{sec:sec3} presents the general aspects of stream processing tasks and the implementation of experimental workloads.
+Section~\ref{sec:sec4}
+compares the performance results of the different workloads. And finally, Section~\ref{sec:sec5} gives the overall conclusion.
 
 \section{Technical Background}
+\label{sec:sec2}
  In this section, we present the architecture, key characteristics, programming APIs  of Apache Spark and Apache Flink. Moreover, a brief overview of the Apache Kafka platform is  provided.
@@
 In other words, the Spark Streaming is a batch-based processing model, which processes the continuous stream of data by dividing it into micro-batches that are  processed by the Spark engine.
-Figure 1 illustrates the general process flow of Spark Streaming.
+Figure~\ref{fig:spark} illustrates the general process flow of Spark Streaming.
 
@@
      \caption{ Process flow of Spark Streaming \cite{spark_streaming}.}
+      \label{fig:spark}
 \end{figure}
@@
 The data flow graphs are composed of stateful, parallel operations and intermediate data stream partitions.
-Figure 2 shows a data flow graph of a Flink program.
+Figure~\ref{fig:fstream} shows a data flow graph of a Flink program.
 
@@
      \caption{ An example of data flow graph in Flink \cite{flink}.}
+      \label{fig:fstream}
 \end{figure} 
@@
 The consumers within a group use the  same group name, which allows Kafka to distribute and balance the stream partitions among the  members of a certain group for the sake of scalability.
-Figure 3 shows how 4 partitions of a stream are distributed between two groups of consumers based on the number of members.
+Figure~\ref{fig:kafka} shows how four partitions of a stream are distributed between two groups of consumers based on the number of members.
 Apache Kafka has been widely  adopted, for example, 
@@
      \caption{ Distribution of a stream partitions for consumer groups in Kafka \cite{kafka}.}
+     \label{fig:kafka}
 \end{figure}
@@
 \section{Experiment Setup and Implementation}
+\label{sec:sec3}
 This section describes the data stream setup, the design of evaluation streaming workloads and the  implementation details of the proposed solution of Apache Spark and Apache Flink.
@@
 \par In our experiment a data stream producer component reads the ADS-B message datasets, then publishes them to the Kafka cluster to be consumed by the workloads in Apache Spark and Apache Flink.
-Figure 4 illustrates the data producer and the Kafka cluster setup in all experiments.
+Figure~\ref{fig:data_stream} illustrates the data producer and the Kafka cluster setup in all experiments.
 The data stream is divided into four portions over two server.
@@
     \includegraphics[width=.9\textwidth, height=.3\textheight]{kafka.png}
-     \caption{The setuop of the Data Stream Producer \& Kafka Cluster.}
+     \caption{The setup of the Data Stream Producer and Kafka Cluster.}
+     \label{fig:data_stream}
 \end{figure}
@@
 \item Handling parallel input streams (e.g., Kafka Stream).
-\item How to aggregate the state of input stream.
+\item How to aggregate the state of an input stream.
 \item Manage the order of stream records.
@@
 \end{itemize}
-We give the description of workloads, the relation with these aspects and  the implementation details of Spark's and Flink's solutions in Sections 3.2.1 \& 3.2.1.
-Afterward, In Section 4 we discuss the performance evaluation and analysis of stream processing workloads in the two frameworks.
+We give the description of workloads, the relation with these aspects and  the implementation details of Spark's and Flink's solutions in Sections~\ref{sec:statistics} and~\ref{sec:sectors}.
+Afterward, In Section~\ref{sec:sec4} we discuss the performance evaluation and analysis of stream processing workloads in the two frameworks.
 \par The two streaming workloads and the data stream producers are implemented in Java and the source code is available in a repository on Github\footnote{\url{https://github.com/ehabqadah/Spark_vs_Flink}}.
 
-\subsubsection{Statistics Computation per Trajectory}
+\subsubsection{Statistics Computation per Trajectory Workload}
+\label{sec:statistics}
 In the first stream processing workload, we construct a stream of trajectories by considering  the position messages (i.e., ADS-B messages) that belong to the same aircraft as trajectory.
 Moreover, we continuously compute and aggregate statistics for each new position in a trajectory .
-As example of computed statistics quantities speed mean, mean of location coordinates, min and max altitude, etc.
-In this workload, we cover the parallel receiving of input data stream, statefull aggregation over the input data stream, and preserving the correct order of the stream's records.
+As example of the computed statistics quantities speed mean, mean of location coordinates, min and max altitude, etc.
+In this workload, we cover the parallel receiving of an input data stream, statefull aggregation over a data stream and preserving the correct order of the stream's records.
 The implementation details of this workload in Spark and Flink are presented in the following: 
@@
 
-\par The implementation of trajectory statistics computation in Spark Streaming, first, it reads the Kafka stream using \texttt{KafkaUtils.createDirectStream} that creates \texttt{DStream} that represents the stream of position messages, in order to scale up and have parallel Kafka stream receivers, a multiple instances of \texttt{DStream} must be created and combined together using the \texttt{union} operation.
-The output stream of the \texttt{union} is further processed using \texttt{mapToPair}, to map each position message to tuple of aircraft ID and position message.
-Then, the irrelevant tuples are filtered based on the position message type using  a \texttt{filter} transformation.
-Afterward, the tuples that share the same aircraft ID are groped together to construct the stream of trajectories (tuples of Id and list of positions), by applying the \texttt{groupByKey} operation.
-Since the processing  model of Spark Streaming is micro-batch based and stateless, \texttt{updateStateByKey} operation must be used to maintain the state of the trajectories stream.
-In the context of this workload, the state is the aggregated statistics of trajectory, so the statistics computation is preformed within the custom \texttt{updateStateByKey} function that is continuously applied for every batch.
- The state update function uses the last position of trajectory from previous batch to aggregate and compute statistics quantities to each trajectory's position in the new data batch, and the new positions list is kept as the new state for the trajectory.
-Inside the state update function, the new positions must be sorted to preserve the correct order, since the tuples after \texttt{groupByKey} operation are shuffled across the cluster's  nodes and sent to the state update function.
+\par The implementation of trajectory statistics computation in Spark Streaming, first, it reads the Kafka stream using \texttt{KafkaUtils.createDirectStream} that creates \texttt{DStream}, which represents the stream of position messages, in order to scale up and have parallel Kafka stream receivers, multiple instances of \texttt{DStream} must be created and combined together using a \texttt{union} operation.
+The output stream of the \texttt{union} is further processed using a \texttt{mapToPair} transformation that  maps each position message to tuple of aircraft ID and position message.
+Then,  irrelevant tuples are filtered based on the position message type using  a \texttt{filter} transformation.
+Afterward, tuples that share the same aircraft ID are groped together to construct a stream of trajectories (tuples of ID and list of positions) by applying the \texttt{groupByKey} operation.
+Since the processing  model of Spark Streaming is micro-batch based and stateless an \texttt{updateStateByKey} operation must be used to maintain the state of the stream of trajectories.
+In the context of this workload, the state is the aggregated statistics of a trajectory, so the statistics computation is preformed within the custom \texttt{updateStateByKey} function that is continuously applied for every batch.
+ The state update function uses the last position of a trajectory from previous batch to aggregate and compute statistics quantities to the received positions within a new batch and new positions are kept as the new state.
+Inside the state update function  the new positions must be sorted to preserve the correct order, since the tuples after \texttt{groupByKey} operation are shuffled across the cluster's  nodes and sent to the state update function.
 
@@
 The details of Flink's implementation of statistics computation workload as follows.
-First, a \texttt{FlinkKafkaConsumer} is used to  construct a \texttt{DataStream} instance that represent the Kafka stream of aircrafts positions, it implicitly runs multiple parallel Kafka stream consumers, which makes the Kafka partitions balanced between the instances.
-Second, the stream records (i.e., ADS-B messages) are parsed and transformed to tuples of ID and position message using a configured \texttt{map} transformation.
+First, a \texttt{FlinkKafkaConsumer} is used to  construct a \texttt{DataStream} instance that represents the Kafka stream of aircrafts positions.
+The  \texttt{FlinkKafkaConsumer} implicitly runs multiple parallel Kafka stream consumers, which makes the Kafka stream's partitions balanced between the parallel instances.
+Second, the stream's records (i.e., ADS-B messages) are parsed and transformed to tuples of ID and position message using a configured \texttt{map} transformation.
 Third, a \texttt{filter} function is used to filter irrelevant tuples based on the message type.
-Fourth, the resulted tuples after the filtration transformation are grouped based on the tuple's ID using \texttt{keyBy} operator that constructs the \texttt{KeyedStream} of trajectories.
+Fourth, the filtered tuples  are grouped based on the tuple's ID using a \texttt{keyBy} operator that constructs a \texttt{KeyedStream} of trajectories.
 Finally,
-a \texttt{reduce} transformation on the  \texttt{KeyedStream} is used to calculate the statistics for each new coming trajectory's position, by utilizing the aggregated statistics of previous position, while old position is discarded and the new position with attached computed statistics is used to aggregate the statistics of a certain trajectory.
+a \texttt{reduce} transformation on the  \texttt{KeyedStream} is used to calculate the statistics for each new arriving trajectory's position by utilizing the aggregated statistics of the previous position.
+The old position is discarded and the new position with computed statistics is used to aggregate the statistics of a trajectory.
 
@@
 
-\par In summary, Flink  handles  the parallel consumers of Kafka stream implicitly, while a multiple \texttt{DStream} must be created in Spark Streaming and union them to have parallel Kafka receivers.
-The Flink's streaming processing is stateful that means there is no need to manage the stream state as the case in Spark streaming.
+\par In summary, Flink  handles  the parallel consumers of Kafka stream implicitly, while multiple \texttt{DStream} must be created in Spark Streaming and union them to have parallel Kafka receivers.
+The Flink's stream processing is stateful, so  the  management of the stream's state is not required in contrast to the Spark streaming.
 A sort action is required in Spark Streaming to preserve the correct  order of the position messages inside the state update function.
 
-\subsubsection{Air Sector Change Detection}
-
+\subsubsection{Air Sector Change Detection Workload}
+\label{sec:sectors}
 The goal of sector change detection workload is to detect the  entering or leaving of an aircraft from one air sector to another one, by processing the real-time stream of aircraft positions.
@@
 Afterward, the tuples that shares a common IDs are grouped together by the \texttt{keyBy} operator.
-Since the Flink streaming processes the stream records as it come, a defined \texttt{reduce} operator assign corresponding sector of position on new tuple, while the old tuple with same ID is used to retrieve the previous sector to attach it to the new tuple, while the old tuple is discarded.
-Finally, the tuples (i.e., tuple of (aircraft ID, position message)) with different previous and current sector are filtered using defined \texttt{filter} transformation.
+Since the Flink processes the stream records as they come, a \texttt{reduce} operator assigns the corresponding sector of position on the new tuple, while the old tuple with same ID is used to retrieve the previous sector to attach it to the new tuple then the old tuple is discarded.
+Finally, the tuples (i.e., tuples of (aircraft ID, position message)) with different previous and current sectors are filtered using a \texttt{filter} transformation.
 
@@
 
- \par To sum up, the same differences between Flink and Spark Streaming regarding the state management and parallel  consumers of Kafka stream as was discussed in Section 3.2.1.
-Another key difference between them, is how  to provide the global data model (e.g., sectors) in a streaming workload.
+ \par To sum up, in this workload we have the  same differences between Flink and Spark Streaming regarding the state management and parallel  consumers of Kafka stream as was discussed in Section~\ref{sec:statistics}.
+Another key difference between them is how  to provide a global data model (e.g., sectors) in a streaming workload.
 Spark Streaming offers the \texttt{Broadcast} feature to handle this issue in efficient way, while the global data model is manually provided  to the stream operations in Flink.
 \section{Performance Results}
-
+\label{sec:sec4}
 \par This section presents the results of performance evaluation in Spark (2.0.2) and Flink (1.1.3) for the streaming workloads that were described in Section 3.2.
@@
 
-Figure 5 shows the results of the latency measurements of the streaming workloads in Spark and Flink.
+Figure~\ref{fig:latency} shows the results of the latency measurements of the streaming workloads in Spark and Flink.
 The presented values present the average value of multiple trails and measurements.
@@
 \caption{Latency Measurements}
-\label{fig:fig}
+\label{fig:latency}
 \end{figure}
@@
 
-The  measurements of throughput for the two streaming workloads are given in Figure 6. The presented values reflect the average of throughput (message/minute) 
+The  measurements of throughput for the two streaming workloads are given in Figure~\ref{fig:throughput}. The presented values reflect the average of throughput (message/minute) 
  in Spark streaming and Flink for multiple executions. Also for Spark streaming we tested different 
@@
 \caption{Throughput Measurements}
-\label{fig:fig}
+\label{fig:throughput}
 \end{figure}
@@
 
+\label{sec:sec5}
 In this report, we present an experimental evaluation of the stream processing performance and capabilities of Apache Spark and Apache Flink.
 We carried our experiment over aircraft positions datasets provided by the Datacorn project, we developed two streaming workloads to cover general aspects of real-time stream processing tasks.
-Additionally, and we measured the latency and throughput of those workloads solution of Spark and Flink.
+Additionally,  we measured the latency and throughput of those workloads solution of Spark and Flink.
 Results show that Flink outperforms Spark Streaming in term of processing latency.
 In contrast, Spark Streaming has showed better throughput rates than Flink.
-\par To sum up, Flink's processing model is well-suited to the stream processing tasks that require low latency, also there is no need to manage the Stream's state and the batch interval in contrast to Spark streaming.
+\par To sum up, Flink's processing model is well-suited to the stream processing tasks that require low latency, also there is no need to manage the Stream's state and the batch interval comparing to Spark streaming.
 Moreover, Flink operates over the individual record of the data stream, while Spark Streaming  processes the input data stream as micro-batches.
@@
 
-In future work, we plan to carry our experiments over cluster of computers not just only on single computer.
+In future work, we plan to carry out our experiments over cluster of computers not just only on single computer.
 In addition, we will cover more aspects of real-time data stream processing and developing more evaluation workloads.

### 2017-03-09 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesSectorChangeDetector.java
finished throughput measurements

@@ -29,7 +29,7 @@ public class TrajectoriesSectorChangeDetector {
 		StreamExecutionEnvironment env = FlinkUtils.getInitializedEnv();
 		KeyedStream<Tuple2<String, PositionMessage>, Tuple> trajectoriesStream = TrajectoriesStreamUtils
 				.getTrajectoriesStream(env);
-
+		LoggerUtils.logMessage("Start Time:" + System.currentTimeMillis());
 		DataSet<Sector> sectorsDataSet = getSectorDataSet();
 
 		// assign sector for each trajectory, then check if there is change in
@@ -52,7 +52,7 @@ public class TrajectoriesSectorChangeDetector {
 			String outputLine = tuple.f0 + ": " + trajectory.getPrevSector().getNameAndAirBlock() + " --> "
 					+ trajectory.getSector().getNameAndAirBlock();
 
-			LoggerUtils.logMessage(outputLine);
+			// LoggerUtils.logMessage(outputLine);
 			return outputLine;
 
 		});

### 2017-03-09 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesSectorsReducer.java

@@ -5,6 +5,7 @@ import java.util.List;
 import org.apache.flink.api.common.functions.ReduceFunction;
 import org.apache.flink.api.java.tuple.Tuple2;
 
+import de.kdml.bigdatalab.spark_and_flink.common_utils.LoggerUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.SectorUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
@@ -42,6 +43,8 @@ public class TrajectoriesSectorsReducer implements ReduceFunction<Tuple2<String,
 
 		newTuple.f1.setNew(false);
 
+		LoggerUtils.logMessage("-" + System.currentTimeMillis());
+		LoggerUtils.logMessage("-" + System.currentTimeMillis());
 		return newTuple;
 	}
 

### 2017-03-09 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesStreamUtils.java

@@ -70,10 +70,10 @@ public class TrajectoriesStreamUtils {
 			// parse ADS-B messages and construct tuple of ID & position
 			// message for each message
 			StreamRecord streamRecord = StreamRecord.parseData(line);
-			PositionMessage trajectory = TrajectoriesUtils.parseDataInput(streamRecord.getValue());
-			trajectory.setStreamedTime(streamRecord.getStreamedTime());
-			trajectory.setNew(true);
-			return new Tuple2<>(trajectory.getID(), trajectory);
+			PositionMessage position = TrajectoriesUtils.parseDataInput(streamRecord.getValue());
+			position.setStreamedTime(streamRecord.getStreamedTime());
+			position.setNew(true);
+			return new Tuple2<>(position.getID(), position);
 
 		}).filter(tuple -> {
 			// filter irrelevant messages

### 2017-03-09 flink_project/src/main/java/kafka/datacorn/AircraftPositionsStreamGenerator.java

@@ -33,15 +33,18 @@ public class AircraftPositionsStreamGenerator implements SourceFunction<String> 
 			for (String file : adsbFiles) {
 
 				System.out.println("streaming of " + file);
-				i++;
+
 				try (BufferedReader br = new BufferedReader(new FileReader(file))) {
 					String messageLine;
 					while ((messageLine = br.readLine()) != null) {
+						i++;
 
 						// append streaming time
 						messageLine = new StreamRecord(messageLine).toString();
+
 						ctx.collect(messageLine);
-						Thread.sleep(LINE_SLIDE_TIME_MS);
+
+						// Thread.sleep(LINE_SLIDE_TIME_MS);
 					}
 				} catch (Exception e) {
 

### 2017-03-09 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesSectorChangeDetector.java

@@ -14,6 +14,7 @@ import org.apache.spark.streaming.api.java.JavaPairDStream;
 import org.apache.spark.streaming.api.java.JavaStreamingContext;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.LoggerUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.SectorUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.TrajectoriesUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
@@ -181,6 +182,7 @@ public class TrajectoriesSectorChangeDetector {
 				 */
 				appendTrajectoryAndAssignSector(aggregatedPositionsOfTrajectory, position);
 				position.setFinishProcessingTime(System.currentTimeMillis());
+				LoggerUtils.logMessage(System.currentTimeMillis() + "-");
 
 			}
 			// update state

### 2017-03-10 paper/paper.tex
[paper] spell checking

@@
 The experiment is done using multiple streaming workloads over real-world datasets.
-Results show that Flink outperform Spark streaming in term of latency.
+Results show that Flink outperforms Spark streaming in term of latency.
 While Spark Streaming gives higher throughput rates than Flink.
@@
 Moreover, many Big Data frameworks have been recently developed such as Apache Spark and Apache Flink that provide efficient data processing functionalities.
- \par Big Data is a term that does not only describe large datasets (volume), but also the data that arrives in high rate (velocity), and it comes in a mix of different data format (variety) (i.e., 3Vs of Big Data (Volume, Velocity, Variety)) \cite{svs}.
+ \par Big Data is a term that does not only describe large datasets (volume), but also the data that arrives at a high rate (velocity), and it comes in a mix of different data format (variety) (i.e., 3Vs of Big Data (Volume, Velocity, Variety)) \cite{svs}.
 
 \par The processing models of Big Data are Batch processing and Stream processing.
-Batch processing involves handling  large, finite volumes of data (e.g., processing of historical transaction records), and is more concerned with throughput.
-While Stream processing  is a method to manage fast and continuously incoming data (real-time), as it arrives to the system (e.g., credit card fraud detection and network monitoring applications).
+Batch processing involves handling  large and finite volumes of data (e.g., processing of historical transaction records), and is more concerned with throughput.
+While Stream processing  is a method to manage fast and continuously incoming data (real-time) as it arrives in the system (e.g., credit card fraud detection and network monitoring applications).
 Stream processing systems usually  provide low latency processing.
@@
 \par There exist some publications that compare Apache Spark and Apache Flink
- such as the Yahoo Streaming Benchmark \cite{yahoo}. But we believe that we still need to carry our evaluation for two reasons: \begin{enumerate*}[label=(\roman*)]
-\item both frameworks are continuously improving in short time 
+ such as the Yahoo Streaming Benchmark \cite{yahoo}. But we believe that we still need to carry out our evaluation for two reasons: \begin{enumerate*}[label=(\roman*)]
+\item both frameworks are continuously improving in a short time 
 \item we aim to study the performance of some complex operations such as \texttt{groupByKey} \& \texttt{keyBy} in Apache Spark and Apache Flink, respectively.
@@
 In order to simulate real-world use cases, we use datasets of airplane trajectories provided by the  Datacron project\footnote{\url{http://www.datacron-project.eu/}} to develop multiple streaming data processing workloads.
-Beside the performance comparison of the stream processing of the two frameworks, we try to cover general aspects of stream processing tasks (see Section~\ref{sec:sec3}) and how they can be solved in each platform.
+Besides the performance comparison of the stream processing of the two frameworks, we try to cover general aspects of stream processing tasks (see Section~\ref{sec:sec3}) and how they can be solved in each platform.
 Informally,  this work aims to help the developers to determine which platform to use for different use cases.
@@
 Section~\ref{sec:sec4}
-compares the performance results of the different workloads. And finally, Section~\ref{sec:sec5} gives the overall conclusion.
+compares the performance results of the different workloads. And finally, Section~\ref{sec:sec5} gives the overall conclusion and future work.
 
@@
 \label{sec:sec2}
- In this section, we present the architecture, key characteristics, programming APIs  of Apache Spark and Apache Flink. Moreover, a brief overview of the Apache Kafka platform is  provided.
+ In this section, we present the architecture, key characteristics and the programming APIs  of Apache Spark and Apache Flink. Moreover, a brief overview of the Apache Kafka platform is  provided.
 
@@
 \par Apache Flink is an open source project that provides a large-scale, distributed stream processing platform \cite{flink}.
-Flink treats the batch processing as special case of streaming applications (i.e., finite stream).
+Flink treats the batch processing as a special case of streaming applications (i.e., finite stream).
 It offers programming APIs in Java and Scala.
@@
 
-\par The main data abstraction of Flink are DataStream and DataSet that represent read-only collection of data elements.
-The list of elements is bounded (i.e., finite) in DataSet, while it is unbounded (i.e., infinite) in case of DataStream.
+\par The main data abstractions of Flink are DataStream and DataSet that represent read-only collection of data elements.
+The list of elements is bounded (i.e., finite) in DataSet, while it is unbounded (i.e., infinite) in the case of DataStream.
 
@@
 The evaluation streaming workloads read an input data stream from Kafka.
-In order to simulate real-world use case we use datasets of Automatic Dependent Surveillance – Broadcast (ADS-B) messages that represent the aircraft's position over time.
+In order to simulate real-world use case we use datasets of Automatic Dependent Surveillance - Broadcast (ADS-B) messages that represent the aircraft's position over time.
 The message line comprises 22 fields of data such as aircraft ID, date message generated, longitude, latitude and altitude.
@@
 Figure~\ref{fig:data_stream} illustrates the data producer and the Kafka cluster setup in all experiments.
-The data stream is divided into four portions over two server.
+The data stream is divided into four portions over two servers.
 The data producer publishes the stream records randomly to a Kafka partition and all ADS-B messages are published to the same topic (\texttt{datacorn}).
@@
 \subsection{Evaluation Streaming Workloads}
-In our experimental evaluation, we developed two real-time stream processing workloads that read a ADS-B messages stream from Kafka in order to perform basic trajectories analysis methods.
+In our experimental evaluation, we developed two real-time stream processing workloads that read a stream of ADS-B messages from Kafka in order to perform basic trajectories analysis methods.
 The workloads are designed to cover general key aspects of real-time stream processing tasks and evaluate the corresponding solutions in Spark and Flink.
-The following are the aspects of streaming data processing were covered by our workloads's design:
+The following are the aspects of streaming data processing were covered by our workloads design:
 
@@
 Afterward, In Section~\ref{sec:sec4} we discuss the performance evaluation and analysis of stream processing workloads in the two frameworks.
-\par The two streaming workloads and the data stream producers are implemented in Java and the source code is available in a repository on Github\footnote{\url{https://github.com/ehabqadah/Spark_vs_Flink}}.
+\par The two streaming workloads and the data stream producer are implemented in Java and the source code is available in a repository on Github\footnote{\url{https://github.com/ehabqadah/Spark_vs_Flink}}.
 
@@
 In the first stream processing workload, we construct a stream of trajectories by considering  the position messages (i.e., ADS-B messages) that belong to the same aircraft as trajectory.
-Moreover, we continuously compute and aggregate statistics for each new position in a trajectory .
-As example of the computed statistics quantities speed mean, mean of location coordinates, min and max altitude, etc.
-In this workload, we cover the parallel receiving of an input data stream, statefull aggregation over a data stream and preserving the correct order of the stream's records.
+Moreover, we continuously compute and aggregate statistics for each new position in a trajectory.
+As an example of the computed statistics quantities speed mean, mean of location coordinates, min and max altitude, etc.
+In this workload, we cover the parallel receiving of an input data stream, stateful aggregation over a data stream and preserving the correct order of the stream's records.
 The implementation details of this workload in Spark and Flink are presented in the following: 
@@
 \par The implementation of trajectory statistics computation in Spark Streaming, first, it reads the Kafka stream using \texttt{KafkaUtils.createDirectStream} that creates \texttt{DStream}, which represents the stream of position messages, in order to scale up and have parallel Kafka stream receivers, multiple instances of \texttt{DStream} must be created and combined together using a \texttt{union} operation.
-The output stream of the \texttt{union} is further processed using a \texttt{mapToPair} transformation that  maps each position message to tuple of aircraft ID and position message.
-Then,  irrelevant tuples are filtered based on the position message type using  a \texttt{filter} transformation.
-Afterward, tuples that share the same aircraft ID are groped together to construct a stream of trajectories (tuples of ID and list of positions) by applying the \texttt{groupByKey} operation.
+The output stream of the \texttt{union} is further processed using a \texttt{mapToPair} transformation that  maps each position message to a tuple of aircraft ID and position message.
+Then,  irrelevant tuples are filtered based on the position message type using  a \texttt{filter} transformation.\par Afterward, tuples that share the same aircraft ID are grouped together to construct a stream of trajectories (tuples of ID and list of positions) by applying the \texttt{groupByKey} operation.
 Since the processing  model of Spark Streaming is micro-batch based and stateless an \texttt{updateStateByKey} operation must be used to maintain the state of the stream of trajectories.
-In the context of this workload, the state is the aggregated statistics of a trajectory, so the statistics computation is preformed within the custom \texttt{updateStateByKey} function that is continuously applied for every batch.
- The state update function uses the last position of a trajectory from previous batch to aggregate and compute statistics quantities to the received positions within a new batch and new positions are kept as the new state.
+In the context of this workload, the state is the aggregated statistics of a trajectory, so the statistics computation is performed within the custom \texttt{updateStateByKey} function that is continuously applied for every batch.
+ \par The state update function uses the last position of a trajectory from previous batch to aggregate and compute statistics quantities to the received positions within a new batch and new positions are kept as the new state.
 Inside the state update function  the new positions must be sorted to preserve the correct order, since the tuples after \texttt{groupByKey} operation are shuffled across the cluster's  nodes and sent to the state update function.
@@
 Third, a \texttt{filter} function is used to filter irrelevant tuples based on the message type.
-Fourth, the filtered tuples  are grouped based on the tuple's ID using a \texttt{keyBy} operator that constructs a \texttt{KeyedStream} of trajectories.
-Finally,
+Fourth, the filtered tuples  are grouped based on the tuple's ID using a \texttt{keyBy} operator that constructs a \texttt{KeyedStream} of trajectories.\par Finally,
 a \texttt{reduce} transformation on the  \texttt{KeyedStream} is used to calculate the statistics for each new arriving trajectory's position by utilizing the aggregated statistics of the previous position.
@@
 The goal of sector change detection workload is to detect the  entering or leaving of an aircraft from one air sector to another one, by processing the real-time stream of aircraft positions.
-Given the fact that a dataset of sectors (i.e., polygons) is available as a reference to assign the corresponding sector for certain aircraft's position.
+Given the fact that a dataset of 20,000 sectors (i.e., polygons) is available as a reference to assign the corresponding sector for certain aircraft's position.
 In this workload, we test the aspect of providing a global data model (sectors dataset) in stream processing workload.
@@
  We implemented the sector change detection workload in Spark Streaming by constructing the stream of trajectories and assigning a sector for each position of a trajectory.
-A change between the assigned sectors of two consecutive aircraft's positions is marked  as a transition of sector for a certain trajectory.
-\par First, multiple instances of \texttt{DStream} are created using \texttt{KafkaUtils.create} \texttt{DirectStream}  are combined using \texttt{union} to read in  parallel from Kafka.
+A change between the assigned sectors of two consecutive aircraft's positions is marked  as a sector's transition for a certain trajectory.
+\par First, multiple instances of \texttt{DStream} are created using the \texttt{KafkaUtils.create} \texttt{DirectStream} and combined using a \texttt{union} to read in  parallel from Kafka.
 Second, the records (i.e., ADS-B messages) of the resulting stream are parsed and mapped to tuples of (aircraft ID, position message) using \texttt{mapToPair} transformation.
@@
 The state update function assigns the corresponding sector for each new position, the old positions of a certain trajectory from the previous batch are discarded, but only the last position of the old batch is kept to be used alongside with the new positions.
-In addition to that the new positions are sorted by th streaming time to maintain the correct order.
+In addition to that the new positions are sorted by the streaming time to maintain the correct order.
  \par The sector data set is provided using the built-in \texttt{Broadcast} feature that distributes a shared value for  all cluster nodes in an efficient matter.
@@
 Irrelevant tuples are filtered based on the message type using a \texttt{filter} operation.
-Afterward, the tuples that shares a common IDs are grouped together by the \texttt{keyBy} operator.
-Since the Flink processes the stream records as they come, a \texttt{reduce} operator assigns the corresponding sector of position on the new tuple, while the old tuple with same ID is used to retrieve the previous sector to attach it to the new tuple then the old tuple is discarded.
+Afterward, the tuples that share a common ID are grouped together by the \texttt{keyBy} operator.
+Since the Flink processes the stream records as they come, a \texttt{reduce} operator assigns the corresponding sector of position on the new tuple, while the old tuple with the same ID is used to retrieve the previous sector to attach it to the new tuple then the old tuple is discarded.
 Finally, the tuples (i.e., tuples of (aircraft ID, position message)) with different previous and current sectors are filtered using a \texttt{filter} transformation.
@@
 Another key difference between them is how  to provide a global data model (e.g., sectors) in a streaming workload.
-Spark Streaming offers the \texttt{Broadcast} feature to handle this issue in efficient way, while the global data model is manually provided  to the stream operations in Flink.
+Spark Streaming offers the \texttt{Broadcast} feature to handle this issue in an efficient way, while the global data model is manually provided  to the stream operations in Flink.
 \section{Performance Results}
@@
 \par We consider the time delay between the streaming time of a record in Kafka and  the time when the record is processed by a workload as the platform's processing latency (measured in milliseconds).
-And the number of processed stream's records (aircrafts position messages) per second within each streaming workload is the throughput measure.
+And the number of processed stream's records (aircraft positions) per second within each streaming workload is the throughput measure.
 Since the Spark Streaming is micro-batch based,  we examine the effect of different batch intervals
@@
 Measurements for different batch durations are given in Spark Streaming, we noticed that making the batch time too small does not give better latency results.
-Results show that Flink outperform Spark streaming in term of processing latency in the two evaluation workloads.
+Results show that Flink outperforms Spark Streaming in term of processing latency in the two evaluation workloads.
 
@@
 Results show that Flink outperforms Spark Streaming in term of processing latency.
-In contrast, Spark Streaming has showed better throughput rates than Flink.
+In contrast, Spark Streaming provides better throughput rates than Flink.
 \par To sum up, Flink's processing model is well-suited to the stream processing tasks that require low latency, also there is no need to manage the Stream's state and the batch interval comparing to Spark streaming.
@@
 In future work, we plan to carry out our experiments over cluster of computers not just only on single computer.
-In addition, we will cover more aspects of real-time data stream processing and developing more evaluation workloads.
+In addition, we will cover more aspects of real-time data stream processing and develop more evaluation workloads.
 

### 2017-03-13 paper/paper.tex
[paper] update report with feedback from PD Dr. Mock

@@
 create large volumes of data.
-Moreover, many Big Data frameworks have been recently developed such as Apache Spark and Apache Flink that provide efficient data processing functionalities.
+Moreover, many Big Data frameworks have been recently developed such as Apache Spark \cite{spark} and Apache Flink \cite{flink} that provide efficient data processing functionalities.
  \par Big Data is a term that does not only describe large datasets (volume), but also the data that arrives at a high rate (velocity), and it comes in a mix of different data format (variety) (i.e., 3Vs of Big Data (Volume, Velocity, Variety)) \cite{svs}.
@@
 \par There exist some publications that compare Apache Spark and Apache Flink
- such as the Yahoo Streaming Benchmark \cite{yahoo}. But we believe that we still need to carry out our evaluation for two reasons: \begin{enumerate*}[label=(\roman*)]
+ such as the Yahoo Streaming Benchmark \cite{yahoo}.
+The benchmarking provides a performance comparison of Apache Flink , Apache Storm \cite{sortm} and  Spark Streaming in terms of latency and throughput over a single data pipeline.
+The results show that Flink and Storm have similar and lower latency, while Spark Streaming has higher latency and it gives higher throughput by increasing the batch duration.
+But we believe that we still need to carry out our evaluation for two reasons: \begin{enumerate*}[label=(\roman*)]
 \item both frameworks are continuously improving in a short time 
@@
 In other words, the Spark Streaming is a batch-based processing model, which processes the continuous stream of data by dividing it into micro-batches that are  processed by the Spark engine.
+The operations of Spark Streaming are stateless due to the micro-batch design, the \texttt{updateStateByKey} transformation can be used to aggregate the state between stream's micro batches.
 Figure~\ref{fig:spark} illustrates the general process flow of Spark Streaming.
@@
 \par The Flink's core is a distributed streaming dataflow engine,  with each Flink program is represented by a dataflow graph (i.e., directed acyclic graph - DAG) that executed by the Flink's engine \cite{flink_paper}.
-The data flow graphs are composed of stateful, parallel operations and intermediate data stream partitions.
+The data flow graphs are composed of stateful (state is maintained per partition), parallel operations and intermediate data stream partitions.
 Figure~\ref{fig:fstream} shows a data flow graph of a Flink program.
@@
 The message line comprises 22 fields of data such as aircraft ID, date message generated, longitude, latitude and altitude.
-And  these datasets contain around \texttt{26 millions} messages.
+And  these datasets (2.4 GB) contain around \texttt{26 millions}  messages.
 \par In our experiment a data stream producer component reads the ADS-B message datasets, then publishes them to the Kafka cluster to be consumed by the workloads in Apache Spark and Apache Flink.
@@
 
-\par The implementation of trajectory statistics computation in Spark Streaming, first, it reads the Kafka stream using \texttt{KafkaUtils.createDirectStream} that creates \texttt{DStream}, which represents the stream of position messages, in order to scale up and have parallel Kafka stream receivers, multiple instances of \texttt{DStream} must be created and combined together using a \texttt{union} operation.
+\par The implementation of trajectory statistics computation in Spark Streaming, first, it reads the Kafka stream using KafkaUtils.createDirectStream that creates \texttt{DStream}, which represents the stream of position messages, in order to scale up and have parallel Kafka stream receivers, multiple instances of \texttt{DStream} must be created and combined together using a \texttt{union} operation.
 The output stream of the \texttt{union} is further processed using a \texttt{mapToPair} transformation that  maps each position message to a tuple of aircraft ID and position message.
@@
 
-\par In summary, Flink  handles  the parallel consumers of Kafka stream implicitly, while multiple \texttt{DStream} must be created in Spark Streaming and union them to have parallel Kafka receivers.
-The Flink's stream processing is stateful, so  the  management of the stream's state is not required in contrast to the Spark streaming.
-A sort action is required in Spark Streaming to preserve the correct  order of the position messages inside the state update function.
+\par In summary, The following are the main key differences between the Flink and Spark Streaming implementation of the statistics computation workload:
+
+\begin{itemize}
+\item  Flink  handles  the parallel consumers of Kafka stream implicitly, while multiple \texttt{DStream} must be created in Spark Streaming and union them to have parallel Kafka receivers.
+\item The Flink's operations of the \texttt{KeyedStream} are stateful, so  the  management of the stream's state is not required as Spark Streaming that requires the usage of \texttt{updateStateByKey} to manage the state of the micro batches.
+\item A sort action is required in Spark Streaming to preserve the correct  order of the position messages inside the state update function.
+while in Flink the sort is not required by using a \texttt{reduce} transformation over the \texttt{KeyedStream} that processes the stream records item by item.
+\end{itemize}
 
@@
 A change between the assigned sectors of two consecutive aircraft's positions is marked  as a sector's transition for a certain trajectory.
-\par First, multiple instances of \texttt{DStream} are created using the \texttt{KafkaUtils.create} \texttt{DirectStream} and combined using a \texttt{union} to read in  parallel from Kafka.
+\par First, multiple instances of \texttt{DStream} are created using the KafkaUtils.create \texttt{DirectStream} and combined using a \texttt{union} to read in  parallel from Kafka.
 Second, the records (i.e., ADS-B messages) of the resulting stream are parsed and mapped to tuples of (aircraft ID, position message) using \texttt{mapToPair} transformation.
@@
 In addition to that the new positions are sorted by the streaming time to maintain the correct order.
- \par The sector data set is provided using the built-in \texttt{Broadcast} feature that distributes a shared value for  all cluster nodes in an efficient matter.
+ \par The sector data set is provided using the built-in \texttt{Broadcast} feature that distributes a shared value for  all cluster nodes in an efficient matter, and it can be update at the driver program by using the \texttt{unpersist} function and then update the sectors.
 Finally, the detection of a sector change is done by an extra filtering transformation using a \texttt{filter} operator, which checks the sectors of consecutive positions in a trajectory to find the ones with a change in the sector assignment.
@@
 
- \par To sum up, in this workload we have the  same differences between Flink and Spark Streaming regarding the state management and parallel  consumers of Kafka stream as was discussed in Section~\ref{sec:statistics}.
-Another key difference between them is how  to provide a global data model (e.g., sectors) in a streaming workload.
+ \par To sum up, The following are the main key differences between the Flink and Spark Streaming implementation of the sector change detection workload:
+
+ \begin{itemize}
+ \item  In this workload we have the  same differences between Flink and Spark Streaming regarding the state management, parallel  consumers of Kafka stream and preserving the correct order of stream records as was discussed in Section~\ref{sec:statistics}.
+ \item Another key difference between them is how  to provide a global data model (e.g., sectors) in a streaming workload.
 Spark Streaming offers the \texttt{Broadcast} feature to handle this issue in an efficient way, while the global data model is manually provided  to the stream operations in Flink.
+ \item The broadcast variable of sectors in Spark Streaming can be updated in the driver program by using the \texttt{unpersist} function and then update it.
+While in Flink the update of the sectors is not possible at run time (the program must be reloaded), since they are provided manually to the transformation.
+ \end{itemize}
+
 \section{Performance Results}
@@
 The presented values present the average value of multiple trails and measurements.
-Measurements for different batch durations are given in Spark Streaming, we noticed that making the batch time too small does not give better latency results.
+Measurements for different batch durations are given in Spark Streaming to find the optimal batch duration, since  the large batch times give higher latency rates, also we noticed that reducing the batch time to smaller values increases the latency because of the overhead of managing more micro batches by Spark, this behavior also was observed in \cite{yahoo}.
 Results show that Flink outperforms Spark Streaming in term of processing latency in the two evaluation workloads.
+The two framework give higher latency in the sectors change detection workload than statistics computation workload due to the sector assignment for each position by traversing the 20,000 sectors until a match is found.
 
@@
 \centering
-\begin{subfigure}{.7\textwidth}
+\begin{subfigure}{.6\textwidth}
   \centering
@@
 
-The  measurements of throughput for the two streaming workloads are given in Figure~\ref{fig:throughput}. The presented values reflect the average of throughput (message/minute) 
- in Spark streaming and Flink for multiple executions. Also for Spark streaming we tested different 
- batch times. Results show that Spark Streaming has higher throughput than Flink, also  increasing the batch duration of Spark Streaming gives better throughput.
+The  measurements of throughput for the two streaming workloads are given in Figure~\ref{fig:throughput}.
+The presented values reflect the average of throughput (message/minute) in Spark streaming and Flink for multiple executions.
+Also for Spark streaming we tested different
+ batch times including the throughput of the best latency batch time.
+Results show that Spark Streaming has higher throughput than Flink; increasing the batch duration of Spark Streaming gives better throughput.
+Moreover, the throughput rates of Spark Streaming and Flink at the low latency are nearly similar (rates of Flink little higher).
+The two framework have a lower throughput in the second workload than the first one due to the sector assignment for each position.
 
@@
 Results show that Flink outperforms Spark Streaming in term of processing latency.
-In contrast, Spark Streaming provides better throughput rates than Flink.
+In contrast, Spark Streaming provides better throughput rates than Flink by increasing the batch duration, while Flink gives a similar throughput to Spark Streaming with small batch sizes.
+Our results in terms of latency and throughput are consistent with the Yahoo Streaming Benchmark results.
 \par To sum up, Flink's processing model is well-suited to the stream processing tasks that require low latency, also there is no need to manage the Stream's state and the batch interval comparing to Spark streaming.
@@
 
-\bibitem[3]{yahoo} 
+\bibitem[3]{spark} 
+Apache Spark. Available: https://spark.apache.org/ [Accessed February 2017].
+
+\bibitem[4]{flink} 
+Apache Flink. Available: https://flink.apache.org/ [Accessed February 2017].
+
+\bibitem[5]{yahoo} 
 Chintapalli, Sanket, et al.
@@
 
-\bibitem[4]{spark} 
-Apache Spark. Available: https://spark.apache.org/ [Accessed February 2017].
-
-\bibitem[5]{rdd}
+\bibitem[6]{sortm} 
+Apache Storm. Available: http://storm.apache.org/ [Accessed March 2017].
+\bibitem[7]{rdd}
 Zaharia, Matei, et al.
@@
 
-\bibitem[6]{spark_streaming}
-Spark Streaming. Available: http://spark.apache.org/docs/latest/streaming-programming-guide.html [Accessed March 2017].
+\bibitem[8]{spark_streaming}
+Spark Streaming. Available: http://spark.apache.org/docs/latest/
+streaming-programming-guide.html [Accessed March 2017].
 
-\bibitem[7]{flink} 
-Apache Flink. Available: https://flink.apache.org/ [Accessed February 2017].
-
-\bibitem[8]{flink_paper} 
+\bibitem[9]{flink_paper} 
 Carbone, P., Katsifodimos, A., Ewen, S., Markl, V., Haridi, S. et al. (2015)
@@
 
-\bibitem[9]{kafka} 
+\bibitem[10]{kafka} 
 Apache Kafka. Available: https://kafka.apache.org/intro.html [Accessed February 2017].

### 2017-03-13 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/SectorUtils.java
handle unknown sectors

@@ -2,8 +2,8 @@ package de.kdml.bigdatalab.spark_and_flink.common_utils;
 
 import java.util.List;
 
-import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
+import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
 
 /**
  * 
@@ -29,7 +29,7 @@ public class SectorUtils {
 
 			}
 		}
-		return null;
+		return new Sector("UNKNOWN_SECTOR", "UNKNOWN_BLOCK");
 
 	}
 }

### 2017-03-13 common_utils/src/main/java/de/kdml/bigdatalab/spark_and_flink/common_utils/data/Sector.java

@@ -2,8 +2,6 @@ package de.kdml.bigdatalab.spark_and_flink.common_utils.data;
 
 import java.io.Serializable;
 
-import javax.sound.sampled.Line;
-
 import com.vividsolutions.jts.geom.Polygon;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.GeoUtils;
@@ -27,6 +25,12 @@ public class Sector implements Serializable {
 	private Polygon polygon;
 
 	public Sector() {
+	}
+
+	public Sector(String name, String airBlockName) {
+		setName(name);
+		setAirBlockName(airBlockName);
+
 	}
 
 	public String getName() {

### 2017-03-13 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesSectorChangeDetector.java
latency and throuput measurements

@@ -52,7 +52,7 @@ public class TrajectoriesSectorChangeDetector {
 			String outputLine = tuple.f0 + ": " + trajectory.getPrevSector().getNameAndAirBlock() + " --> "
 					+ trajectory.getSector().getNameAndAirBlock();
 
-			// LoggerUtils.logMessage(outputLine);
+			LoggerUtils.logMessage(outputLine);
 			return outputLine;
 
 		});

### 2017-03-13 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesSectorsReducer.java

@@ -5,7 +5,6 @@ import java.util.List;
 import org.apache.flink.api.common.functions.ReduceFunction;
 import org.apache.flink.api.java.tuple.Tuple2;
 
-import de.kdml.bigdatalab.spark_and_flink.common_utils.LoggerUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.SectorUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.Sector;
@@ -43,8 +42,6 @@ public class TrajectoriesSectorsReducer implements ReduceFunction<Tuple2<String,
 
 		newTuple.f1.setNew(false);
 
-		LoggerUtils.logMessage("-" + System.currentTimeMillis());
-		LoggerUtils.logMessage("-" + System.currentTimeMillis());
 		return newTuple;
 	}
 

### 2017-03-13 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesStatistics.java

@@ -40,16 +40,9 @@ public class TrajectoriesStatistics {
 					Tuple2<String, PositionMessage> newTuple = tuple2.f1.getStreamedTime() < tuple1.f1.getStreamedTime()
 							? tuple1 : tuple2;
 
-					// to count first tuple in the throughput
-					// if (oldTuple.f1.getSpeed() == null) {
-
-					// LoggerUtils.logMessage("-" + System.currentTimeMillis());
-					// }
-
 					StatisticsUtils.computeStatistics(oldTuple.f1, newTuple.f1);
 					newTuple.f1.setNew(false);
 
-					LoggerUtils.logMessage("-" + System.currentTimeMillis());
 					return newTuple;
 				});
 

### 2017-03-13 flink_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/flink/datacorn/TrajectoriesStreamUtils.java

@@ -45,7 +45,7 @@ public class TrajectoriesStreamUtils {
 		props.setProperty("zookeeper.connect", configs.getStringProp("zookeeper"));
 		props.setProperty("bootstrap.servers", configs.getStringProp("bootstrap.servers"));
 		props.setProperty("group.id", configs.getStringProp("kafkaGroupId"));
-		// Always read topicfrom start 
+		// Always read topicfrom start
 		props.setProperty("auto.offset.reset", "earliest");
 
 		// create a Kafka consumer
@@ -73,6 +73,8 @@ public class TrajectoriesStreamUtils {
 			PositionMessage position = TrajectoriesUtils.parseDataInput(streamRecord.getValue());
 			position.setStreamedTime(streamRecord.getStreamedTime());
 			position.setNew(true);
+
+			// LoggerUtils.logMessage("-=-" + System.currentTimeMillis());
 			return new Tuple2<>(position.getID(), position);
 
 		}).filter(tuple -> {

### 2017-03-13 flink_project/src/main/java/kafka/datacorn/AircraftPositionsStreamGenerator.java

@@ -51,7 +51,7 @@ public class AircraftPositionsStreamGenerator implements SourceFunction<String> 
 					System.out.println("exp" + e);
 				}
 			}
-			// running = false;
+			running = false;
 
 		}
 	}

### 2017-03-13 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesSectorChangeDetector.java

@@ -14,7 +14,6 @@ import org.apache.spark.streaming.api.java.JavaPairDStream;
 import org.apache.spark.streaming.api.java.JavaStreamingContext;
 
 import de.kdml.bigdatalab.spark_and_flink.common_utils.Configs;
-import de.kdml.bigdatalab.spark_and_flink.common_utils.LoggerUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.SectorUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.TrajectoriesUtils;
 import de.kdml.bigdatalab.spark_and_flink.common_utils.data.PositionMessage;
@@ -52,7 +51,7 @@ public class TrajectoriesSectorChangeDetector {
 		JavaPairDStream<String, Iterable<PositionMessage>> runningTrajectories = trajectories
 				.updateStateByKey(updateTrajectoriesAndAssignSectors);
 
-		TrajectoriesStreamUtils.printLatencies(runningTrajectories);
+		// TrajectoriesStreamUtils.printLatencies(runningTrajectories);
 		// find trajectories with change in sector
 		JavaDStream<String> changedTrajectories = runningTrajectories.filter(trajectoryTuple -> {
 
@@ -182,7 +181,7 @@ public class TrajectoriesSectorChangeDetector {
 				 */
 				appendTrajectoryAndAssignSector(aggregatedPositionsOfTrajectory, position);
 				position.setFinishProcessingTime(System.currentTimeMillis());
-				LoggerUtils.logMessage(System.currentTimeMillis() + "-");
+				// LoggerUtils.logMessage(System.currentTimeMillis() + "-");
 
 			}
 			// update state

### 2017-03-13 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesStatistics.java

@@ -37,7 +37,6 @@ public class TrajectoriesStatistics {
 
 		// configure spark streaming context
 		JavaSparkContext sc = SparkConfigsUtils.getSparkContext("Trajectories Statistics Computation ");
-
 		long batchTime = configs.getIntProp("batchDuration");
 		// configure the stream batch interval
 		JavaStreamingContext jssc = new JavaStreamingContext(sc, Durations.milliseconds(batchTime));
@@ -81,8 +80,9 @@ public class TrajectoriesStatistics {
 				count += Iterators.size(tuple._2().iterator());
 			}
 
-			LoggerUtils.logMessage(time + " count=" + count + "Current throughput = " + (count / batchTimeInSeconds)
-					+ " records / second");
+			// LoggerUtils.logMessage(time + " count=" + count + "Current
+			// throughput = " + (count / batchTimeInSeconds)
+			// + " records / second");
 		});
 	}
 
@@ -129,7 +129,7 @@ public class TrajectoriesStatistics {
 				StatisticsUtils.computeStatistics(lastOldPosition, position);
 				long currentTime = System.currentTimeMillis();
 				position.setFinishProcessingTime(currentTime);
-				LoggerUtils.logMessage(currentTime + "-");
+
 				lastOldPosition = position;
 
 				/**

### 2017-03-13 spark_project/src/main/java/de/kdml/bigdatalab/spark_and_flink/spark_project/datacorn/TrajectoriesStreamUtils.java

@@ -87,7 +87,7 @@ public class TrajectoriesStreamUtils {
 			PositionMessage trajectory = TrajectoriesUtils.parseDataInput(streamRecord.getValue());
 			trajectory.setStreamedTime(streamRecord.getStreamedTime());
 			trajectory.setNew(true);
-
+			LoggerUtils.logMessage(System.currentTimeMillis() + "-=");
 			// create a tuple of Id & position message for each stream line
 			return new Tuple2<>(trajectory.getID(), trajectory);
 		}).filter(tuple -> {
